### 2021-07-19 paper.tex
Initial commit

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{url}
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{algorithm2e}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{
  What should I notice? Evaluating unexpectedness of events to generate surprising causal candidates.
}
\author{
  \IEEEauthorblockN{
  \IEEEauthorrefmark{1}\IEEEauthorrefmark{2} Étienne Houzé,
  \IEEEauthorrefmark{1} Jean-Louis Dessalles,
  \IEEEauthorrefmark{1} Ada Diaconescu,
  \IEEEauthorrefmark{2} David Menga
  }

  \IEEEauthorblockA{\IEEEauthorrefmark{1}
    Télécom Paris, IP Paris, \emph{Palaiseau}, France\\
  email: \{first\}.\{second\}@telecom-paris.fr}
  \IEEEauthorblockA{\IEEEauthorrefmark{2} EDF R\&{}D, \emph{Palaiseau}, France\\
  email: \{first\}.\{second\}@edf.fr}
}

\maketitle

\begin{abstract}
  When confronted to an unprecedented situation, humans typically show good performance in quickly identifying noticeable past events and proposing them as possible causal hypotheses. Our research focuses on emulating this behaviour into devices, as to deal with rare, unusual situations where previously acquired knowledge fails to provide hypotheses. We rely on a definition of the unexpectedness of events based on Algorithmic Information Theory and the notion of Kolmogorov Complexity. We illustrate our proposal on a simulated smart homes undergoing various unusual events over the course of more than a year.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}

The ability to infer new causal relations is one of the many tasks in which humans prevail but machines still fall short of completing. This ability is prime in making discoveries, or acquiring new knowledge by making educated guesses.

Without any knowledge about how physics or electricty work, it would be possible to infer a possible link between a memorable thunderstorm and a general black-out only by using the noticeability score of the first. While this kind of abductive inference can yield many false-positives, its application can be to propose new hypotheses to be tested out when other possible methods of inference fail to provide any hypothesis (due to a lack of prior similar situations, or to the importance of the context). The difficulty of this approach to reasoning is, however, to quickly identify the relevant candidate hypotheses. In this regards, we may use a basic instinct: without prior knowledge, the most relevant hypothesis might simply be the most memorable recent event. However this definition is highly subjective, and does not seem to fit well into the canvas of computing and AI.

A possible approach would be to rely on the naming complexity of events: memorable events are more likely to be shorter to be named than boring usual events. Think, for instance, of how ``last year's hottest day'' description appears much simpler than ``the 182th day of 7 years ago''. Our idea is to rely on this metrics to estimate the description complexity of events. Then, by comparing the description complexity of a given event $e$ to the average complexity of similar events, we induce a measure of unexpectedness that can be used to assess the most memorable events.

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{figures/filters}
  \caption{An example of retrieving an event through successive predicative filters. The ``hottest day of last year'' is accessed by selecting day-type events, from last year, and looking for the one ranked first by maximum temperature.}
  \label{fig:filters}
\end{figure}

To programatically achieve what humans are easily capable to do, we rely on a basic model intuition of how memory works. We assume that a memory is a unordered set of various distinct objects, upon which can be applied boolean predicates to describe them. These predicates thus provide filters that can be applied to the memory. By applying successive filters, we narrow down the size of the memory, until a single event remains. The complexity of this remaining events can thus be estimated as the simplest combination of predicative filter that uniquely describe it.

The rest of this paper is as follows: we first briefly introduce some notions of Algorithmic Information Theory and its applications in section \ref{sec:theory}, then we explicit our methods of computing the unnexpectedness of past events in cyber-physical systems \ref{sec:computing}. We illustrate this approach by providing an implementation relying on a few predicates on events from a smart home simulation \ref{sec:example}. In section \ref{sec:related} we briefly review other related works using complexity theory or trying to explain smart homes and how our work can be linked with them.

\section{Theoretical Background}
\label{sec:theory}
The notion of complexity formally quantifies the amount of information required for the computation of a binary string (or any object representable by a finite bynary string)\cite{kolmogorov_three_1965,li_introduction_2008}. For such a binary string $s$, its complexity $K(s)$, is the bit length of the shortest program $p$ which, if given as input to a universal Turing Machine $U$, outputs $s$.
\begin{equation}
  K_{U}(s) = \min_{p}\left\{l(p)|U(p)=s\right\}
\end{equation}
The first notable property of this definition is its universality: given that a Turing machine $U'$ can be encoded into $U$ using a program $p_{U}$, we have the following inequality:
\begin{equation}
  K_{U}(s) \le l(p_{u}) + K_{U}(s)
\end{equation}
since the additional constant does not depend on the string $s$, we can drop the requirement on a given machine $U$ and use a universal complexity $K(s)$, such that:
\begin{equation}
  \forall U\in\text{TM}, \forall s, |K(s) - K_{u}(s)| \le C_{U}
\end{equation}
It is important to note that the notion of Kolmogorov complexity is not related to the computation complexity, as there is no requirement on the execution time of the programs, only their length in bits matters for the computation of complexity. In fact, it can be shown that Kolmogorov complexity is not computable\cite{li_introduction_2008}. The sketch of the proof is that computing the Kolmogorov complexity would require to run all programs, up to a certain length, which includes non-terminating ones. As the termination of programs is non-decidable, the computation of complexity is not.

While the introducion of Kolmogorov complexity was purely theoretical, many applications were found, notably because of its close relation with the intuitive notions of the complexity of items, events, situations. For instance,~\cite{murena_solving_2020} studied the use of Kolmogorov Complexity to identify and solve analogies.

Algorithmic Information Theory is also used to define the notion of unexpectedness: in \cite{dessalles2011coincidences}, the unexpectedness $U(e)$ of an event is defined as the difference between an a-priori expected causal complexity $C_{w}(e)$ and the actual observed complexity $C(e)$
\begin{equation}
  \label{eq:unexpected}
  U(e) = |C_{w}(e) - C(e)|
\end{equation}
Using this measure, it is possible to model phenomenons such as coincidences: imagine that you happen to run into a friend in a park. The unnexpectedness will be much larger if this park is in Singapore than if it is in your common neighbourhood: the causal complexity $C_{w}$ is much higher if you a priori have no knowledge of your friend being around, while the description complexity remains the same.

\section{Computing the unexpectedness of events}
\label{sec:computing}
\subsection{A memory model}
We model the memory as an unordered set $M$ of events $e$. Individual events can be seen as data points, augmented with a label indicating their nature (temperature events, failure events, addition/removal events), and a timestamp:
\begin{equation}
  e \equiv (l, t,\mathcal{D})
\end{equation}
where $l$ is the label, $t$ a timestamp and $\mathcal{D}$ a multi-dimensional datapoint representing various characteristics of the events (e.g.~its duration, the maximum temperature reached, the sensor name, its position).

We define \emph{predicates} as boolean function operating on these events, and taking an additional argument, encoded as a finite binary string
\begin{equation}
  \pi : \begin{cases}
    \mathcal{M}\times \{0,1\}^{*} &\mapsto \{0,1\} \\
    (e, k) &\mapsto \pi_{k}(e)
    \end{cases}
\end{equation}
Using this construction allows for structures such as \emph{max\_temperature(e, T)}, where $e$ is the event argument and $T$ a temperature described by the finite binary string.

From these predicates, we can construct filter operations on the memory. A filter operates by simply selecting the events satisfying a given predicate $\pi_{k}$ and constructing a resulting memory from these events.
\begin{equation}
 f_{\pi, k}(\mathcal{M}) = \left\{e \in \mathcal{M} | \pi_{k}(e) \right\}
\end{equation}
For example, the filter ``last year'' corresponds to $\pi = \mathit{date}$ and $k = \mathtt{last year}$.

\subsection{Description complexity of events}

From the previous definitions, we then estimate the complexity of events as the ``simplest'' way of retrieving them from the memory, by applying successive filters. A retrieval path composed of various filters $p = f_{1} = f_{\pi_{1}, k_{1}}, f_{2} = f_{\pi_{2}, k_{2}}, \dots, f_{n} = f_{\pi_{n}, k_{n}}$ is said to \emph{retrieve} an event $e$ from a memory $\mathcal{M}$ if and only if $p(\mathcal{M}) = f_{1} \circ f_{2} \circ \dots \circ f_{n} (\mathcal{M}) = \{e\}$. For an event $e$, if such a path exist, its complexity can be bounded by $K(p) = \sum_{f_{\pi, k} \in p} l(\pi) + l(k)$, the complexity of the retrieval path being the sum of the bit lengths required to describe the predicates $\pi$ and their arguments $k$.

The description complexity of an event $e$ can thus be defined as the minimum complexity of a retrieval path retrieving $e$:

\begin{equation}
  \label{eq:desc_k}
  K(e) = \min_{p | p(\mathcal{M}) = e} K(p) = \min_{p | p(\mathcal{M})=e} \sum_{f_{\pi, k} \in p} l(\pi) + l(k)
\end{equation}

\subsection{Computing an unexpectedness measure}
Equation \ref{eq:unexpected} gives a universal description of unexpectedness, as the difference between the complexity of the phenomenon generating an event and the complexity required to describe this event. In our particular application case, we may directly use the description complexity $K(e)$ as defined in \ref{eq:desc_k}.

\begin{algorithm}
  \caption{Iterative computation of the complexity}
  \label{alg:complex_iter}
  $\mathtt{current_{explore}} \leftarrow [\mathcal{M}]$ \;
  $\mathtt{future_{explore} \leftarrow} [\;]$ \;
  \While{$\mathtt{current_{explore}} \neq [\;]$ \textbf{and} $\mathtt{pass} < \mathtt{max\_pass}$}{
    \For{$(\mathtt{M_{prev}}, \mathtt{K_{prev}}) \in \mathtt{current_{explore}}$}{
      \For{$\mathtt{p \in \mathcal{P}}$}{
        \For{$\mathtt{k \in \{0,1\}^{*}}$}{
          $\mathtt{K_{current} \leftarrow l(p) + l(k) + K_{prev}}$ \;
          \If{$\mathtt{K_{current}} > \mathtt{max_{complex}}$}{
            \textbf{break} \;
          }
          $\mathtt{M'} \leftarrow f_{p,k}(\mathtt{M_{prev}})$ \;
          \eIf{$\mathtt{M'} = \mathtt{\{e\}}$}{
            $\mathtt{K(e) \leftarrow \min(K(e), K_{current})}$ \;
          }{
            $\mathtt{future_{explore}.append((M', K_{current}))}$\;
          }
        }
      }
    }
    $\mathtt{current_{explore}} \leftarrow \mathtt{future_{explore}}$ \;
    $\mathtt{future_{explore}} \leftarrow [\;]$ \;
    $\texttt{pass} \leftarrow \mathtt{pass} + 1$\;
  }
\end{algorithm}

\section{Example}
\label{sec:example}
\subsection{Setup}
In order to provide an acceptable setup to our experiments, we used the scenario of smart homes. This kind of scenario is a prime example of situations where innovative methods of abduction can prove useful, for various reasons: i) as the number of sensors grows with the number of equipped devices in the house, not all recorded events are useful and should be remembered over long period of time; ii) atypical situations, which are the ones where abduction is most likely to be used (to explain situations to the user, or to solve a conflict), are also where the lack of past data makes knowledge acquisition hard.

We evaluate our approach in these two prominent use cases: recognizing memorable events from a wide array of past measures, and finding relevant hypotheses for abductive reasoning in unusual situations. To achieve both these situtations, we rely on simulations provided by a smart home simulator: iCasa \cite{lalanda_self-aware_2017}.

Using this basis, we implemented a scenario spanning over 420 days, and comprising a daily cycle of outdoor weather (temperature and sunlight), as well as user's movements. All these daily changes create non-noticeable events, serving as a background noise for our experiments. To produce outstanding events, we randomly generated around twenty events, spanning over the whole duration of the simulation, of different kinds:
\begin{itemize}
    \item Unusual weather: the outdoor conditions are set to unusually high or low temeperatures.
    \item Heater failures: heater can fail, making them turning off regardless of the command they receive.
    \item User's vacation: the user go out of the building for an extended period of time.
\end{itemize}
\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{figures/simulator}
  \caption{View of the simulator's web interface provided by iCasa. The four rooms are visible, with their equipment and the user.}
  \label{fig:view}
\end{figure}

\begin{figure}[ht]
  \includegraphics[width=\linewidth]{figures/ts_example}
  \caption{Time series data from the simulation: outdoor temperature and controller temperature of a room. We can see two temperature drops, one for each room. What could be the causes of these drops?}
  \label{fig:ts_example}
\end{figure}

The values of all devices and zones variables was regularly monitored throughout the simulation run, and the resulting data, which an excerpt is shown in figure \ref{fig:ts_exaple} can later be used as a basis for our experiments.
\subsection{Implementation}

For the implementation of our method, we first needed to identify and characterize events from the time series data generated by the iCasa simulation. Since this is not the focus point of our present work (see sec. \ref{sec:related}), we simply apply threshold and precomputed conditions based detections to create a base of events.

This base of events consists the basis of the initial memory $\mathcal{M}$ used for computations.

\subsection{Results}

\begin{figure}[ht]
  \centering
  \includegraphics[width=\linewidth]{figures/surprises}
  \caption{Unnexpectedness score for events in the memory}
  \label{fig:result1}
\end{figure}

\section{Related Works}
\label{sec:related}

\section{Perspectives}

\section{Conclusion}

\bibliographystyle{IEEEtran}
\bibliography{biblio.bib}

\end{document}

### 2021-07-21 paper.tex
Working on the paper...

@@
 \usepackage{textcomp}
-\usepackage{algorithm2e}
+\usepackage[ruled,linesnumbered]{algorithm2e}
 \usepackage{xcolor}
@@
 \title{
-  What should I notice? Evaluating unexpectedness of events to generate surprising causal candidates.
+  What should I notice? Evaluating memorability of events to generate surprising causal candidates.
 }
@@
 \begin{abstract}
-  When confronted to an unprecedented situation, humans typically show good performance in quickly identifying noticeable past events and proposing them as possible causal hypotheses.
-Our research focuses on emulating this behaviour into devices, as to deal with rare, unusual situations where previously acquired knowledge fails to provide hypotheses.
-We rely on a definition of the unexpectedness of events based on Algorithmic Information Theory and the notion of Kolmogorov Complexity.
-We illustrate our proposal on a simulated smart homes undergoing various unusual events over the course of more than a year.
+  When confronted to an unprecedented situation, humans typically show good
+  performance in quickly identifying noticeable past events and proposing them
+  as possible causal hypotheses. This kind of abductive inference is widely
+  overlooked in modern AI approaches which rely on massive datasets to learn
+  associative patterns. Our proposal is to formalize and compute a ``memorability''
+  score over a memory of various recorded events from a cyber-physical system.
+ This score can later be used either to select only relevant information to be
+ remembered, or to propose causal hypotheses in unusual situations, on demand.
+ As such, the approach aims at being complementary to more traditional
+ learning-focused techinques. We provide theoretical ground for our approach by
+ using and extending existing results and ideas from Algorithmic Information
+ Theory and provide an implementation example, showing practical results in a
+ smart-home scenario.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-component, formatting, style, styling, insert
+  Complexity, Algorithmic Information Theory, Simplicity, Abduction, Surprise
 \end{IEEEkeywords}
@@
 
-The ability to infer new causal relations is one of the many tasks in which humans prevail but machines still fall short of completing.
-This ability is prime in making discoveries, or acquiring new knowledge by making educated guesses.
+The ability to infer new causal relations is one of the many tasks in which
+humans prevail but machines still fall short of completing. This ability is
+prime in making discoveries, or acquiring new knowledge by making educated
+guesses.
 
-Without any knowledge about how physics or electricty work, it would be possible to infer a possible link between a memorable thunderstorm and a general black-out only by using the noticeability score of the first.
-While this kind of abductive inference can yield many false-positives, its application can be to propose new hypotheses to be tested out when other possible methods of inference fail to provide any hypothesis (due to a lack of prior similar situations, or to the importance of the context).
-The difficulty of this approach to reasoning is, however, to quickly identify the relevant candidate hypotheses.
-In this regards, we may use a basic instinct: without prior knowledge, the most relevant hypothesis might simply be the most memorable recent event.
-However this definition is highly subjective, and does not seem to fit well into the canvas of computing and AI.
+Without any knowledge about how physics or electricty work, it would be possible
+to infer a possible link between a memorable thunderstorm and a general
+black-out only by using the noticeability score of the first. While this kind of
+abductive inference can yield many false-positives, its application can be to
+propose new hypotheses to be tested out when other possible methods of inference
+fail to provide any hypothesis (due to a lack of prior similar situations, or to
+the importance of the context). The difficulty of this approach to reasoning is,
+however, to quickly identify the relevant candidate hypotheses. In this regards,
+we may use a basic instinct: without prior knowledge, the most relevant
+hypothesis might simply be the most memorable recent event. However this
+definition is highly subjective, and does not seem to fit well into the canvas
+of computing and AI.
 
-A possible approach would be to rely on the naming complexity of events: memorable events are more likely to be shorter to be named than boring usual events.
-Think, for instance, of how ``last year's hottest day'' description appears much simpler than ``the 182th day of 7 years ago''.
-Our idea is to rely on this metrics to estimate the description complexity of events.
-Then, by comparing the description complexity of a given event $e$ to the average complexity of similar events, we induce a measure of unexpectedness that can be used to assess the most memorable events.
+The difficulty of this computation comes from different factors. First, events
+can be of different nature, and not directly comparable. In fact, for the use
+case of smart homes, or other cyber physical systems, events range from device
+removal to a detected presented or an unusually high temperature. How can one
+can then assess which ones are more memorable? Furhtermore, even for comparable
+events, different characterics can be put forward to argue for the most
+memorable: is a record-high temperature 47 days ago more memorable than the
+small deviation recorded just 3 minutes ago? To this day, no current system
+proposes to use all this information from various event types from different
+devices to compute a unified metrics of ``memorability''.
+
+% The difficulty of doing so comes from the wide variety of events, parameters and
+% characteristics to take into account. How can one assess if a small recent
+% perturbation is more important than a larger but older one? In addition, one
+% needs a common metrics to evaluate the relative importance of data coming from
+% various devices. In the case of a smart home, for instance, devices range from
+% presence sensors to humidity and temperature sensors, power meters, etc. To this
+% day, no current system proposes to analyze all events coming from such various
+% devices and come to a unified metrics of ``memorability''.
+%
+
+To tackle this challenge, we propose to refer to events using a technique used
+by humans: naming. It appears that humans are able to name events and describe
+them, regardless of their nature or their characteristics. Furthermore, there is
+a noticeable correlation between the lenght of the description, in terms of the
+complexity of concepts and arguments, and the ``memorability'' of the event.
+Think, for instance, of how ``last year's hottest day'' appears simpler than
+``the 182th day of 7 years ago''. Using this method, the day described by the
+first sentence would be deemed simpler than the target of the second
+description. By comparing the description length of the event with the average
+description lenght of similar events, one would then have a measure of the
+``memorability'' of events, as memorable events stands out as much simpler (or
+more complex) than their neighbours.
+
+% A possible approach would be to rely on the naming complexity of events:
+% memorable events are more likely to be shorter to be named than boring usual
+% events. Think, for instance, of how ``last year's hottest day'' description
+% appears much simpler than ``the 182th day of 7 years
+% ago''\cite{robles_applications_2010}. Our idea is to rely on this metrics to
+% estimate the description complexity of events. Then, by comparing the
+% description complexity of a given event $e$ to the average complexity of similar
+% events, we induce a measure of unexpectedness that can be used to assess the
+% most memorable events.
 
@@
   \includegraphics[width=\linewidth]{figures/filters}
-  \caption{An example of retrieving an event through successive predicative filters.
-The ``hottest day of last year'' is accessed by selecting day-type events, from last year, and looking for the one ranked first by maximum temperature.}
+  \caption{An example of retrieving an event through successive predicative
+    filters. From the base memory (yellow), successive filters select the events
+    satisfying their predicates (grey arrows). The operations yields successive
+    subsets of the memory (yellow), until a final singleton memory is reach,
+    meaning that the successive filters retrieve a unique event from the memory.
+    The predicates used for this retrival path and their arguments,
+    $\pi_{1}, \pi_{2}, \pi_{3}, k_{1}, k_{2}, k_{3}$, can be used as a retrieval
+    program. Its length gives an upper bound to the description complexity of
+    the event $e_{h}$: ``last year's hottest day''.}
   \label{fig:filters}
@@
 
-To programatically achieve what humans are easily capable to do, we rely on a basic model intuition of how memory works.
-We assume that a memory is a unordered set of various distinct objects, upon which can be applied boolean predicates to describe them.
-These predicates thus provide filters that can be applied to the memory.
-By applying successive filters, we narrow down the size of the memory, until a single event remains.
-The complexity of this remaining events can thus be estimated as the simplest combination of predicative filter that uniquely describe it.
+To programatically achieve what humans are easily capable to do, we rely on a
+basic model intuition of how memory works. Figure \ref{fig:filters} shows this
+process for the example description ``last year's hottest day''. We assume that
+a memory $\mathcal{M}$ is a unordered set of various distinct objects, upon
+which can be applied boolean predicates $\pi$ to describe them. These predicates thus
+provide filters $f_{\pi}$ that can be applied to the memory, selecting all events
+satifying predicate $\pi$. By applying successive
+filters, we narrow down the size of the memory, until a single event remains.
+The complexity of this remaining events can thus be estimated as the simplest
+combination of predicative filters used to retrieve it.
 
-The rest of this paper is as follows: we first briefly introduce some notions of Algorithmic Information Theory and its applications in section \ref{sec:theory}, then we explicit our methods of computing the unnexpectedness of past events in cyber-physical systems \ref{sec:computing}.
-We illustrate this approach by providing an implementation relying on a few predicates on events from a smart home simulation \ref{sec:example}.
-In section \ref{sec:related} we briefly review other related works using complexity theory or trying to explain smart homes and how our work can be linked with them.
+The rest of this paper is organized as follows: we first briefly introduce some notions of
+Algorithmic Information Theory and its applications in section~\ref{sec:theory},
+then we explicit our methods of computing the unnexpectedness of past events in
+cyber-physical systems~\ref{sec:computing}. We illustrate this approach by
+providing an implementation relying on a few predicates on events from a smart
+home simulation~(section \ref{sec:example}). In section~\ref{sec:related} we briefly
+review other related works using complexity theory or trying to explain smart
+homes and how our work can be linked with them before exploring potential
+extensions of our work in section \ref{sec:future}
 
@@
 \label{sec:theory}
-The notion of complexity formally quantifies the amount of information required for the computation of a binary string (or any object representable by a finite bynary string)\cite{kolmogorov_three_1965,li_introduction_2008}.
-For such a binary string $s$, its complexity $K(s)$, is the bit length of the shortest program $p$ which, if given as input to a universal Turing Machine $U$, outputs $s$.
+Kolmorov complexity formally quantifies the amount of information required
+for the computation of a finite binary string\footnote{While the defintion still
+holds for some inifinite binary strings (think of the representation of the
+decimals of $\pi$), we restrict ourselves to finite strings in this paper.} (or
+any object represented by a finite bynary
+string)\cite{kolmogorov_three_1965,li_introduction_2008}. For such a binary
+string $s$, its complexity $K(s)$, is the bit length of the shortest program $p$
+which, if given as input to a universal Turing Machine $U$, outputs $s$.
 \begin{equation}
+  \label{eq:kolmo_def}
   K_{U}(s) = \min_{p}\left\{l(p)|U(p)=s\right\}
 \end{equation}
-The first notable property of this definition is its universality: given that a Turing machine $U'$ can be encoded into $U$ using a program $p_{U}$, we have the following inequality:
+The first notable property of this definition is its universality: while the
+choice of the Turing machine $U$ used for the computations appears in the
+definition of eq. \ref{eq:kolmo_def}, all results stand, up to an additional
+constant. Think, for instance, how any Turing-complete programming language can
+be turned into any other language, using a compiler program. In fact, since any
+Turing machine $U'$ can be emulated by $U$ from a
+finite program $p_{U}$, we have the following inequality:
 \begin{equation}
+  \label{eq:inequality_univ}
   K_{U}(s) \le l(p_{u}) + K_{U}(s)
 \end{equation}
-since the additional constant does not depend on the string $s$, we can drop the requirement on a given machine $U$ and use a universal complexity $K(s)$, such that:
+From this first result, we can then define a universal complexity $K(s)$, which
+no longer depends on the choice of a Turing machine $U$, such that, for any
+machine $U$,
 \begin{equation}
@@
 \end{equation}
-It is important to note that the notion of Kolmogorov complexity is not related to the computation complexity, as there is no requirement on the execution time of the programs, only their length in bits matters for the computation of complexity.
-In fact, it can be shown that Kolmogorov complexity is not computable\cite{li_introduction_2008}.
-The sketch of the proof is that computing the Kolmogorov complexity would require to run all programs, up to a certain length, which includes non-terminating ones.
-As the termination of programs is non-decidable, the computation of complexity is not.
+where the additional constant $C_{U}$ does not depend on the object $s$.
 
-While the introducion of Kolmogorov complexity was purely theoretical, many applications were found, notably because of its close relation with the intuitive notions of the complexity of items, events, situations.
-For instance,~\cite{murena_solving_2020} studied the use of Kolmogorov Complexity to identify and solve analogies.
+It is important to note that the notion of Kolmogorov complexity is not related
+to the computation complexity, as there is no requirement on the execution time
+of the programs, only their length in bits matters for the computation of
+complexity. In fact, it can be shown that Kolmogorov complexity is not
+computable\cite{li_introduction_2008}. The sketch of the proof is the following:
+since the definition of complexity from eq. \ref{eq:kolm_def} requires to find
+the shortest of all programs on a given Turing machine outputting $s$, it
+requires to run all programs up to a given length and compare their result to
+$s$. However, being able to do so means that we would be able to tell whether
+these programs terminate. Since the termination of programs is famously
+undecidable, by extension, the computation of Kolmogorov complexity is
+impossible. One can however easily approximate it with upper bounds, by
+exhibiting a program outputting $s$.
 
-Algorithmic Information Theory is also used to define the notion of unexpectedness: in \cite{dessalles2011coincidences}, the unexpectedness $U(e)$ of an event is defined as the difference between an a-priori expected causal complexity $C_{w}(e)$ and the actual observed complexity $C(e)$
+While the introducion of Kolmogorov complexity was purely theoretical, many
+applications were found, notably because of its close relation with the
+intuitive notions of the complexity of items, events, situations. For
+instance,~\cite{murena_solving_2020} studied the use of Kolmogorov Complexity to
+identify and solve analogies.
+
+Algorithmic Information Theory is also used to define the notion of
+unexpectedness: in \cite{dessalles2011coincidences}, the unexpectedness $U(e)$
+of an event is defined as the difference between an a-priori expected causal
+complexity $C_{w}(e)$ and the actual observed complexity $C(e)$.
 \begin{equation}
@@
 \end{equation}
-Using this measure, it is possible to model phenomenons such as coincidences: imagine that you happen to run into a friend in a park.
-The unnexpectedness will be much larger if this park is in Singapore than if it is in your common neighbourhood: the causal complexity $C_{w}$ is much higher if you a priori have no knowledge of your friend being around, while the description complexity remains the same.
+
+This result comes from the understanding that, while Kolmogorov complexity is
+computed with Turing machine, it can be used as an elegant proxy for modeling
+information processing in the human brain, and thus helps design a notion of
+simplicty or complexity of events\cite{}.
+
+Using this measure, it is possible to model phenomenons such as coincidences:
+imagine that you happen to run into a friend in a park. The unnexpectedness will
+be much larger if this park is in Singapore than if it is in your common
+neighbourhood: the causal complexity $C_{w}$ is much higher if you a priori have
+no knowledge of your friend being around, while the description complexity
+remains the same.
 
@@
 \subsection{A memory model}
-We model the memory as an unordered set $M$ of events $e$.
-Individual events can be seen as data points, augmented with a label indicating their nature (temperature events, failure events, addition/removal events), and a timestamp:
+We model the memory as an unordered set $M$ of events $e$. Individual events can
+be seen as data points, augmented with a label indicating their nature
+(temperature events, failure events, addition/removal events), and a timestamp:
 \begin{equation}
+  \label{eq:event}
   e \equiv (l, t,\mathcal{D})
 \end{equation}
-where $l$ is the label, $t$ a timestamp and $\mathcal{D}$ a multi-dimensional datapoint representing various characteristics of the events (e.g.~its duration, the maximum temperature reached, the sensor name, its position).
+where $l$ is the label, $t$ a timestamp and $\mathcal{D}$ a multi-dimensional
+datapoint representing various characteristics of the events (e.g.~its duration,
+the maximum temperature reached, the sensor name, its position).
 
-We define \emph{predicates} as boolean function operating on these events, and taking an additional argument, encoded as a finite binary string
+We define \emph{predicates} as boolean function operating on these events, and
+taking an additional argument, encoded as a finite binary string:
 \begin{equation}
+  \label{eq:predicate}
   \pi : \begin{cases}
@@
 \end{equation}
-Using this construction allows for structures such as \emph{max\_temperature(e, T)}, where $e$ is the event argument and $T$ a temperature described by the finite binary string.
+An example of predicate is to use $\pi = \mathtt{year}$ and $k=1$, thus
+constructing the predicate $\mathtt{year()}e, Y\mathtt{)}$, which tells whether
+the event $e$ occurred $Y$ years ago.
 
-From these predicates, we can construct filter operations on the memory.
-A filter operates by simply selecting the events satisfying a given predicate $\pi_{k}$ and constructing a resulting memory from these events.
+From these predicates, we can construct filter operations on the memory. A
+filter operates by selecting the events satisfying a given predicate
+$\pi_{k}$ and constructing a resulting memory from these events.
 \begin{equation}
+  \label{eq:filter}
  f_{\pi, k}(\mathcal{M}) = \left\{e \in \mathcal{M} | \pi_{k}(e) \right\}
 \end{equation}
-For example, the filter ``last year'' corresponds to $\pi = \mathit{date}$ and $k = \mathtt{last year}$.
+For instance, using the same $\pi = \mathtt{year}$ and $k=1$ as above, we would
+build the filter $f_{\pi, k} = \mathtt{last\_{}year}$, which selects all events
+that occurred last year.
 
@@
 
-From the previous definitions, we then estimate the complexity of events as the ``simplest'' way of retrieving them from the memory, by applying successive filters.
-A retrieval path composed of various filters $p = f_{1} = f_{\pi_{1}, k_{1}}, f_{2} = f_{\pi_{2}, k_{2}}, \dots, f_{n} = f_{\pi_{n}, k_{n}}$ is said to \emph{retrieve} an event $e$ from a memory $\mathcal{M}$ if and only if $p(\mathcal{M}) = f_{1} \circ f_{2} \circ \dots \circ f_{n} (\mathcal{M}) = \{e\}$.
-For an event $e$, if such a path exist, its complexity can be bounded by $K(p) = \sum_{f_{\pi, k} \in p} l(\pi) + l(k)$, the complexity of the retrieval path being the sum of the bit lengths required to describe the predicates $\pi$ and their arguments $k$.
+From the previous definitions, we then estimate the complexity of events as the
+``simplest'' way of retrieving them from the memory, by applying successive
+filters. A retrieval path composed of various filters
+$p = f_{1} = f_{\pi_{1}, k_{1}}, f_{2} = f_{\pi_{2}, k_{2}}, \dots, f_{n} = f_{\pi_{n}, k_{n}}$
+is said to \emph{retrieve} an event $e$ from a memory $\mathcal{M}$ if and only
+if
+$p(\mathcal{M}) = f_{1} \circ f_{2} \circ \dots \circ f_{n} (\mathcal{M}) = \{e\}$.
+For an event $e$, if such a path exist, its complexity can be bounded by
+$K(p) = \sum_{f_{\pi, k} \in p} l(\pi) + l(k)$, the complexity of the retrieval
+path being the sum of the bit lengths required to describe the predicates $\pi$
+and their arguments $k$.
 
@@
 
-\subsection{Computing an unexpectedness measure}
-Equation \ref{eq:unexpected} gives a universal description of unexpectedness, as the difference between the complexity of the phenomenon generating an event and the complexity required to describe this event.
-In our particular application case, we may directly use the description complexity $K(e)$ as defined in \ref{eq:desc_k}.
-
 \begin{algorithm}
-  \caption{Iterative computation of the complexity}
-  \label{alg:complex_iter}
   $\mathtt{current_{explore}} \leftarrow [\mathcal{M}]$ \;
   $\mathtt{future_{explore} \leftarrow} [\;]$ \;
+  $\mathtt{pass} \leftarrow 0$ \;
   \While{$\mathtt{current_{explore}} \neq [\;]$ \textbf{and} $\mathtt{pass} < \mathtt{max\_pass}$}{
@@
   }
+  \caption{Iterative computation of the complexity}
+  \label{alg:complex_iter}
 \end{algorithm}
 
-\section{Example}
+The definition used for description complexity in eq. \ref{eq:desc_k} allows for
+a direct implemenation, which is shown in alg. \ref{alg:complex_iter}. This
+algorithm operates iteratively: starting with the base memory $\mathcal{M}$
+(line 1), we apply all possible predicate concepts $\pi$ and programs $k$ (lines
+6-7), up to a given length $\mathtt{max\_len}$ bits, and apply them:
+$M' = f_{\pi, k}(M)$ (line 12). We then store the pairs
+$(M', \mathtt{len(}\pi, k\mathtt{)})$ in an array $\mathtt{future_{explore}}$. At
+the end of the iteration, the results of the filters become the memories which
+will be explored during the next iteration (lines 21--23). Each pass thus explore
+retrieval paths of increasing length. When a singleton memory is reached, the
+complexity of its unique element is upper-bounded with the length of the
+corresponding retrieval path (line 14).
+
+\subsection{Computing an unexpectedness measure}
+Equation \ref{eq:unexpected} gives a universal definition of unexpectedness, as
+the difference between the complexity of the phenomenon generating an event and
+the complexity required to describe this event. In our particular application
+case, we may directly use the description complexity $K(e)$ as defined in
+\ref{eq:desc_k}.
+
+The term $C_{w}(e)$ from eq. \ref{eq:unexpected} is harder to formally
+fit into our canvas. As $C_{w}$ is defined as the shortest program a
+``W-machine'' replicating the world's behaviour, based on the person's
+knowledge\cite{dessalles2011coincidences}, it is convenient to understand this
+term as the expectation one would have regarding the complexity of the event
+$e$, \emph{before} actually evaluating it.
+
+A possible proxy for this quantity is to compute the average complexity
+of \emph{similar} events. Still, difficulty remains in the definition of what
+can be considered \emph{similar} events. Given that we have already used the
+notion of predicates to describe the events and compute their description
+complexity, we can use them as simple metrics for similarity: for each predicate
+$\pi$ satisfied by $e$, we can define a set of similar events $N_{\pi}(e)$ as
+all the events also satisfying $\pi$.
+
+\begin{equation}
+\label{eq:similar}
+N_{\pi}(e) = \{e'\in \mathcal{M}, \pi(e') \wedge e' \neq e\}
+\end{equation}
+
+Now, when considering, for all possible predicates $\pi$, the corresponding
+neighbourhoods $N_{\pi}(e)$, with the convention that $N_{\pi}(e) = \emptyset$
+if $\pi(e) = \bot$, we can compute an average expected complexity for $e$:
+
+\begin{equation}
+\label{eq:expected}
+K_{exp}(e) = \frac{
+                    \sum_{\pi} \sum_{e' \in N_{\pi}(e)} K(e')
+                  }{
+                    \sum_{\pi} |N_{\pi}(e)|
+                  }
+\end{equation}
+
+The definition used for the expected complexity of the event $e$ in eq.
+\ref{eq:expected} holds the intuition of favorizing more similar events: if $e'$
+is \emph{very} similar to $e$, it will appear in many neighbouhoods, since it
+satisfies mostly the same predicates as $e$. Therefore, its complexity term will
+be present in more terms in eq. \ref{eq:expected}, and will weight more towards
+the final result.
+
+\subsection{From memorability to abduction}
+% On peut étendre la définition utilisée pour la surprise en faisant intervenir
+% les prédicats'' `sympathiques` par rapport à ce que l'on questionne.
+The problem of the abductive inference is different from the basic computation
+of the memorability score. Here, \emph{knowing} that we want to find a cause $c$
+to an observed effect $e$, we try to find the most remarkable event in past
+memory in that regards. While our ``memorability'' score aims to identify the
+most remarkable events, it does not take into account the added knowledge about
+the effect $e$ that is available for abductive inference.
+
+This added knowledge would however influence the unexpectedness of a possible
+cause $c$, from eq. \ref{eq:unexpected}: even though the causal complexity
+$C_{W}(c)$ is likely to remain unchanged, as it relies on the model of the world in
+the person's mind, the description complexity $C(c)$ is replaced with the
+conditional complexity $C(c|e)$, as the knowledge of $e$ is included in the
+problem's data.
+
+In our canvas of predicative filtering, this corresponds to adding for free, in
+terms of program length, information about $e$ to predicates. This translates
+into considering relative order between events: when investigating a cause for
+an anomaly in the living-rooms, other anomalies occurring in the same
+living-room will be simpler, since the location ``living-room'' is already known
+from knowing the consequnce. In terms of predicates, we therefore append a
+description of $e$ to all programs $k$ of predicates:
+$\pi_{k}(c) = \pi_{k::e}(c)$, where $::$ is the append operation.
+
+This leads to another score $Abd(c,e)$ (abduction score of effect $c$ for cause
+$e$), which is defined similarly to the description complexity in eq.
+\ref{eq:desc_k} and using relative filters.
+
+\begin{equation}
+  \label{eq:abd_k}
+  Abd(c) = \min_{p | p(\mathcal{M}|e) = c} K(p) = \min_{p | p(\mathcal{M}|e)=c} \sum_{f_{\pi, k} \in p} l(\pi) + l(k)
+\end{equation}
+This abduction score translates the additional information provided to the
+system when answering a user's request. As such, it encapsulates the idea
+presented as the motivation of this paper: when confronted to a surprising
+situation, and with no other source of information, it can be used as a metrics
+to find relevant posisble surprising causes to the observed consequence.
+
+\section{Implementation Example}
 \label{sec:example}
 \subsection{Setup}
-In order to provide an acceptable setup to our experiments, we used the scenario of smart homes.
-This kind of scenario is a prime example of situations where innovative methods of abduction can prove useful, for various reasons: i) as the number of sensors grows with the number of equipped devices in the house, not all recorded events are useful and should be remembered over long period of time; ii) atypical situations, which are the ones where abduction is most likely to be used (to explain situations to the user, or to solve a conflict), are also where the lack of past data makes knowledge acquisition hard.
+To test our approach, we set up an experimental smart home setup. This choice of
+configuration is motivated by the challenges posed by smart homes for abductive
+inference: i) as the number of connected devices increases, more events are
+recorded, making the detection of memorable events more important; ii) smart
+homes are prone to undergo atypical situations, highly dependant on the context,
+for which pre-established relations might fail to find good abduction
+candidates. Furthermore, the choice was also motivated by the ease of application: as
+previous works exist on smart homes, it is possible to simulate their behaviour
+and quickly generate data to extract events from and test our methods.
 
-We evaluate our approach in these two prominent use cases: recognizing memorable events from a wide array of past measures, and finding relevant hypotheses for abductive reasoning in unusual situations.
-To achieve both these situtations, we rely on simulations provided by a smart home simulator: iCasa \cite{lalanda_self-aware_2017}.
+% In order to provide an acceptable setup to our experiments, we used the scenario
+% of smart homes. This kind of scenario is a prime example of situations where
+% innovative methods of abduction can prove useful, for various reasons: i) as the
+% number of sensors grows with the number of equipped devices in the house, not
+% all recorded events are useful and should be remembered over long period of
+% time; ii) atypical situations, which are the ones where abduction is most likely
+% to be used (to explain situations to the user, or to solve a conflict), are also
+% where the lack of past data makes knowledge acquisition hard.
 
-Using this basis, we implemented a scenario spanning over 420 days, and comprising a daily cycle of outdoor weather (temperature and sunlight), as well as user's movements.
-All these daily changes create non-noticeable events, serving as a background noise for our experiments.
-To produce outstanding events, we randomly generated around twenty events, spanning over the whole duration of the simulation, of different kinds:
-\begin{itemize}
-    \item Unusual weather: the outdoor conditions are set to unusually high or low temeperatures.
-    \item Heater failures: heater can fail, making them turning off regardless of the command they receive.
-    \item User's vacation: the user go out of the building for an extended period of time.
-\end{itemize}
+To carry out the simulation, we built custom modules into the existing iCasa
+smart home simulator\cite{lalanda_self-aware_2017}. This simulation platform
+allows for simulating autonomic systems, handling internal communications,
+injection of new components at runtime, deletion or change of existing
+components. As such, we used a basic scenario consisting of a house of 4 rooms,
+a single user, and an outdoor zone. All four rooms of the house are equipped
+with a temperature controller system, monitoring and controlling heaters (fig.
+\ref{fig:view}).
+
 \begin{figure}[ht]
@@
   \includegraphics[width=\linewidth]{figures/simulator}
-  \caption{View of the simulator's web interface provided by iCasa. The four rooms are visible, with their equipment and the user.}
+  \caption{View of the simulator's web interface provided by iCasa. The four
+    rooms are visible, with their equipment and the user.}
   \label{fig:view}
@@
 
+Using this basis, we implemented a scenario spanning over 420 days, and
+comprising a daily cycle of outdoor weather (temperature and sunlight), as well
+as user's movements. All these daily changes create non-noticeable events,
+serving as a background noise for our experiments. To produce outstanding
+events, we randomly generated around twenty events, spanning over the whole
+duration of the simulation, of different kinds:
+
+\begin{itemize}
+    \item Unusual weather: the outdoor conditions are set to unusually high or
+        low temeperatures.
+    \item Heater failures: heater can fail, making them turning off regardless
+        of the command they receive.
+    \item User's vacation: the user go out of the building for an extended
+        period of time.
+    \item Device removal/addition: a device is removed, or another one is added
+        to the system.
+\end{itemize}
+
 \begin{figure}[ht]
   \includegraphics[width=\linewidth]{figures/ts_example}
-  \caption{Time series data from the simulation: outdoor temperature and controller temperature of a room.
-We can see two temperature drops, one for each room.
-What could be the causes of these drops?}
+  \caption{Time series data from the simulation: outdoor temperature and
+    controller temperature of a room. We can see two temperature drops, one for
+    each room. What could be the causes of these drops?}
   \label{fig:ts_example}
@@
 
-The values of all devices and zones variables was regularly monitored throughout the simulation run, and the resulting data, which an excerpt is shown in figure \ref{fig:ts_exaple} can later be used as a basis for our experiments.
-\subsection{Implementation}
+The values of all devices and zones variables was regularly monitored throughout
+the simulation run, and the resulting data, which an excerpt is shown in figure
+\ref{fig:ts_example} can later be used as a basis for our experiments.
 
-For the implementation of our method, we first needed to identify and characterize events from the time series data generated by the iCasa simulation.
+\subsection{Implementing the complexity computation}
+
+For the implementation of our method, we first needed to identify and
+characterize events from the time series data generated by the iCasa simulation.
 Since this is not the focus point of our present work (see sec.
-\ref{sec:related}), we simply apply threshold and precomputed conditions based detections to create a base of events.
+\ref{sec:related}), we simply apply threshold and precomputed conditions based
+detections to create a base of events.
 
-This base of events consists the basis of the initial memory $\mathcal{M}$ used for computations.
+This base of events consists the basis of the initial memory $\mathcal{M}$ used
+for computations.
+We then implemented a small number of predicate concepts:
+\begin{itemize}
+        \item \texttt{label(e, k)}: whether the event $e$ has the label ranked
+        $k^{th}$ in the memory.
+        \item \texttt{rank(e, k)}: whether the event $e$ has the rank $k_{1}$
+        along axis $k_{2}$, where $k$ is a unique encoding for $(k_{1}, k_{2})$.
+        \item \texttt{day(e, k)}: whether the event $e$ occured $k$ days ago.
+        \item \texttt{month(e, k)}: whether the event $e$ occured $k$ months
+        ago.
+        \item \texttt{location(e, k)}: whether the event $e$ occured in zone
+        $k$.
+\end{itemize}
+
+With a straightforward implemention of memory, predicates and filters, the
+implementation of alg. \ref{alg:complex_iter} worked, but, as expected, took too
+long to be usable in realistic scenarios with hundreds or thousands of events to
+consider. In order to facilitate and speed up computaitons, we also implemented
+the following improvements:
+\begin{itemize}
+  \item The memory object was augmented with various built-in rankings, allowing
+  for faster operations during future filtering. For instance, since the memory
+  object keeps a mapping from timestamps to events, which allows to quickly
+  filter by date without having to loop over each stored element. This mapping,
+  however, is not directly used to retrieve an event from the outside, as to
+  preserve the theoretical model of memory as an unordered set presented in
+  section \ref{sec:computing}.
+
+  \item Each of these predicates holds the property that, in addition to
+  \texttt{True} and \texttt{False}, they can return another value,
+  \texttt{None}, which is theoretically treated as \texttt{False} but carries
+  the additional information that this predicate concept will also be false for
+  any other element of the memory for any subsequent program $k$. This allows to
+  effectively break the innermost loop in alg. \ref{alg:complex_iter}.
+
+  \item Some of the filters, for instance date or rank filters, were
+  hard-written to select from the pre-computed mappings of the memory objects
+  rather than testing a predicate over all memory elements.
+\end{itemize}
 
@@
 
+\subsubsection{Memorability}
+
+The results of the descritpion complexity evaluation, for the described setup,
+are shown in fig. \ref{fig:computed_cplx}. The entire computation, over 4
+iterations (meaning that retrieval paths contained at most 4 filters), took
+around 30 seconds on a modern commercial laptop with an i7-8700u CPU.
+
+A main sequence of ``usual'' events, which complexity is roughly a logarithm of
+the elapsed time since their occurence, is visible. This corresponds to events
+for which the best retrieval path consists of a time description (e.g. ``2
+months and 12 days ago''). On the other hand, some events stand out in terms of
+complexity: some appear simpler, as they can be distinguished by using their
+rank alongside an axis (``the hottest day'', ``the second longest user's
+absence''), or the rare occurence of their kind (``the only fault on the heater'').
+
 \begin{figure}[ht]
   \centering
-  \includegraphics[width=\linewidth]{figures/surprises}
+  \includegraphics[width=\linewidth]{figures/complexities_computed}
+  \caption{The computed complexities of events with at most retrieval paths of
+    length at most 4. Events of type ``day'' (blue), ``hot'' (green), ``cold'' (red) are shown.}
+  \label{fig:computed_cplx}
+\end{figure}
+
+Computing the ``memorability score'', which is shown in fig. \ref{fig:result1},
+highlights these aforementioned events from the main ``usual'' sequence. Since
+the computation of this measure treats unusually complex or simple events the
+same way (from the absolute value operation in eq. \ref{eq:unexpected}), we
+observe that some events are memorabale due to their context only. For instance,
+a temperature anomaly occurring simultaneously to many other anomalies is
+notable, as it is costlier than expected to distinguish it from its neighbours.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\linewidth]{figures/complexities_surprises}
   \caption{Unnexpectedness score for events in the memory}
@@
 
+Given that we generated the data used for this experiment, it is possible to
+flag all perturbation events from the usual daily events and evaluate how a
+detection based on ``memorability'' score would succeed in distinguishing these
+events. The result is presented as a ROC curve in fig. \ref{fig:roc}.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\linewidth]{example-image-a}
+  \caption{Experimental ROC curve (True Positive Rate against False Positive Rate) for a
+    classifier based on our memorability score.}
+  \label{fig:roc}
+\end{figure}
+\subsubsection{Abduction}
+
+As an illustration of the abductive inference possible by using the memorability
+score, we used as a targed consequence the temperature drop visible in the raw
+time series data from our scenario (fig \ref{fig:ts_example}).
+
+% TODO
+\textbf{TODO}
+
 \section{Related Works}
 \label{sec:related}
+\textbf{TODO}
 
 \section{Perspectives}
+\label{sec:future}
+The main purpose of the present paper is to show the possible connections
+between exising definitions of simplicity from cognitive science, Algorithmic
+Information Theory and a practical use case in cyber-physical systems. As such,
+many further improvements can be done to pave the way towards a better
+integration and performance for anomaly detection or abduction.
+
+First, the main limitation of the current approach is the requirement of
+predefined predicate concepts, from which the different filters are constructed.
+As an extension, we suggest that in the future, we could explore online
+generation of such predicate. A posisbility would be to analyze discriminating
+dimensions of incoming data and create predicate as to name these differences,
+similar to the contrast operations proposed in \cite{dessalles_conceptual_2015,
+  gardenfors2004conceptual}. For instance, the predicate concept $\mathtt{hot}$
+can be discovered by discriminating a recent hot day along the
+temperature axis and naming the difference with the prototypical day.
+
+While the execution time is not part of the theoretical view of complexity, it
+is of prime importance for practical applications, especially when one considers
+implementation into real-time systems or embedded devices. While the computation
+we propose appear to be heavy, and possibly heavier as the number of allowed
+predicates grows, significant time savings can be achieve by trimming the base
+memory of past events deemed the most ``non memorable''. For instance, one can
+only retains the 100 most memorable events from the past. The difficulty with
+this approach is that such operations should be done in a manner to not
+interfere with the complexity computations for new elements: by forgetting some
+past events, even uninteresting ones, one should make sure to keep track of what
+made the interesting ones, interesting. Investigation of how to do so can pave
+the way towards practical implementations and dynamic selection of interesting
+events and help reducing the memory and computation cost of data-driven applications.
 

### 2021-07-26 paper.tex
Working on the document

@@
  As such, the approach aims at being complementary to more traditional
- learning-focused techinques. We provide theoretical ground for our approach by
+ learning-focused techniques. We provide theoretical ground for our approach by
  using and extending existing results and ideas from Algorithmic Information
@@
 
-The ability to infer new causal relations is one of the many tasks in which
-humans prevail but machines still fall short of completing. This ability is
-prime in making discoveries, or acquiring new knowledge by making educated
-guesses.
+As the number of connected devices and sensors grows, so does the amount of
+produced data. As a consequence, the ability to filter out a few events from past recordings
+as ``memorable'' becomes prime. While this task is often overlooked by many
+statistical approaches, which instead benefit from large quantity of
+information, it is part of the learning process for humans
+\cite{dehaene_how_2020}. Faced with new situations, humans are prone to infer
+previously seen unusual events as possible causes, and carry out further testing
+to assess the existence of a causal relation.
 
-Without any knowledge about how physics or electricty work, it would be possible
-to infer a possible link between a memorable thunderstorm and a general
-black-out only by using the noticeability score of the first. While this kind of
-abductive inference can yield many false-positives, its application can be to
-propose new hypotheses to be tested out when other possible methods of inference
-fail to provide any hypothesis (due to a lack of prior similar situations, or to
-the importance of the context). The difficulty of this approach to reasoning is,
-however, to quickly identify the relevant candidate hypotheses. In this regards,
-we may use a basic instinct: without prior knowledge, the most relevant
-hypothesis might simply be the most memorable recent event. However this
-definition is highly subjective, and does not seem to fit well into the canvas
-of computing and AI.
+For instance, it is possible, for humans without any prior knowledge about
+physics or electrical engineering, to infer a memorable thunderstorm as a
+possible cause for a general black-out just by using the memorability of
+the former. While this kind of abductive inference (i.e. inference to the cause
+of an observation \cite{magnani_abduction_2011}) can yield many
+false-positives, it can be considered in situations where the lack of knowledge
+from previous occurrences penalizes other approaches.
 
-The difficulty of this computation comes from different factors. First, events
-can be of different nature, and not directly comparable. In fact, for the use
-case of smart homes, or other cyber physical systems, events range from device
-removal to a detected presented or an unusually high temperature. How can one
-can then assess which ones are more memorable? Furhtermore, even for comparable
-events, different characterics can be put forward to argue for the most
-memorable: is a record-high temperature 47 days ago more memorable than the
-small deviation recorded just 3 minutes ago? To this day, no current system
-proposes to use all this information from various event types from different
-devices to compute a unified metrics of ``memorability''.
+% The difficulty of this approach to reasoning is, however, to quickly identify
+% the relevant candidate hypotheses. In this regards, we may use a basic
+% instinct: without prior knowledge, the most relevant hypothesis might simply be
+% the most memorable recent event. However this definition is highly subjective,
+% and does not seem to fit well into the canvas of computing and AI.
+
+This kind of abduction seems however hard to translate into machine, as the difficulty
+comes from different factors. First, events can be of different nature, and not
+directly comparable. In fact, even for specific systems such as smart homes,
+events range from device removal to a presence detection or an unusually high
+temperature. How can one can then assess which ones are more memorable?
+Furthermore, even for comparable events, different characteristics can be put
+forward to argue for the most memorable: is a record-high temperature 47 days
+ago more memorable than the small deviation recorded just 3 minutes ago? To this
+day, no current system proposes to use all this information from various event
+types from different devices to compute a unified metrics of ``memorability''.
 
@@
 % devices and come to a unified metrics of ``memorability''.
-%
 
-To tackle this challenge, we propose to refer to events using a technique used
-by humans: naming. It appears that humans are able to name events and describe
-them, regardless of their nature or their characteristics. Furthermore, there is
-a noticeable correlation between the lenght of the description, in terms of the
-complexity of concepts and arguments, and the ``memorability'' of the event.
-Think, for instance, of how ``last year's hottest day'' appears simpler than
-``the 182th day of 7 years ago''. Using this method, the day described by the
-first sentence would be deemed simpler than the target of the second
-description. By comparing the description length of the event with the average
-description lenght of similar events, one would then have a measure of the
-``memorability'' of events, as memorable events stands out as much simpler (or
-more complex) than their neighbours.
+To tackle this challenge, we use as a starting point the observation that while
+all events, regardless of their characteristics or nature, can be uniquely
+described using a combination of qualifiers, the most memorable ones are likely
+to require less words to be described. Think, for instance, of ``last year's
+hottest day'' and ``the 182th day of 7 years ago''. By evaluating the lenght of
+each description, taking into account both the complexity of concept words (a
+time ranking, a temperature ranking), and the arguments used in both
+descriptions (the hottest, the 182th, 7), we can assign each event a complexity
+score. Memorable events, as they stand out, would then differ from their
+neighbors by being much simpler (or more complex).
 
@@
     filters. From the base memory (yellow), successive filters select the events
-    satisfying their predicates (grey arrows). The operations yields successive
+    satisfying their predicates (gray arrows). The operations yields successive
     subsets of the memory (yellow), until a final singleton memory is reach,
     meaning that the successive filters retrieve a unique event from the memory.
-    The predicates used for this retrival path and their arguments,
+    The predicates used for this retrieval path and their arguments,
     $\pi_{1}, \pi_{2}, \pi_{3}, k_{1}, k_{2}, k_{3}$, can be used as a retrieval
@@
 
-To programatically achieve what humans are easily capable to do, we rely on a
-basic model intuition of how memory works. Figure \ref{fig:filters} shows this
-process for the example description ``last year's hottest day''. We assume that
+To programmatically achieve this computation, we rely on principles from
+Algorithmic Information Theory, which provides formal definitions of complexity
+and simplicity of objects, which appear to be consistent with human perception
+\cite{li_introduction_2008, dessalles2011coincidences, delahaye_numerical_2012}.
+ Figure \ref{fig:filters} shows our solution applied to the example description
+ ``last year's hottest day''. We assume that
 a memory $\mathcal{M}$ is a unordered set of various distinct objects, upon
-which can be applied boolean predicates $\pi$ to describe them. These predicates thus
-provide filters $f_{\pi}$ that can be applied to the memory, selecting all events
-satifying predicate $\pi$. By applying successive
+which can be applied boolean predicates $\pi_{k}$ to describe them. These predicates thus
+provide filters $f_{\pi, k}$ that can be applied to the memory, selecting all events
+satisfying predicate $\pi_{k}$. By applying successive
 filters, we narrow down the size of the memory, until a single event remains.
-The complexity of this remaining events can thus be estimated as the simplest
-combination of predicative filters used to retrieve it.
+All the successive filters used to come to this singleton memory constitutes a
+\emph{retrieval path}, whose length, in bits, gives an upper bound of the
+description complexity of the retrieved event.
 
@@
 Algorithmic Information Theory and its applications in section~\ref{sec:theory},
-then we explicit our methods of computing the unnexpectedness of past events in
+then we explicit our methods of computing the unexpectedness of past events in
 cyber-physical systems~\ref{sec:computing}. We illustrate this approach by
@@
 \label{sec:theory}
-Kolmorov complexity formally quantifies the amount of information required
-for the computation of a finite binary string\footnote{While the defintion still
-holds for some inifinite binary strings (think of the representation of the
+Kolmogorov complexity formally quantifies the amount of information required
+for the computation of a finite binary string\footnote{While the definition still
+holds for some infinite binary strings (think of the representation of the
 decimals of $\pi$), we restrict ourselves to finite strings in this paper.} (or
-any object represented by a finite bynary
+any object represented by a finite binary
 string)\cite{kolmogorov_three_1965,li_introduction_2008}. For such a binary
@@
 \end{equation}
+
 The first notable property of this definition is its universality: while the
@@
 \end{equation}
+
 From this first result, we can then define a universal complexity $K(s)$, which
@@
 It is important to note that the notion of Kolmogorov complexity is not related
-to the computation complexity, as there is no requirement on the execution time
+to computation complexity, as there is no requirement on the execution time
 of the programs, only their length in bits matters for the computation of
@@
 
-While the introducion of Kolmogorov complexity was purely theoretical, many
-applications were found, notably because of its close relation with the
-intuitive notions of the complexity of items, events, situations. For
-instance,~\cite{murena_solving_2020} studied the use of Kolmogorov Complexity to
-identify and solve analogies.
+Interestingly, Kolmogorov's definition of complexity matches the intuitive
+notion and perception of complexity from a human standpoint. For instance, the
+complexity of short binary strings evaluated in \cite{delahaye_numerical_2012}
+shows similar results to human perception of complex strings and patterns. More
+recently, \cite{murena_solving_2020} used Kolmogorov complexity to solve
+analogies and showed results close to human expectations.
 
-Algorithmic Information Theory is also used to define the notion of
-unexpectedness: in \cite{dessalles2011coincidences}, the unexpectedness $U(e)$
-of an event is defined as the difference between an a-priori expected causal
+This bridge between Algorithmic Information Theory and human perception of
+complexity is extended to define the notion of simplicity and unexpectedness,
+which are considered of uttermost importance in cognitive science\cite{chater_simplicity_2003}.
+\cite{dessalles2011coincidences} proposes a formal definition of the unexpectedness $U(e)$
+of an event, as the difference between an a-priori expected causal
 complexity $C_{w}(e)$ and the actual observed complexity $C(e)$.
@@
 \end{equation}
-
 This result comes from the understanding that, while Kolmogorov complexity is
@@
 information processing in the human brain, and thus helps design a notion of
-simplicty or complexity of events\cite{}.
+simplicity or complexity of events.
 
 Using this measure, it is possible to model phenomenons such as coincidences:
-imagine that you happen to run into a friend in a park. The unnexpectedness will
-be much larger if this park is in Singapore than if it is in your common
-neighbourhood: the causal complexity $C_{w}$ is much higher if you a priori have
-no knowledge of your friend being around, while the description complexity
-remains the same.
+imagine that you happen to run into someone in a park. If this person is not
+noticeable, the event will be quite trivial. On the other hand, this person is
+your best friend, as the description complexity $C(e)$ drops, the causal
+complexity $C_{W}(e)$ remains unchanged, and the event becomes noticeable.
+
+Using these insights from AIT, we propose to measure the complexity of past
+events $e$ from a memory $\mathcal{M}$ by considering the quantity
+$K(e|\mathcal{M})$, that is the length of the shortest program outputting $e$,
+given the entire memory of past events $\mathcal{M}$ as an input. To limit the
+search and obtain a computable measure, we limit ourselves to only a subset of
+possible programs, which we will now describe.
 
@@
 \label{sec:computing}
-\subsection{A memory model}
+\subsection{Retrieving an event}
 We model the memory as an unordered set $M$ of events $e$. Individual events can
@@
   \label{eq:event}
-  e \equiv (l, t,\mathcal{D})
+  e = (l, t,\mathcal{D})
 \end{equation}
 where $l$ is the label, $t$ a timestamp and $\mathcal{D}$ a multi-dimensional
-datapoint representing various characteristics of the events (e.g.~its duration,
+data point representing various characteristics of the events (e.g.~its duration,
 the maximum temperature reached, the sensor name, its position).
@@
 
+As the output of a filter applied to the memory $\mathcal{M}$ is another memory
+object $\mathcal{M'} = f_{\pi, k}(\mathcal{M})$, we can compose filter
+functions. A sequence of such filters, applied in succession, is called a
+\emph{retrieval path}
+\begin{equation}
+\label{eq:ret_def}
+p = (f_{\pi_{1}, k_{1}}, \dots, f_{\pi_{n}, k_{n}})
+\end{equation}
+and by definition
+$p(\mathcal{M}) = f_{\pi_{n}, k_{n}}(\dots(f_{\pi_{1}, k_{1}}(\mathcal{M})))$.
+In case the result of the operation $p(\mathcal{M})$ contains a single element
+$e$, we say that the path $p$ \emph{retrieves} the element $e$ from $\mathcal{M}$, and write
+$p(\mathcal{M}) = e$.
+
 \subsection{Description complexity of events}
 
-From the previous definitions, we then estimate the complexity of events as the
-``simplest'' way of retrieving them from the memory, by applying successive
-filters. A retrieval path composed of various filters
-$p = f_{1} = f_{\pi_{1}, k_{1}}, f_{2} = f_{\pi_{2}, k_{2}}, \dots, f_{n} = f_{\pi_{n}, k_{n}}$
-is said to \emph{retrieve} an event $e$ from a memory $\mathcal{M}$ if and only
-if
-$p(\mathcal{M}) = f_{1} \circ f_{2} \circ \dots \circ f_{n} (\mathcal{M}) = \{e\}$.
-For an event $e$, if such a path exist, its complexity can be bounded by
-$K(p) = \sum_{f_{\pi, k} \in p} l(\pi) + l(k)$, the complexity of the retrieval
-path being the sum of the bit lengths required to describe the predicates $\pi$
-and their arguments $k$.
+% From the previous definitions, we then estimate the complexity of events as the
+% ``simplest'' way of retrieving them from the memory, by applying successive
+% filters. A retrieval path composed of various filters
+% $p = f_{1} = f_{\pi_{1}, k_{1}}, f_{2} = f_{\pi_{2}, k_{2}}, \dots, f_{n} = f_{\pi_{n}, k_{n}}$
+% is said to \emph{retrieve} an event $e$ from a memory $\mathcal{M}$ if and only
+% if
+% $p(\mathcal{M}) = f_{1} \circ f_{2} \circ \dots \circ f_{n} (\mathcal{M}) = \{e\}$.
+% For an event $e$, if such a path exist, its complexity can be bounded by
+% $K(p) = \sum_{f_{\pi, k} \in p} l(\pi) + l(k)$, the complexity of the retrieval
+% path being the sum of the bit lengths required to describe the predicates $\pi$
+% and their arguments $k$.
 
-The description complexity of an event $e$ can thus be defined as the minimum complexity of a retrieval path retrieving $e$:
+As presented in sec. \ref{sec:theory}, we are interested in computing the
+Kolmogorov complexity of an event $e$, conditionned on the memory $\mathcal{M}$.
+As this quantity is not computable, we use the previous definitions of retrieval
+paths and approximate the quantity $K(e|\mathcal{M})$ by the lenght of the
+shortest retrieval path $p$ retrieving $e$ from $\mathcal{M}$.
+
+Given that all predicate concepts can be described by a prefix program of length
+$l(\pi)$, and their arguments in length $l(k)$, each retrieval path $p$ can be
+described by the sequence of prefix description of all the predicate concepts and
+arguments used to construct its filters:
+
+\begin{equation}
+K(p) = l((f_{\pi_{1}, k_{1}}, \dots, f_{\pi{n}, k_{n}})) = \sum_{i=0}^{n} l(\pi_{i}) + l(k_{i})
+\end{equation}
+
+Using this definition of the lenght of a retrieval path, we can compute the
+estimation for the complexity $K(e|\mathcal{M})$, which, for clarity, we will
+denote simply by $K(e)$ for the rest of this paper.
 
@@
   \label{eq:desc_k}
-  K(e) = \min_{p | p(\mathcal{M}) = e} K(p) = \min_{p | p(\mathcal{M})=e} \sum_{f_{\pi, k} \in p} l(\pi) + l(k)
+  K(e) \approx \min_{p | p(\mathcal{M}) = e} K(p) = \min_{p | p(\mathcal{M})=e} \sum_{f_{\pi, k} \in p} l(\pi) + l(k)
 \end{equation}
@@
 \begin{algorithm}
-  $\mathtt{current_{explore}} \leftarrow [\mathcal{M}]$ \;
+  $\mathtt{current_{explore}} \leftarrow [(\mathcal{M}, 0)]$ \;
   $\mathtt{future_{explore} \leftarrow} [\;]$ \;
   $\mathtt{pass} \leftarrow 0$ \;
+  $\mathtt{K(e)} \leftarrow +\infty$ \;
   \While{$\mathtt{current_{explore}} \neq [\;]$ \textbf{and} $\mathtt{pass} < \mathtt{max\_pass}$}{
     \For{$(\mathtt{M_{prev}}, \mathtt{K_{prev}}) \in \mathtt{current_{explore}}$}{
-      \For{$\mathtt{p \in \mathcal{P}}$}{
+      \For{$\mathtt{\pi \in \mathcal{P}}$}{
         \For{$\mathtt{k \in \{0,1\}^{*}}$}{
-          $\mathtt{K_{current} \leftarrow l(p) + l(k) + K_{prev}}$ \;
+          $\mathtt{K_{current} \leftarrow l(\pi) + l(k) + K_{prev}}$ \;
           \If{$\mathtt{K_{current}} > \mathtt{max_{complex}}$}{
@@
           }
-          $\mathtt{M'} \leftarrow f_{p,k}(\mathtt{M_{prev}})$ \;
+          $\mathtt{M'} \leftarrow f_{\pi,k}(\mathtt{M_{prev}})$ \;
           \eIf{$\mathtt{M'} = \mathtt{\{e\}}$}{
@@
   }
-  \caption{Iterative computation of the complexity}
+  \caption{Iterative computation of the approximate complexity}
   \label{alg:complex_iter}
@@
 The definition used for description complexity in eq. \ref{eq:desc_k} allows for
-a direct implemenation, which is shown in alg. \ref{alg:complex_iter}. This
+a direct implementation, which is shown in alg. \ref{alg:complex_iter}. This
 algorithm operates iteratively: starting with the base memory $\mathcal{M}$
-(line 1), we apply all possible predicate concepts $\pi$ and programs $k$ (lines
-6-7), up to a given length $\mathtt{max\_len}$ bits, and apply them:
-$M' = f_{\pi, k}(M)$ (line 12). We then store the pairs
-$(M', \mathtt{len(}\pi, k\mathtt{)})$ in an array $\mathtt{future_{explore}}$. At
-the end of the iteration, the results of the filters become the memories which
-will be explored during the next iteration (lines 21--23). Each pass thus explore
-retrieval paths of increasing length. When a singleton memory is reached, the
-complexity of its unique element is upper-bounded with the length of the
-corresponding retrieval path (line 14).
+(line 1), we apply all possible predicate concepts $\pi$ from a given set
+$\mathcal{P}$ and programs $k$ (lines 6-7), up to a given length
+$\mathtt{max\_len}$ bits, and apply them: $M' = f_{\pi, k}(M)$ (line 12). We
+then store the pairs $(M', \mathtt{len(}\pi, k\mathtt{)})$ in an array
+$\mathtt{future_{explore}}$. At the end of the iteration, the results of the
+filters become the memories which will be explored during the next iteration
+(lines 21--23). Each pass thus explore retrieval paths of increasing length.
+When a singleton memory is reached, the complexity of its unique element is
+upper-bounded with the length of the corresponding retrieval path (line 14).
 
@@
 fit into our canvas. As $C_{w}$ is defined as the shortest program a
-``W-machine'' replicating the world's behaviour, based on the person's
+``W-machine'' replicating the world's behavior, based on the person's
 knowledge\cite{dessalles2011coincidences}, it is convenient to understand this
@@
 Now, when considering, for all possible predicates $\pi$, the corresponding
-neighbourhoods $N_{\pi}(e)$, with the convention that $N_{\pi}(e) = \emptyset$
+neighborhoods $N_{\pi}(e)$, with the convention that $N_{\pi}(e) = \emptyset$
 if $\pi(e) = \bot$, we can compute an average expected complexity for $e$:
@@
 The definition used for the expected complexity of the event $e$ in eq.
-\ref{eq:expected} holds the intuition of favorizing more similar events: if $e'$
-is \emph{very} similar to $e$, it will appear in many neighbouhoods, since it
+\ref{eq:expected} holds the intuition of favoring more similar events: if $e'$
+is \emph{very} similar to $e$, it will appear in many neighborhoods, since it
 satisfies mostly the same predicates as $e$. Therefore, its complexity term will
@@
 living-room will be simpler, since the location ``living-room'' is already known
-from knowing the consequnce. In terms of predicates, we therefore append a
+from knowing the consequence. In terms of predicates, we therefore append a
 description of $e$ to all programs $k$ of predicates:
@@
 situation, and with no other source of information, it can be used as a metrics
-to find relevant posisble surprising causes to the observed consequence.
+to find relevant possible surprising causes to the observed consequence.
 
@@
 recorded, making the detection of memorable events more important; ii) smart
-homes are prone to undergo atypical situations, highly dependant on the context,
+homes are prone to undergo atypical situations, highly dependent on the context,
 for which pre-established relations might fail to find good abduction
 candidates. Furthermore, the choice was also motivated by the ease of application: as
-previous works exist on smart homes, it is possible to simulate their behaviour
+previous works exist on smart homes, it is possible to simulate their behavior
 and quickly generate data to extract events from and test our methods.
@@
     \item Unusual weather: the outdoor conditions are set to unusually high or
-        low temeperatures.
+        low temperatures.
     \item Heater failures: heater can fail, making them turning off regardless
@@
   \includegraphics[width=\linewidth]{figures/ts_example}
-  \caption{Time series data from the simulation: outdoor temperature and
-    controller temperature of a room. We can see two temperature drops, one for
-    each room. What could be the causes of these drops?}
+  \caption{Time series data from the simulation: outdoor temperature (blue) and
+    controller temperature of a room (blue). On different occasion, the
+    controlled temperature deviates from its setting (21°C). We will evaluate if
+  our system finds these deviations as memorable, and if it can propose causal
+  hypotheses.}
   \label{fig:ts_example}
@@
 Since this is not the focus point of our present work (see sec.
-\ref{sec:related}), we simply apply threshold and precomputed conditions based
-detections to create a base of events.
+\ref{sec:related}), we simply apply threshold and pre-computed conditions based
+detection to create a base of events.
 
@@
         along axis $k_{2}$, where $k$ is a unique encoding for $(k_{1}, k_{2})$.
-        \item \texttt{day(e, k)}: whether the event $e$ occured $k$ days ago.
-        \item \texttt{month(e, k)}: whether the event $e$ occured $k$ months
+        \item \texttt{day(e, k)}: whether the event $e$ occurred $k$ days ago.
+        \item \texttt{month(e, k)}: whether the event $e$ occurred $k$ months
         ago.
-        \item \texttt{location(e, k)}: whether the event $e$ occured in zone
+        \item \texttt{location(e, k)}: whether the event $e$ occurred in zone
         $k$.
@@
 
-With a straightforward implemention of memory, predicates and filters, the
+With a straightforward implementation of memory, predicates and filters, the
 implementation of alg. \ref{alg:complex_iter} worked, but, as expected, took too
 long to be usable in realistic scenarios with hundreds or thousands of events to
-consider. In order to facilitate and speed up computaitons, we also implemented
+consider. In order to facilitate and speed up computations, we also implemented
 the following improvements:
@@
 
-The results of the descritpion complexity evaluation, for the described setup,
+The results of the description complexity evaluation, for the described setup,
 are shown in fig. \ref{fig:computed_cplx}. The entire computation, over 4
@@
 A main sequence of ``usual'' events, which complexity is roughly a logarithm of
-the elapsed time since their occurence, is visible. This corresponds to events
+the elapsed time since their occurrence, is visible. This corresponds to events
 for which the best retrieval path consists of a time description (e.g. ``2
@@
 rank alongside an axis (``the hottest day'', ``the second longest user's
-absence''), or the rare occurence of their kind (``the only fault on the heater'').
+absence''), or the rare occurrence of their kind (``the only fault on the heater'').
 
@@
   \caption{The computed complexities of events with at most retrieval paths of
-    length at most 4. Events of type ``day'' (blue), ``hot'' (green), ``cold'' (red) are shown.}
+    length at most 4. Events of type ``day'' (blue), ``hot'' (green), ``cold''
+    (red), ``device removal''(cyan) are shown.}
   \label{fig:computed_cplx}
@@
 same way (from the absolute value operation in eq. \ref{eq:unexpected}), we
-observe that some events are memorabale due to their context only. For instance,
+observe that some events are memorable due to their context only. For instance,
 a temperature anomaly occurring simultaneously to many other anomalies is
-notable, as it is costlier than expected to distinguish it from its neighbours.
+notable, as it is costlier than expected to distinguish it from its neighbors.
 
@@
   \includegraphics[width=\linewidth]{figures/complexities_surprises}
-  \caption{Unnexpectedness score for events in the memory}
+  \caption{Memorability score for events in the memory}
   \label{fig:result1}
@@
   \centering
-  \includegraphics[width=\linewidth]{example-image-a}
+  \includegraphics[width=\linewidth]{./figures/roc}
   \caption{Experimental ROC curve (True Positive Rate against False Positive Rate) for a
@@
 As an illustration of the abductive inference possible by using the memorability
-score, we used as a targed consequence the temperature drop visible in the raw
+score, we used as a target consequence the temperature drop visible in the raw
 time series data from our scenario (fig \ref{fig:ts_example}).
@@
 The main purpose of the present paper is to show the possible connections
-between exising definitions of simplicity from cognitive science, Algorithmic
+between existing definitions of simplicity from cognitive science, Algorithmic
 Information Theory and a practical use case in cyber-physical systems. As such,
@@
 As an extension, we suggest that in the future, we could explore online
-generation of such predicate. A posisbility would be to analyze discriminating
+generation of such predicate. A possibility would be to analyze discriminating
 dimensions of incoming data and create predicate as to name these differences,

### 2021-08-16 paper.tex
some changes

@@
 
-The rest of this paper is organized as follows: we first briefly introduce some notions of
-Algorithmic Information Theory and its applications in section~\ref{sec:theory},
-then we explicit our methods of computing the unexpectedness of past events in
-cyber-physical systems~\ref{sec:computing}. We illustrate this approach by
-providing an implementation relying on a few predicates on events from a smart
-home simulation~(section \ref{sec:example}). In section~\ref{sec:related} we briefly
-review other related works using complexity theory or trying to explain smart
-homes and how our work can be linked with them before exploring potential
+The rest of this paper is organized as follows: we first briefly introduce
+some notions of Algorithmic Information Theory and its applications in
+section~\ref{sec:theory}, then we explicit our methods of computing the
+unexpectedness of past events in cyber-physical systems~\ref
+{sec:computing}. We illustrate this approach by providing an implementation
+relying on a few predicates on events from a smart home simulation~
+(section \ref{sec:example}). In section~\ref{sec:related} we briefly review
+other related works using complexity theory or trying to explain smart homes
+and how our work can be linked with them before exploring potential
 extensions of our work in section \ref{sec:future}
@@
 It is important to note that the notion of Kolmogorov complexity is not related
-to computation complexity, as there is no requirement on the execution time
-of the programs, only their length in bits matters for the computation of
+to computation complexity, as there is no requirement on the execution time of
+the programs, only their length in bits matters for the computation of
 complexity. In fact, it can be shown that Kolmogorov complexity is not
-computable\cite{li_introduction_2008}. The sketch of the proof is the following:
-since the definition of complexity from eq. \ref{eq:kolm_def} requires to find
-the shortest of all programs on a given Turing machine outputting $s$, it
-requires to run all programs up to a given length and compare their result to
-$s$. However, being able to do so means that we would be able to tell whether
-these programs terminate. Since the termination of programs is famously
-undecidable, by extension, the computation of Kolmogorov complexity is
-impossible. One can however easily approximate it with upper bounds, by
-exhibiting a program outputting $s$.
+computable\cite{li_introduction_2008}. The sketch of the proof is the
+following: since the definition of complexity from eq. \ref
+{eq:kolm_def} requires to find the shortest of all programs on a given Turing
+machine outputting $s$, it requires to run all programs up to a given length
+and compare their result to $s$. However, being able to do so means that we
+would be able to tell whether these programs terminate. Since the termination
+of programs is famously undecidable, by extension, the computation of
+Kolmogorov complexity is impossible. One can however easily approximate it with
+upper bounds, by exhibiting a program outputting $s$.
 
@@
 complexity is extended to define the notion of simplicity and unexpectedness,
-which are considered of uttermost importance in cognitive science\cite{chater_simplicity_2003}.
-\cite{dessalles2011coincidences} proposes a formal definition of the unexpectedness $U(e)$
-of an event, as the difference between an a-priori expected causal
-complexity $C_{w}(e)$ and the actual observed complexity $C(e)$.
+which are considered of uttermost importance in cognitive science\cite
+{chater_simplicity_2003}.
+\cite{dessalles2011coincidences} proposes a formal definition of the
+ unexpectedness $U(e)$ of an event, as the difference between an a-priori
+ expected causal complexity $C_{w}(e)$ and the actual observed complexity $C
+ (e)$.
 \begin{equation}
@@
 (line 1), we apply all possible predicate concepts $\pi$ from a given set
-$\mathcal{P}$ and programs $k$ (lines 6-7), up to a given length
-$\mathtt{max\_len}$ bits, and apply them: $M' = f_{\pi, k}(M)$ (line 12). We
-then store the pairs $(M', \mathtt{len(}\pi, k\mathtt{)})$ in an array
-$\mathtt{future_{explore}}$. At the end of the iteration, the results of the
-filters become the memories which will be explored during the next iteration
-(lines 21--23). Each pass thus explore retrieval paths of increasing length.
-When a singleton memory is reached, the complexity of its unique element is
-upper-bounded with the length of the corresponding retrieval path (line 14).
+$\mathcal{P}$ and programs $k$ (lines 6-7), up to a given length $\mathtt
+{max\_len}$ bits, and apply them: $M' = f_{\pi, k}(M)$ (line 12). We then store
+the pairs $(M', \mathtt{len(}\pi, k\mathtt{)})$ in an array $\mathtt{future_
+{explore}}$. At the end of the iteration, the results of the filters become the
+memories which will be explored during the next iteration(lines 21--23). Each
+pass thus explore retrieval paths of increasing length. When a singleton memory
+is reached, the complexity of its unique element is upper-bounded with the
+length of the corresponding retrieval path (line 14).
 
@@
   \includegraphics[width=\linewidth]{./figures/roc}
-  \caption{Experimental ROC curve (True Positive Rate against False Positive Rate) for a
-    classifier based on our memorability score.}
+  \caption{Experimental ROC curve (True Positive Rate against False Positive
+   Rate) for a classifier based on our memorability score. Measures consider 23
+   manually flagged events as memorable (events added to the background noise
+   as described in Section~\ref{sec:example}.)}
   \label{fig:roc}
@@
 \label{sec:related}
-\textbf{TODO}
+Our work is intended to be integrated into larger-scale frameworks to monitor 
+and detect events in complex environment such as smart homes. In these works, the
+approach to smart homes is often regarded as self-organizing systems \cite{kramer_rigorous_2009,kounev_notion_2017}. As such, they present capacities of adaptation to new goals, 
+new components, new environment. A commonly used approach is the principle of 
+autonomic system, which minimizes user's intervention for management of the 
+system \cite{kounev_notion_2017,kephart_vision_2003}.
+
+We want to inscribe our work among other classical concurrent approaches to 
+abduction. In situations where more data is available, we could for instance rely
+on correlation or causal inference from known relations \cite{peters_elements_2017,fadiga_or_2021}.
+Previous relations between inference and complexity have been studied. In fact, the 
+case of inference was one of the motivations for R. Solomonoff to introduce his 
+universal algorithmic probability \cite{solomonoff_formal_1964} as a tool to 
+reach an idealized inference machine, creating the notion of complexity concomitantly 
+to Kolmogorov. Subsequently, notions of complexity re-emerged in causal inference: 
+\cite{janzing_causal_2010} found, when a causal link exists between two random variables,
+considering the direct joint probability is simpler, in terms of Kolmogorov complexity, 
+than the inverse direction.
+
+The topic of finding events from streams of time series data has already been explored in many ways. 
+\cite{aggarwal_outlier_2017} provides good review of modern approaches and techniques
+in the field.
+
+While all these works advocate in favor of a strong link between complexity and the 
+discovery of causes, not other works extends the notion up to the point we propose in this
+work, using complexity only as a tool for inference.
 

### 2021-08-17 paper.tex
Starting to take JL's corrections into account

@@
 \begin{document}
-
 \title{
-  What should I notice? Evaluating memorability of events to generate surprising causal candidates.
+  What should I notice? Evaluating memorability of events for abductive inference.
 }
@@
 
-As the number of connected devices and sensors grows, so does the amount of
-produced data. As a consequence, the ability to filter out a few events from past recordings
-as ``memorable'' becomes prime. While this task is often overlooked by many
-statistical approaches, which instead benefit from large quantity of
-information, it is part of the learning process for humans
-\cite{dehaene_how_2020}. Faced with new situations, humans are prone to infer
-previously seen unusual events as possible causes, and carry out further testing
-to assess the existence of a causal relation.
+% As the number of connected devices and sensors grows, so does the amount of
+% produced data. As a consequence, the ability to filter out a few events from past recordings
+% as ``memorable'' becomes prime. While this task is often overlooked by many
+% statistical approaches, which instead benefit from large quantity of
+% information, it is part of the learning process for humans
+% \cite{dehaene_how_2020}. Faced with new situations, humans are prone to infer
+% previously seen unusual events as possible causes, and carry out further testing
+% to assess the existence of a causal relation.
 
-For instance, it is possible, for humans without any prior knowledge about
-physics or electrical engineering, to infer a memorable thunderstorm as a
-possible cause for a general black-out just by using the memorability of
-the former. While this kind of abductive inference (i.e. inference to the cause
-of an observation \cite{magnani_abduction_2011}) can yield many
-false-positives, it can be considered in situations where the lack of knowledge
-from previous occurrences penalizes other approaches.
+As a user has just turned on the TV in her all-equipped living-room, the
+lights dim and the window blinds lower. Intrigued by this behavior, she
+quickly infers that both light dimming and the blinds closing occurred as a
+consequence of the TV being turned on. How did she come to this conclusion?
+By performing \emph{abductive inference} \cite{magnani_abduction_2011}, which
+is a key element of human's ability to understand the world: from the
+observed consequences, infer the possible causes.
+
+In this example, there are mainly three possibilities to come to the
+conclusion. If the user knows how the smart living-room system works, if she
+knows the set of applied rules or parameters, she has access to a causal
+knowledge she can use for the abduction. Otherwise, if she has no knowledge
+about the system but the behavior occurs frequently, examining past
+correlations will reveal that the television frequently causes the blinds to
+close and the lights to dim. Else, if this is the first instance of the TV
+being turned on in this living-room, there is no previous occurrence upon
+which the user may rely to draw conclusions. However, she remains able to
+infer the TV as a possible cause for the observed reaction: by noticing the
+TV as a memorable recent event (since it is its first occurrence), she
+proposes it as a possible cause for the questioned phenomenon. This example
+shows how humans are able to use distinct methods to perform abductive tasks
+and infer new knowledge. While abduction of the first two kinds can be
+automated using knowledge bases and statistical approaches, this paper
+proposes an approach to evaluate the memorability of events, allowing
+machines to perform ``memorability-based'' abduction.
+
+ % For instance, it is
+% possible, for humans without any prior knowledge about
+% physics or electrical engineering, to infer a memorable thunderstorm as a
+% possible cause for a general black-out just by using the memorability of
+% the former. While this kind of abductive inference (i.e. inference to the cause
+% of an observation \cite{magnani_abduction_2011}) can yield many
+% false-positives, it can be considered in situations where the lack of knowledge
+% from previous occurrences penalizes other approaches.
 
@@
 
-This kind of abduction seems however hard to translate into machine, as the difficulty
-comes from different factors. First, events can be of different nature, and not
-directly comparable. In fact, even for specific systems such as smart homes,
-events range from device removal to a presence detection or an unusually high
+The difficulty of implementing this kind of abduction for machines comes from
+several factors. First, events can be of different nature, and not directly
+comparable. In fact, even for specific systems such as smart homes, events
+range from device removal to a presence detection or an unusually high
 temperature. How can one can then assess which ones are more memorable?
@@
 forward to argue for the most memorable: is a record-high temperature 47 days
-ago more memorable than the small deviation recorded just 3 minutes ago? To this
-day, no current system proposes to use all this information from various event
-types from different devices to compute a unified metrics of ``memorability''.
+ago more memorable than the small deviation recorded just 3 minutes ago? To
+this day, no current system proposes to use all this information from various
+event types from different devices to compute a unified metrics of
+``memorability''.
 
@@
 
-To tackle this challenge, we use as a starting point the observation that while
-all events, regardless of their characteristics or nature, can be uniquely
-described using a combination of qualifiers, the most memorable ones are likely
-to require less words to be described. Think, for instance, of ``last year's
-hottest day'' and ``the 182th day of 7 years ago''. By evaluating the lenght of
-each description, taking into account both the complexity of concept words (a
-time ranking, a temperature ranking), and the arguments used in both
-descriptions (the hottest, the 182th, 7), we can assign each event a complexity
+% To tackle this challenge, we use as a starting point the observation that while
+% all events, regardless of their characteristics or nature, can be uniquely
+% described using a combination of qualifiers, the most memorable ones are likely
+% to require less words to be described. Think, for instance, of ``last year's
+% hottest day'' and ``the 182th day of 7 years ago''. By evaluating the lenght of
+% each description, taking into account both the complexity of concept words (a
+% time ranking, a temperature ranking), and the arguments used in both
+% descriptions (the hottest, the 182th, 7), we can assign each event a complexity
+% score. Memorable events, as they stand out, would then differ from their
+% neighbors by being much simpler (or more complex).
+
+To tackle this challenge, we rely on the following observation: while all
+events, regardless of their characteristics or nature, can be uniquely
+described using a combination of qualifiers, the most memorable ones are
+likely to require less words to be described. Think, for instance, of how
+``last year's hottest day'' and ``the 182nd day from 7 years ago'': the
+former seems simpler to describe. By evaluating the length of each
+description, taking into account both the complexity of concept words (a date
+of occurrence, a temperature ranking), and the arguments (the hottest, the
+182ndThis
+kind of inference, named \emph{abductive inference} \cite
+{magnani_abduction_2011}, is a key element of humans's ability to understand the world., 7) used in both descriptions, we can assign each event a \emph{description complexity}
 score. Memorable events, as they stand out, would then differ from their
-neighbors by being much simpler (or more complex).
+neighbors and require simpler and less numerous qualifiers to be
+unambiguously described. As such, they will stand out as simpler.
 
@@
   \centering
-  \includegraphics[width=\linewidth]{figures/filters}
+  \includegraphics[width=0.9\linewidth]{figures/filters}
   \caption{An example of retrieving an event through successive predicative
-    filters. From the base memory (yellow), successive filters select the events
-    satisfying their predicates (gray arrows). The operations yields successive
-    subsets of the memory (yellow), until a final singleton memory is reach,
-    meaning that the successive filters retrieve a unique event from the memory.
-    The predicates used for this retrieval path and their arguments,
-    $\pi_{1}, \pi_{2}, \pi_{3}, k_{1}, k_{2}, k_{3}$, can be used as a retrieval
-    program. Its length gives an upper bound to the description complexity of
-    the event $e_{h}$: ``last year's hottest day''.}
+   filters. From the base memory (yellow), successive filters select the
+   events satisfying the associated predicate (grey arrow). For example,
+   filter $f_1$ selects events from last year, i.e. which satisfy the
+   predicate $\mathtt{date}(event, 1\_year)$. In this case, successively
+   applying filters $f_1$, $f_2$ and $f_3$ yields a unique event $e_h$, last
+   year's hottest day. The complexity of this event can then be upper bounded
+   by the complexity of the three filters, since they give a unambiguous way
+   to describe the event in the memory.}
+   % their predicates (gray arrows). The
+   % operations yields successive subsets of the memory (yellow), until a final
+   % singleton memory is reach, meaning that the successive filters retrieve a
+   % unique event from the memory. The predicates used for this retrieval path
+   % and their arguments, $\pi_{1}, \pi_{2}, \pi_{3}, k_{1}, k_{2}, k_{3}$, can
+   % be used as a retrieval program. Its length gives an upper bound to the
+   % description complexity of the event $e_{h}$: ``last year's hottest
+   % day''.}
   \label{fig:filters}
@@
 
-To programmatically achieve this computation, we rely on principles from
-Algorithmic Information Theory, which provides formal definitions of complexity
-and simplicity of objects, which appear to be consistent with human perception
-\cite{li_introduction_2008, dessalles2011coincidences, delahaye_numerical_2012}.
- Figure \ref{fig:filters} shows our solution applied to the example description
- ``last year's hottest day''. We assume that
-a memory $\mathcal{M}$ is a unordered set of various distinct objects, upon
-which can be applied boolean predicates $\pi_{k}$ to describe them. These predicates thus
-provide filters $f_{\pi, k}$ that can be applied to the memory, selecting all events
-satisfying predicate $\pi_{k}$. By applying successive
-filters, we narrow down the size of the memory, until a single event remains.
-All the successive filters used to come to this singleton memory constitutes a
-\emph{retrieval path}, whose length, in bits, gives an upper bound of the
-description complexity of the retrieved event.
+For machines to use description complexity to distinguish memorable events from others, we need a formal canvas and computation methods which coincide with human intuition .
+To do so, we rely on principles from Algorithmic Information Theory, which also appear to be consistent with the human perception of complexity\cite{li_introduction_2008,dessalles2011coincidences,delahaye_numerical_2012}.
+Our method is as follows: we
+consider events as being unordered elements from a base \emph{memory}. To
+reproduce the language features applicable to events, we use \emph
+{predicates}, i.e. functions assigning a boolean value to an event and
+eventual arguments. For instance, the predicate $\mathtt{date}
+(\cdot, 1\_year)$ assigns \emph{true} to events that occurred last year. Selecting
+all events, from the memory, that satisfy a given predicate corresponds to
+a \emph{filter} operation, and yields another memory, a subset of the former.
+This filtering operation can then be repeated, each iteration, selecting
+fewer events, until a singleton memory is reached, meaning that the sequence
+of predicates can \emph{retrieve} unambiguously the unique remaining event.
+The description complexity of this event can thus be upper-bounded by the
+number of bits required to describe the filters used in the retrieval
+process. Figure \ref{fig:filters} illustrates this evaluation for the event
+$e_h$: ``last year's hottest day''.
+
+% To programmatically achieve this computation, we rely on principles from
+% Algorithmic Information Theory, which provides formal definitions of complexity
+% and simplicity of objects, which appear to be consistent with human perception
+% \cite{li_introduction_2008, dessalles2011coincidences, delahaye_numerical_2012}.
+%  Figure \ref{fig:filters} shows our solution applied to the example description
+%  ``last year's hottest day''. We assume that
+% a memory $\mathcal{M}$ is a unordered set of various distinct objects, upon
+% which can be applied boolean predicates $\pi_{k}$ to describe them. These predicates thus
+% provide filters $f_{\pi, k}$ that can be applied to the memory, selecting all events
+% satisfying predicate $\pi_{k}$. By applying successive
+% filters, we narrow down the size of the memory, until a single event remains.
+% All the successive filters used to come to this singleton memory constitutes a
+% \emph{retrieval path}, whose length, in bits, gives an upper bound of the
+% description complexity of the retrieved event.
 

### 2021-09-17 paper.tex
commit

@@
 By performing \emph{abductive inference} \cite{magnani_abduction_2011}, which
-is a key element of human's ability to understand the world: from the
+is a key element of humans' ability to understand the world: from the
 observed consequences, infer the possible causes.
@@
 described using a combination of qualifiers, the most memorable ones are
-likely to require less words to be described. Think, for instance, of how
+likely to require less words to be described. Think, for instance, of
 ``last year's hottest day'' and ``the 182nd day from 7 years ago'': the
-former seems simpler to describe. By evaluating the length of each
+former seems simpler to describe. To quantify the notion, we can evaluate the complexity of each
 description, taking into account both the complexity of concept words (a date
 of occurrence, a temperature ranking), and the arguments (the hottest, the
-182ndThis
-kind of inference, named \emph{abductive inference} \cite
-{magnani_abduction_2011}, is a key element of humans's ability to understand the world., 7) used in both descriptions, we can assign each event a \emph{description complexity}
-score. Memorable events, as they stand out, would then differ from their
-neighbors and require simpler and less numerous qualifiers to be
-unambiguously described. As such, they will stand out as simpler.
+182nd, 7). The resulting value defines the  \emph{description complexity}
+of the events. As our intuition is that memorable events require simpler and less numerous qualifiers to be
+unambiguously described than other unremarkable ones, they will stand out as simpler in terms of description complexity.
 
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{figures/filters}
-  \caption{An example of retrieving an event through successive predicative
-   filters. From the base memory (yellow), successive filters select the
-   events satisfying the associated predicate (grey arrow). For example,
+\includegraphics[width=\linewidth]{figures/filters}
+\caption{Retrieving an event through successive predicative
+filters. From the base memory (yellow), successive filters select events satisfying the associated predicate (grey arrows). For example,
    filter $f_1$ selects events from last year, i.e. which satisfy the
@@
 to describe the event in the memory.}
-   % their predicates (gray arrows). The
-   % operations yields successive subsets of the memory (yellow), until a final
-   % singleton memory is reach, meaning that the successive filters retrieve a
-   % unique event from the memory. The predicates used for this retrieval path
-   % and their arguments, $\pi_{1}, \pi_{2}, \pi_{3}, k_{1}, k_{2}, k_{3}$, can
-   % be used as a retrieval program. Its length gives an upper bound to the
-   % description complexity of the event $e_{h}$: ``last year's hottest
-   % day''.}
   \label{fig:filters}
@@
 
-For machines to use description complexity to distinguish memorable events from others, we need a formal canvas and computation methods which coincide with human intuition .
-To do so, we rely on principles from Algorithmic Information Theory, which also appear to be consistent with the human perception of complexity\cite{li_introduction_2008,dessalles2011coincidences,delahaye_numerical_2012}.
+For machines to implement and compute description complexity, we need a formal canvas and computation methods which coincide with human intuition .
+To do so, we rely on Algorithmic Information Theory, which also appears to be consistent with the human perception of complexity\cite{li_introduction_2008,dessalles2011coincidences,delahaye_numerical_2012}.
 Our method is as follows: we
@@
 some notions of Algorithmic Information Theory and its applications in
-section~\ref{sec:theory}, then we explicit our methods of computing the
-unexpectedness of past events in cyber-physical systems~\ref
+section~\ref{sec:theory}, then we explicit our methods to compute the
+memorability of past events in cyber-physical systems~\ref
 {sec:computing}. We illustrate this approach by providing an implementation
@@
 computable\cite{li_introduction_2008}. The sketch of the proof is the
-following: since the definition of complexity from eq. \ref
-{eq:kolm_def} requires to find the shortest of all programs on a given Turing
+following: since the definition of complexity from Equation. \ref
+{eq:kolmo_def} requires to find the shortest of all programs on a given Turing
 machine outputting $s$, it requires to run all programs up to a given length
@@
 \begin{equation}
-  \label{eq:unexpected}
-  U(e) = |C_{w}(e) - C(e)|
+\label{eq:unexpected} U(e) = C_{w}(e) - C(e)
 \end{equation}
@@
 
-Using this measure, it is possible to model phenomenons such as coincidences:
-imagine that you happen to run into someone in a park. If this person is not
-noticeable, the event will be quite trivial. On the other hand, this person is
-your best friend, as the description complexity $C(e)$ drops, the causal
-complexity $C_{W}(e)$ remains unchanged, and the event becomes noticeable.
+The definition of Equation \ref{eq:unexpected} allows to model phenomenons
+such as coincidences: imagine that you happen to run into someone in a park.
+If this person has no particular link to you, the event will be quite
+trivial: the complexity of describing this person will be equivalent to
+distinguishing it from the global population, which will be roughly
+equivalent as the complexity of describing that this person happens to be in
+the same park as you. On the other hand, if you run into your best friend in
+a park, as the complexity of describing your best friend is significantly
+lower, the description complexity $C(e)$ drops while the causal complexity
+remains $C_W (e)$ unchanged. Therefore the event becomes unexpected.
 
-Using these insights from AIT, we propose to measure the complexity of past
-events $e$ from a memory $\mathcal{M}$ by considering the quantity
-$K(e|\mathcal{M})$, that is the length of the shortest program outputting $e$,
-given the entire memory of past events $\mathcal{M}$ as an input. To limit the
-search and obtain a computable measure, we limit ourselves to only a subset of
-possible programs, which we will now describe.
+Using these insights from AIT, we define the memorability $M(e)$ of an event as the absolute difference between the description complexity $K_d(e)$ of an event and its expected description complexity $K_{exp}(e)$:
+\begin{equation}
+  \label{eq:memorability}
+  M(e) = |K_{exp}(e) - K_d(e)|
+\end{equation}
+Contrary to the definition of unexpectedness from Equation \ref{eq:unexpected}, we use an absolute value: this is done to acknowledge events more complex than expected as memorable\footnote{In the original paper \cite{dessalles2011coincidences}, unusually complex events could be handled by considering complexity itself as a way to describe the event: see ``the Pisa Tower effect''\cite{dessalles_pisa_nodate}}.
+In the next section, we formally define the description complexity $K_d$ and the expected complexity $K_{exp}$ of events and detail how we can compute approximations of these quantities.
 
-\section{Computing the unexpectedness of events}
+\section{Computing the memorability of events}
 \label{sec:computing}
 \subsection{Retrieving an event}
-We model the memory as an unordered set $M$ of events $e$. Individual events can
-be seen as data points, augmented with a label indicating their nature
-(temperature events, failure events, addition/removal events), and a timestamp:
+
+Computing the memorability of events begins with the introduction of formal definitions of events, how they are stored, how we can describe them and how these descriptions can be used to retrieve them.
+
+We define \emph{events} as data points augmented with a \emph{label} indicating their nature (temperature event, failure event, addition/removal of a device) and a timestamp of occurrence. Formally:
 \begin{equation}
@@
 \end{equation}
-where $l$ is the label, $t$ a timestamp and $\mathcal{D}$ a multi-dimensional
-data point representing various characteristics of the events (e.g.~its duration,
-the maximum temperature reached, the sensor name, its position).
+where $l$ is the label, $t$ the timestamp and $\mathcal{D}$ a multi-dimensional data point representing the various characteristics of $e$: its duration, the maximum temperature reached, the sensor name, its position, etc.
+Labels can also be considered as classes of events, of which each event is a particular instance.
 
-We define \emph{predicates} as boolean function operating on these events, and
-taking an additional argument, encoded as a finite binary string:
+To model how humans are able to describe events by using qualifiers, we use \emph{predicates}: function operating on events, eventually taking additional parameters, returning a boolean value: $\pi(e, a_1, a_2, \dots, a_n) \mapsto \{O,1\}$ is a predicate of arity $n$ operating on event $e$.
+In the rest of this paper, we will prefer the equivalent notation $\pi(e, k) \mapsto \{0,1\}$, where $k$ is a binary string encoding the sequence of arguments $a_1, \dots, a_n$.
+Using this notation, the predicate $\pi$ becomes a boolean function operating on $\mathbf{E} \times \{0,1\}^*$:
 \begin{equation}
@@
 \end{equation}
-An example of predicate is to use $\pi = \mathtt{year}$ and $k=1$, thus
-constructing the predicate $\mathtt{year()}e, Y\mathtt{)}$, which tells whether
-the event $e$ occurred $Y$ years ago.
+An example of predicate is to take $\pi = \mathtt{year}$ and $k$ a string encoding the number $1$, thus
+constructing the predicate $\mathtt{year(}e, 1\mathtt{)}$, which tells whether
+the event $e$ occurred $1$ years ago.
 
-From these predicates, we can construct filter operations on the memory. A
-filter operates by selecting the events satisfying a given predicate
-$\pi_{k}$ and constructing a resulting memory from these events.
+As events occur, they are stored in a \emph{memory} $M_0$.
+As they are not directly comparable, the memory $M_0$ can be considered as having the structure of an unordered set.
+We denote by $\mathcal{M}$ the subsets of $M_0$.
+By extension, elements of $\mathcal{M}$, i.e.
+subsets of $M_0$, are also called \emph{memories}.
+
+By applying a given predicate $\pi_k$ to all events contained in a memory $M \subseteq M_0$, and selecting only events satisfying $\pi_k$ yields another memory $M_1 \subseteq M \subseteq M_0$.
+We call this operation a \emph{filter}:
 \begin{equation}
   \label{eq:filter}
- f_{\pi, k}(\mathcal{M}) = \left\{e \in \mathcal{M} | \pi_{k}(e) \right\}
+f_{\pi, k}: \begin{cases}
+  \mathcal{M} & \mapsto \mathcal{M}             \\
+  M           & \mapsto \{e \in M | \pi_k(e) \}
+\end{cases}
 \end{equation}
-For instance, using the same $\pi = \mathtt{year}$ and $k=1$ as above, we would
+For instance, using the same $\pi = \mathtt{year}$ and $k=1$ as above, we can
 build the filter $f_{\pi, k} = \mathtt{last\_{}year}$, which selects all events
@@
 
-As the output of a filter applied to the memory $\mathcal{M}$ is another memory
-object $\mathcal{M'} = f_{\pi, k}(\mathcal{M})$, we can compose filter
-functions. A sequence of such filters, applied in succession, is called a
-\emph{retrieval path}
+As the output of a filter applied to a memory $M$ is another memory
+object $M' \subseteq M$, we can compose filter
+functions. A sequence of such filters is called a \emph{retrieval path}
 \begin{equation}
@@
 and by definition
-$p(\mathcal{M}) = f_{\pi_{n}, k_{n}}(\dots(f_{\pi_{1}, k_{1}}(\mathcal{M})))$.
-In case the result of the operation $p(\mathcal{M})$ contains a single element
-$e$, we say that the path $p$ \emph{retrieves} the element $e$ from $\mathcal{M}$, and write
-$p(\mathcal{M}) = e$.
+$p(M) = f_{\pi_{n}, k_{n}}(\dots(f_{\pi_{1}, k_{1}}(M)))$.
+In case the result of the operation $p(M)$ contains a single element
+$e$, we say that the path $p$ \emph{retrieves} the element $e$ from $M$, and write
+$p(M) = e$. In the example shown in Figure \ref{fig:filters}, the three filters $f_1, f_2, f_3$ form a retrieval path retrieving the event ``last year's hottest day'' from the base memory $M_0$.
 
@@
 
-% From the previous definitions, we then estimate the complexity of events as the
-% ``simplest'' way of retrieving them from the memory, by applying successive
-% filters. A retrieval path composed of various filters
-% $p = f_{1} = f_{\pi_{1}, k_{1}}, f_{2} = f_{\pi_{2}, k_{2}}, \dots, f_{n} = f_{\pi_{n}, k_{n}}$
-% is said to \emph{retrieve} an event $e$ from a memory $\mathcal{M}$ if and only
-% if
-% $p(\mathcal{M}) = f_{1} \circ f_{2} \circ \dots \circ f_{n} (\mathcal{M}) = \{e\}$.
-% For an event $e$, if such a path exist, its complexity can be bounded by
-% $K(p) = \sum_{f_{\pi, k} \in p} l(\pi) + l(k)$, the complexity of the retrieval
-% path being the sum of the bit lengths required to describe the predicates $\pi$
-% and their arguments $k$.
-
-As presented in sec. \ref{sec:theory}, we are interested in computing the
-Kolmogorov complexity of an event $e$, conditionned on the memory $\mathcal{M}$.
-As this quantity is not computable, we use the previous definitions of retrieval
-paths and approximate the quantity $K(e|\mathcal{M})$ by the lenght of the
-shortest retrieval path $p$ retrieving $e$ from $\mathcal{M}$.
-
-Given that all predicate concepts can be described by a prefix program of length
-$l(\pi)$, and their arguments in length $l(k)$, each retrieval path $p$ can be
-described by the sequence of prefix description of all the predicate concepts and
-arguments used to construct its filters:
+As presented in sec.
+\ref{sec:theory}, we are interested in computing an approximation of the description complexity of an event $e$.
+From the above definitions, if there is a path $p$ retrieving $e$ from the base memory $M_0$, i.e.
+$p(M_0) = e$, this path provides a possible unambiguous description for $e$.
+We therefore define the description complexity of $e$ as the minimum complexity of a path $p$ retrieving $e$ from the base memory $M_0$.
 
 \begin{equation}
-K(p) = l((f_{\pi_{1}, k_{1}}, \dots, f_{\pi{n}, k_{n}})) = \sum_{i=0}^{n} l(\pi_{i}) + l(k_{i})
+  \label{eq:k_desc}
+  K_d(e) = \min_{p \in P_\infty} \{L(p) | p(M_0) = e\}
 \end{equation}
 
-Using this definition of the lenght of a retrieval path, we can compute the
-estimation for the complexity $K(e|\mathcal{M})$, which, for clarity, we will
-denote simply by $K(e)$ for the rest of this paper.
+where the bit-length $L(p)$ of a retrieval path is defined as the number of bits of a string encoding the path.
+If we limit ourselves to prefix-free strings encoding predicates and arguments, the total bit length is given by:
+\begin{align}
+  \label{eq:bit_lenght_p}
+  L(p) & = L((f_{\pi_1,k_1}, \dots, f_{\pi_n, k_n}))     \\
+       & = L(\pi_1) + L(k_1) + \dots + L(\pi_n) + L(k_n)
+\end{align}
+
+By considering only a finite number of possible predicates $\pi$ and arguments $k$, and a maximum path length, we can construct a finite set $P$ of possible retrieval paths.
+By limiting the search over this set, we can upper bound the description complexity, and use this upper bound as an approximation:
 
 \begin{equation}
-  \label{eq:desc_k}
-  K(e) \approx \min_{p | p(\mathcal{M}) = e} K(p) = \min_{p | p(\mathcal{M})=e} \sum_{f_{\pi, k} \in p} l(\pi) + l(k)
+\label{eq:approx_k_desc}
+K_d(e) \leq \min_{p \in P \land p(M_0) = e} L(p) = \min_{p \in P \land p(M_0)=e} \sum_{f_{\pi, k} \in p} L(\pi) + L(k)
 \end{equation}
@@
 
-The definition used for description complexity in eq. \ref{eq:desc_k} allows for
-a direct implementation, which is shown in alg. \ref{alg:complex_iter}. This
-algorithm operates iteratively: starting with the base memory $\mathcal{M}$
-(line 1), we apply all possible predicate concepts $\pi$ from a given set
-$\mathcal{P}$ and programs $k$ (lines 6-7), up to a given length $\mathtt
+The approximation of description complexity from Equation \ref{eq:approx_k_desc} allows for
+a direct implementation, which is shown in Algorithm \ref{alg:complex_iter}. This
+algorithm operates iteratively: starting with the base memory $M_0$
+(line 1), we apply all possible predicate concepts $\pi$ from a given finite set
+$\Pi$ and programs $k$ (lines 6-7), up to a given length $\mathtt
 {max\_len}$ bits, and apply them: $M' = f_{\pi, k}(M)$ (line 12). We then store
@@
 
-\subsection{Computing an unexpectedness measure}
-Equation \ref{eq:unexpected} gives a universal definition of unexpectedness, as
-the difference between the complexity of the phenomenon generating an event and
-the complexity required to describe this event. In our particular application
-case, we may directly use the description complexity $K(e)$ as defined in
-\ref{eq:desc_k}.
+\subsection{Computing Memorability}
+As stated in Equation \ref{eq:memorability}, we define memorability $M(e)$ as the absolute difference between the description complexity of an event and its expected value.
+As we've just defined $K_d(e)$ and provided an approximation in Equation \ref{eq:approx_k_desc}, we now focus on defining the description expected complexity of an event, $K_{exp}(e)$.
 
-The term $C_{w}(e)$ from eq. \ref{eq:unexpected} is harder to formally
-fit into our canvas. As $C_{w}$ is defined as the shortest program a
-``W-machine'' replicating the world's behavior, based on the person's
-knowledge\cite{dessalles2011coincidences}, it is convenient to understand this
-term as the expectation one would have regarding the complexity of the event
-$e$, \emph{before} actually evaluating it.
+As intended in the Equation \ref{eq:memorability}, this term evaluates the complexity the user, or the system, would expect the event $e$ to have without actually describing it, based on their previous knowledge.
+In our canvas, this prior knowledge consists of the base memory $M_0$.
+The expected complexity of the event $e$ can be computed with a simple first-order approximation, i.e.
+estimating the average complexity of ``similar events'' over the base memory $M_0$.
 
-A possible proxy for this quantity is to compute the average complexity
-of \emph{similar} events. Still, difficulty remains in the definition of what
-can be considered \emph{similar} events. Given that we have already used the
-notion of predicates to describe the events and compute their description
-complexity, we can use them as simple metrics for similarity: for each predicate
-$\pi$ satisfied by $e$, we can define a set of similar events $N_{\pi}(e)$ as
-all the events also satisfying $\pi$.
+Still, difficulty remains in the definition of what
+can be considered \emph{similar} events.
+Given that we deal with non comparable events, the only possibility to universally define the notion of similarity is to once again refer to \emph{predicates}.
+Thus, for a given event $e$, and a given predicate $\pi_k$, we define a $\pi_k$ neighborhood of $e$ as the set $N_{\pi, k}(e)$ of all other events satisfying $\pi_k$.
 
@@
 \label{eq:similar}
-N_{\pi}(e) = \{e'\in \mathcal{M}, \pi(e') \wedge e' \neq e\}
+N_{\pi, k}(e) = \{e'\in M_0, \pi_k(e') \wedge e' \neq e\}
 \end{equation}
 
-Now, when considering, for all possible predicates $\pi$, the corresponding
-neighborhoods $N_{\pi}(e)$, with the convention that $N_{\pi}(e) = \emptyset$
-if $\pi(e) = \bot$, we can compute an average expected complexity for $e$:
+Now, when considering, for all possible predicates $\pi_k$, the corresponding
+neighborhoods $N_{\pi, k}(e)$, with the convention that $N_{\pi, k}(e) = \emptyset$
+if $e$ does not satisfy $\pi_k$, we can compute an average expected complexity for $e$:
 
@@
 K_{exp}(e) = \frac{
-                    \sum_{\pi} \sum_{e' \in N_{\pi}(e)} K(e')
+\sum_{\pi, k} \sum_{e' \in N_{\pi, k}(e)} K_d(e')
                   }{
-                    \sum_{\pi} |N_{\pi}(e)|
+\sum_{\pi, k} |N_{\pi, k}(e)|
                   }
@@
 
-The definition used for the expected complexity of the event $e$ in eq.
-\ref{eq:expected} holds the intuition of favoring more similar events: if $e'$
+This definition is consistent with the intuitive ideas that more similar events should weight more in the computation. Indeed, if $e'$
 is \emph{very} similar to $e$, it will appear in many neighborhoods, since it
-satisfies mostly the same predicates as $e$. Therefore, its complexity term will
-be present in more terms in eq. \ref{eq:expected}, and will weight more towards
+satisfies mostly the same predicates as $e$. Therefore, it will
+be present in more terms in Equation \ref{eq:expected}, and will weight more in
 the final result.
@@
 to an observed effect $e$, we try to find the most remarkable event in past
-memory in that regards. While our ``memorability'' score aims to identify the
+memory in that regard. While our ``memorability'' score aims to identify the
 most remarkable events, it does not take into account the added knowledge about
@@
 
-This added knowledge would however influence the unexpectedness of a possible
-cause $c$, from eq. \ref{eq:unexpected}: even though the causal complexity
-$C_{W}(c)$ is likely to remain unchanged, as it relies on the model of the world in
-the person's mind, the description complexity $C(c)$ is replaced with the
-conditional complexity $C(c|e)$, as the knowledge of $e$ is included in the
-problem's data.
+This added knowledge can be integrated into the description complexity definition by using conditioned predicates.
+That is considering as given, and therefore ``free'' in terms of complexity, information contained in the cause effect $e$.
+For instance, when looking for a cause for
+an anomaly in the living-room, other anomalies occurring in the same
+living-room will be simpler, since the location ``living-room'' is already known
+from knowing the consequence.
 
-In our canvas of predicative filtering, this corresponds to adding for free, in
-terms of program length, information about $e$ to predicates. This translates
-into considering relative order between events: when investigating a cause for
-an anomaly in the living-rooms, other anomalies occurring in the same
-living-room will be simpler, since the location ``living-room'' is already known
-from knowing the consequence. In terms of predicates, we therefore append a
-description of $e$ to all programs $k$ of predicates:
-$\pi_{k}(c) = \pi_{k::e}(c)$, where $::$ is the append operation.
+Formally, we now consider only predicates where knowledge of the effect event $e$  is appended to all programs $k$:
+$\pi_{k::e}(c)$, where $::$ is the append operation.
+The set of paths obtained with such predicates is $P^\infty_e$.
+This append operation is free in terms of bit-length in the computation of complexity, since the effect event $e$ is an input of the problem.
+Therefore, we have $L'(\pi_{k::e}) = L(\pi_k) = L(\pi) + L(k)$.
+Which finally gives the definition for the conditional description complexity:
 
-This leads to another score $Abd(c,e)$ (abduction score of effect $c$ for cause
-$e$), which is defined similarly to the description complexity in eq.
-\ref{eq:desc_k} and using relative filters.
+\begin{align}
+  \label{eq:abd_k}
+  K_d(c | e) & = \min_{p \in P^\infty_e} \{L'(p), \quad p(M_0)=c \}                                                \\
+             & = \min_{p \in P^\infty_e} \left\{\sum_{f_{\pi, k::e} \in p} L(\pi) + L(k), \quad p(M_0) = c\right\}
+\end{align}
+
+This new conditional description complexity translates the additional information provided to the
+system when answering a user's request.
+It can then be averaged over similar events to compute the expected conditional description complexity, $K_{exp}(c|e)$.
+From this, we come to the definition of the conditional memorability, which measures how memorable an event $c$ is, considering the knowledge of another event $e$:
 
 \begin{equation}
-  \label{eq:abd_k}
-  Abd(c) = \min_{p | p(\mathcal{M}|e) = c} K(p) = \min_{p | p(\mathcal{M}|e)=c} \sum_{f_{\pi, k} \in p} l(\pi) + l(k)
+\label{eq:cond_mem}
+M(c|e) = |K_{exp}(c|e) - K_d(c|e)|
 \end{equation}
-This abduction score translates the additional information provided to the
-system when answering a user's request. As such, it encapsulates the idea
+
+Conditional memorability encapsulates the idea
 presented as the motivation of this paper: when confronted to a surprising
-situation, and with no other source of information, it can be used as a metrics
-to find relevant possible surprising causes to the observed consequence.
+situation, and with no other source of information, we can identify events that appear more memorable than others, with regards to the target effect event.
+As such, our conditional memorability score provides a ranking that can be used for abductive inference, proposing as best candidates the most memorable ones.
 
@@
   \caption{Time series data from the simulation: outdoor temperature (blue) and
-    controller temperature of a room (blue). On different occasion, the
+controller temperature of a room (orange). On different occasion, the
     controlled temperature deviates from its setting (21°C). We will evaluate if
@@
 \ref{sec:related}), we simply apply threshold and pre-computed conditions based
-detection to create a base of events.
+detection to create a set of events.
 
-This base of events consists the basis of the initial memory $\mathcal{M}$ used
+This set of events constitutes the basis of the initial memory $M_0$ used
 for computations.
@@
 \begin{itemize}
-        \item \texttt{label(e, k)}: whether the event $e$ has the label ranked
+\item $\mathtt{label(e, k)}$: whether the event $e$ has the label ranked
         $k^{th}$ in the memory.
-        \item \texttt{rank(e, k)}: whether the event $e$ has the rank $k_{1}$
-        along axis $k_{2}$, where $k$ is a unique encoding for $(k_{1}, k_{2})$.
-        \item \texttt{day(e, k)}: whether the event $e$ occurred $k$ days ago.
-        \item \texttt{month(e, k)}: whether the event $e$ occurred $k$ months
+\item $\mathtt{rank(e, a_1, a_2)}$: whether the event $e$ has the rank $a_1$
+along axis $a_2$.
+\item $\mathtt{day(e, k)}$: whether the event $e$ occurred $k$ days ago.
+\item $\mathtt{month(e, k)}$: whether the event $e$ occurred $k$ months
         ago.
-        \item \texttt{location(e, k)}: whether the event $e$ occurred in zone
+\item $\mathtt{location(e, k)}$: whether the event $e$ occurred in zone
         $k$.
@@
 iterations (meaning that retrieval paths contained at most 4 filters), took
-around 30 seconds on a modern commercial laptop with an i7-8700u CPU.
+around 30 seconds on a commercial laptop with an i7-8700u CPU.
 
@@
 the computation of this measure treats unusually complex or simple events the
-same way (from the absolute value operation in eq. \ref{eq:unexpected}), we
+same way (from the absolute value operation in Equation \ref{eq:unexpected}), we
 observe that some events are memorable due to their context only. For instance,
@@
 
-% TODO
-\textbf{TODO}
+\textbf{TODO: the experiment still needs to be done: some additional data is required here (creating the scenario and finding causes.)}
+
+\subsubsection{Discussion}
+\label{ssub:discussion}
 
@@
 \label{sec:related}
+\textbf{TODO: this section is still WIP: I've added most references and short comments, but I still need to write the story linking them.}
 Our work is intended to be integrated into larger-scale frameworks to monitor 
@@
 
+More recently, \cite{marx_causal_2018} used Minimum Description Lenght to determine, given a joint probability distribution over $(X,Y)$, whether $X$ causes $Y$ or $Y$ causes $X$.
+Their method is based on tree models, and implying that a model respecting the causal relation will be simpler to describe.
+
+\cite{tatti_finding_2008} PACK algorithm to cluster data using the simplest possible trees: maybe link this to isolation forests?
+
 The topic of finding events from streams of time series data has already been explored in many ways. 
@@
 in the field.
+Some previous work can also be noted for having used AIT techniques to qualify and detect events in time series data.
+For instance, \cite{batista_complexity-invariant_2011,fadlallah_weighted-permutation_2013} propose weighted permutation entropy as a proxy for complexity measures in time series data, and use it to find relations between different time series.
+\cite{hu_discovering_2011} proposes a MDL approach to find the intrinsic dimensions of time series.
+All these approaches are interesting, and can be integrated into our canvas as tools to detect events using only complexity.
+As such, one could acheive a purely complexity-driven process for detecting and qualifying events.
+
+Among various methods, the approach of ``Isolation forests''\cite{liu_isolation_2008,hariri_extended_2021} is closely related, by design, to ours.
+It evaluates the isolation of data points by constructing random binary tree classifiers.
+On average, outlier points will require less operations to be singled out.
+Using the average height of leaves in the tree as a metrics, this approach succeeds in identifying outlier points without having to define a ``typical'' point.
+This approach can be understood in terms of complexity: each node of binary tree classifier needs a fixed amount of information to be described (which variable and threshold are used).
+So nodes that place higher need less information to be described.
+As such, outliers need less information to be singled out.
+Compared to ours, this method is tailored for data-points living in the same metric space.
+By using predicates as a proxy for complexity computation, our methods is more general, as it is agnostic regarding the nature of events.
+However, the introdcution of predicates adds a subjectivity in the determination of memorable points, as we will discuss later.
 
 While all these works advocate in favor of a strong link between complexity and the 
-discovery of causes, not other works extends the notion up to the point we propose in this
-work, using complexity only as a tool for inference.
+discovery of causes, not other works extends the notion up to the point we propose in this paper, using complexity only as a tool to express the intuitive notion of memorability, and using it for inference.
 

### 2021-10-04 paper.tex
pre-JL-relecture

@@
   This score can later be used either to select only relevant information to be
- remembered, or to propose causal hypotheses in unusual situations, on demand.
+  remembered, or to propose causal hypotheses in unusual situations on demand.
   As such, the approach aims at being complementary to more traditional
@@
   \pi : \begin{cases}
-    \mathcal{M}\times \{0,1\}^{*} &\mapsto \{0,1\} \\
-    (e, k) &\mapsto \pi_{k}(e)
+    \mathcal{M}\times \{0,1\}^{*} & \mapsto \{0,1\}    \\
+    (e, k)                        & \mapsto \pi_{k}(e)
   \end{cases}

### 2021-10-04 paper.tex
Adding modifications from JL's version

--- 2021-09-17 paper.tex
@@
 \title{
-  What should I notice? Evaluating memorability of events for abductive inference.
+What should I notice? Evaluating the memorability of events for abductive inference.
 }
@@
   associative patterns. Our proposal is to formalize and compute a ``memorability''
-  score over a memory of various recorded events from a cyber-physical system.
- This score can later be used either to select only relevant information to be
- remembered, or to propose causal hypotheses in unusual situations, on demand.
- As such, the approach aims at being complementary to more traditional
- learning-focused techniques. We provide theoretical ground for our approach by
- using and extending existing results and ideas from Algorithmic Information
- Theory and provide an implementation example, showing practical results in a
- smart-home scenario.
+score over a set of events recorded from a cyber-physical system.
+This score can then be used to select relevant information to be
+remembered and to propose causal hypotheses in unusual situations on demand.
+The approach is meant to be complementary to more traditional
+learning-focused abduction techniques. We provide theoretical ground for our approach
+based on Algorithmic Information
+Theory. We also provide an implementation example, a
+smart-home scenario, to show how our approach could translate in practice.
 \end{abstract}
@@
 \begin{IEEEkeywords}
-  Complexity, Algorithmic Information Theory, Simplicity, Abduction, Surprise
+Kolmogorov complexity, Algorithmic Information Theory, Simplicity, Abduction, Surprise
 \end{IEEEkeywords}
@@
 As a user has just turned on the TV in her all-equipped living-room, the
-lights dim and the window blinds lower. Intrigued by this behavior, she
-quickly infers that both light dimming and the blinds closing occurred as a
+lights dim and the window blinds go down. Intrigued by this behavior, she
+quickly infers that both light dimming and blind closing occurred as a
 consequence of the TV being turned on. How did she come to this conclusion?
-By performing \emph{abductive inference} \cite{magnani_abduction_2011}, which
+By performing \emph{abductive inference} \cite{magnani_abduction_2011}. This mental operation
 is a key element of humans' ability to understand the world: from the
-observed consequences, infer the possible causes.
+observed consequences, they infer the possible causes.
 
 In this example, there are mainly three possibilities to come to the
-conclusion. If the user knows how the smart living-room system works, if she
-knows the set of applied rules or parameters, she has access to a causal
-knowledge she can use for the abduction. Otherwise, if she has no knowledge
-about the system but the behavior occurs frequently, examining past
-correlations will reveal that the television frequently causes the blinds to
-close and the lights to dim. Else, if this is the first instance of the TV
-being turned on in this living-room, there is no previous occurrence upon
-which the user may rely to draw conclusions. However, she remains able to
-infer the TV as a possible cause for the observed reaction: by noticing the
-TV as a memorable recent event (since it is its first occurrence), she
-proposes it as a possible cause for the questioned phenomenon. This example
-shows how humans are able to use distinct methods to perform abductive tasks
-and infer new knowledge. While abduction of the first two kinds can be
-automated using knowledge bases and statistical approaches, this paper
-proposes an approach to evaluate the memorability of events, allowing
-machines to perform ``memorability-based'' abduction.
+conclusion. (1) If the user knows how the smart living-room system works, if she
+knows the underlying rules or parameters, she may use this causal
+knowledge to perform abduction. (2) If she has no knowledge
+about the system but made several observation of the same behavior, she may examine past
+correlations and figure out that turning on the TV set often leads the blinds to
+close and the lights to dim. (3) If there are no previous occurrences of the event (e.g. it is the first time she turns on the TV
+in the living-room), she may still be able to
+suspect that the TV is a possible cause for the observed event, just because it appears to her as a memorable recent event (as it is its first occurrence). This example
+suggests that human beings are able to use distinct methods to perform abductive tasks
+and infer new knowledge. While the first two mechanisms can be
+automated using knowledge bases and statistical methods, ``memorability-based'' abduction needs to be investigated in more depth.
 
@@
 
-The difficulty of implementing this kind of abduction for machines comes from
-several factors. First, events can be of different nature, and not directly
-comparable. In fact, even for specific systems such as smart homes, events
-range from device removal to a presence detection or an unusually high
-temperature. How can one can then assess which ones are more memorable?
-Furthermore, even for comparable events, different characteristics can be put
-forward to argue for the most memorable: is a record-high temperature 47 days
+Defining a score of memorability is not straightforward. First, events can be of different nature, and not directly
+comparable. Even for restricted systems such as smart homes, noticeable events
+range from device removal to presence detection or unusually high
+temperatures. Even for comparable events, the problem is to weigh different characteristics: is a record-high temperature 47 days
 ago more memorable than the small deviation recorded just 3 minutes ago? To
-this day, no current system proposes to use all this information from various
+our knowledge, no current system proposes to combine various
 event types from different devices to compute a unified metrics of
@@
 
-To tackle this challenge, we rely on the following observation: while all
+To address this issue, we started from the following intuition: while all
 events, regardless of their characteristics or nature, can be uniquely
-described using a combination of qualifiers, the most memorable ones are
-likely to require less words to be described. Think, for instance, of
-``last year's hottest day'' and ``the 182nd day from 7 years ago'': the
-former seems simpler to describe. To quantify the notion, we can evaluate the complexity of each
-description, taking into account both the complexity of concept words (a date
-of occurrence, a temperature ranking), and the arguments (the hottest, the
-182nd, 7). The resulting value defines the  \emph{description complexity}
-of the events. As our intuition is that memorable events require simpler and less numerous qualifiers to be
-unambiguously described than other unremarkable ones, they will stand out as simpler in terms of description complexity.
+described using a combination of quantitative or qualitative qualifiers, the most memorable ones are
+likely to require less words to be described.
+``Last year's hottest day'' is simpler than ``the 182\textsuperscript{nd} day 7 years ago''.
+How can we quantify this relative simplicity?
+We could evaluate the complexity of each
+description, taking into account both the complexity of the concept words (a date
+of occurrence, a temperature ranking), and of the arguments (1st hottest,
+182 and 7). This approach, The resulting values define the  \emph{description complexity}
+of events. Our intuition is that memorable events require simpler and less numerous qualifiers to be
+unambiguously described than unremarkable ones.
 
@@
 applying filters $f_1$, $f_2$ and $f_3$ yields a unique event $e_h$, last
-   year's hottest day. The complexity of this event can then be upper bounded
-   by the complexity of the three filters, since they give a unambiguous way
-to describe the event in the memory.}
+year's hottest day. The complexity of this event can then be upper-bounded
+by the complexity of the three filters, since they give an unambiguous way
+to describe the event within the memory.}
   \label{fig:filters}
@@
 
-For machines to implement and compute description complexity, we need a formal canvas and computation methods which coincide with human intuition .
-To do so, we rely on Algorithmic Information Theory, which also appears to be consistent with the human perception of complexity\cite{li_introduction_2008,dessalles2011coincidences,delahaye_numerical_2012}.
+For machines to implement and compute description complexity, we need a formal framework and computation methods that are in line with human intuition.
+Algorithmic Information Theory (AIT) appears to be such a framework, as it is consistent with the human perception of complexity\cite{li_introduction_2008,dessalles2011coincidences,delahaye_numerical_2012}.
 Our method is as follows: we
-consider events as being unordered elements from a base \emph{memory}. To
+consider events as being elements stored in what we call \emph{base memory}. To
 reproduce the language features applicable to events, we use \emph
-{predicates}, i.e. functions assigning a boolean value to an event and
-eventual arguments. For instance, the predicate $\mathtt{date}
-(\cdot, 1\_year)$ assigns \emph{true} to events that occurred last year. Selecting
+{predicates}, i.e. functions assigning a boolean value to events. For instance, the predicate $\mathtt{date}
+  (\cdot, 1\_year)$ is \emph{true} of events that occurred last year. Selecting
 all events, from the memory, that satisfy a given predicate corresponds to
-a \emph{filter} operation, and yields another memory, a subset of the former.
-This filtering operation can then be repeated, each iteration, selecting
-fewer events, until a singleton memory is reached, meaning that the sequence
-of predicates can \emph{retrieve} unambiguously the unique remaining event.
+a \emph{filter} operation. It generates another memory that is a subset of the previous one.
+The filtering operation can then be repeated, selecting
+fewer events at each iteration, until a singleton memory is reached. This means that the sequence
+of predicates could unambiguously \emph{retrieve} the unique remaining event.
 The description complexity of this event can thus be upper-bounded by the
 number of bits required to describe the filters used in the retrieval
-process. Figure \ref{fig:filters} illustrates this evaluation for the event
-$e_h$: ``last year's hottest day''.
+process. Figure \ref{fig:filters} illustrates this process for the event: ``last year's hottest day''.
 
@@
 
-The rest of this paper is organized as follows: we first briefly introduce
-some notions of Algorithmic Information Theory and its applications in
-section~\ref{sec:theory}, then we explicit our methods to compute the
-memorability of past events in cyber-physical systems~\ref
-{sec:computing}. We illustrate this approach by providing an implementation
-relying on a few predicates on events from a smart home simulation~
-(section \ref{sec:example}). In section~\ref{sec:related} we briefly review
-other related works using complexity theory or trying to explain smart homes
-and how our work can be linked with them before exploring potential
+We will first briefly introduce
+some basic notions of Algorithmic Information Theory. Then, in
+section~\ref{sec:computing}, we show how we compute the
+memorability of past events. We illustrate our approach in~
+(section \ref{sec:example}) by describing an implementation
+in a smart home simulation context. In section~\ref{sec:related} we review
+a few studies that use complexity theory to generate explanations for the smart home
+and we discuss how our work can be linked with them. We eventually explore potential
 extensions of our work in section \ref{sec:future}
@@
 Kolmogorov complexity formally quantifies the amount of information required
-for the computation of a finite binary string\footnote{While the definition still
+for the computation of a finite binary string\footnote{Though the definition
 holds for some infinite binary strings (think of the representation of the
-decimals of $\pi$), we restrict ourselves to finite strings in this paper.} (or
+decimals of $\pi$), we restrict ourselves here to finite strings.} (or
 any object represented by a finite binary
-string)\cite{kolmogorov_three_1965,li_introduction_2008}. For such a binary
-string $s$, its complexity $K(s)$, is the bit length of the shortest program $p$
+string)\cite{kolmogorov_three_1965,li_introduction_2008}. The complexity $K(s)$ of a (finite) binary string is the length in bits of the shortest program $p$
 which, if given as input to a universal Turing Machine $U$, outputs $s$.
@@
 choice of the Turing machine $U$ used for the computations appears in the
-definition of eq. \ref{eq:kolmo_def}, all results stand, up to an additional
-constant. Think, for instance, how any Turing-complete programming language can
-be turned into any other language, using a compiler program. In fact, since any
+definition of eq. \ref{eq:kolmo_def}, all results hold, up to an additional
+constant, if we change the machine. Think how any Turing-complete programming language can
+be turned into any other language, using an interpreter or a compiler program. Since any
 Turing machine $U'$ can be emulated by $U$ from a
@@
   \label{eq:inequality_univ}
-  K_{U}(s) \le l(p_{u}) + K_{U}(s)
+K_{U'}(s) \le l(p_{u}) + K_{U}(s)
 \end{equation}
 
-From this first result, we can then define a universal complexity $K(s)$, which
-no longer depends on the choice of a Turing machine $U$, such that, for any
-machine $U$,
+From this first result, we can then define complexity $K(s)$, based on the choice of a reference Turing machine, such that, for any other
+machine $U$ taken from the set $\text{TM}$ of Turing machines:
 \begin{equation}
@@
 \end{equation}
-where the additional constant $C_{U}$ does not depend on the object $s$.
+where the additional constant $C_{U}$ does not depend on $s$.
 
-It is important to note that the notion of Kolmogorov complexity is not related
-to computation complexity, as there is no requirement on the execution time of
-the programs, only their length in bits matters for the computation of
-complexity. In fact, it can be shown that Kolmogorov complexity is not
-computable\cite{li_introduction_2008}. The sketch of the proof is the
-following: since the definition of complexity from Equation. \ref
-{eq:kolmo_def} requires to find the shortest of all programs on a given Turing
-machine outputting $s$, it requires to run all programs up to a given length
-and compare their result to $s$. However, being able to do so means that we
-would be able to tell whether these programs terminate. Since the termination
-of programs is famously undecidable, by extension, the computation of
-Kolmogorov complexity is impossible. One can however easily approximate it with
-upper bounds, by exhibiting a program outputting $s$.
+Note that the notion of Kolmogorov complexity involves no requirement on the execution time of
+programs, only their length in bits matters for the computation of
+complexity. Though Kolmogorov complexity can be shown to be
+incomputable\cite{li_introduction_2008},
+%The sketch of the proof is the
+%following: since the definition of complexity from Equation. \ref
+%{eq:kolmo_def} requires to find the shortest of all programs on a given Turing
+%machine outputting $s$, it requires to run all programs up to a given length
+%and compare their result to $s$. However, being able to do so means that we
+%would be able to tell whether these programs terminate. Since the termination
+%of programs is famously undecidable, by extension, the computation of
+%Kolmogorov complexity is impossible.
+it can be approximated with
+upper bounds by exhibiting a program outputting $s$.
 
-Interestingly, Kolmogorov's definition of complexity matches the intuitive
+Interestingly, Kolmogorov complexity matches the intuitive
 notion and perception of complexity from a human standpoint. For instance, the
@@
 
-This bridge between Algorithmic Information Theory and human perception of
-complexity is extended to define the notion of simplicity and unexpectedness,
-which are considered of uttermost importance in cognitive science\cite
+The bridge between Algorithmic Information Theory (AIT) and human perception of
+complexity can be pushed farther thanks to the notions of simplicity and unexpectedness,
+which are sometimes regarded of uttermost importance in cognitive science\cite
 {chater_simplicity_2003}.
@@
 unexpectedness $U(e)$ of an event, as the difference between an a-priori
- expected causal complexity $C_{w}(e)$ and the actual observed complexity $C
+expected causal complexity $K_{w}(e)$ and the actual observed complexity $K
   (e)$.
 \begin{equation}
-\label{eq:unexpected} U(e) = C_{w}(e) - C(e)
+\label{eq:unexpected} U(e) = K_{w}(e) - K(e)
 \end{equation}
+
 This result comes from the understanding that, while Kolmogorov complexity is
-computed with Turing machine, it can be used as an elegant proxy for modeling
-information processing in the human brain, and thus helps design a notion of
+ideally computed using a Turing machine, it can be used as a proxy for modeling
+information processing in the human brain, and thus can be used to define a notion of
 simplicity or complexity of events.
 
-The definition of Equation \ref{eq:unexpected} allows to model phenomenons
+Definition \ref{eq:unexpected} allows to model phenomena
 such as coincidences: imagine that you happen to run into someone in a park.
@@
 trivial: the complexity of describing this person will be equivalent to
-distinguishing it from the global population, which will be roughly
-equivalent as the complexity of describing that this person happens to be in
-the same park as you. On the other hand, if you run into your best friend in
+distinguishing it from the global population, which is also roughly
+equivalent to the (causal) complexity of describing the circumstances having brought this person to be in
+that park as the same time as you. On the other hand, if you run into your best friend in
 a park, as the complexity of describing your best friend is significantly
-lower, the description complexity $C(e)$ drops while the causal complexity
-remains $C_W (e)$ unchanged. Therefore the event becomes unexpected.
+lower, the description complexity $K(e)$ drops while the causal complexity
+$K_W (e)$ remains unchanged. This is why this latter  event appears unexpected.
 
@@
 \end{equation}
-Contrary to the definition of unexpectedness from Equation \ref{eq:unexpected}, we use an absolute value: this is done to acknowledge events more complex than expected as memorable\footnote{In the original paper \cite{dessalles2011coincidences}, unusually complex events could be handled by considering complexity itself as a way to describe the event: see ``the Pisa Tower effect''\cite{dessalles_pisa_nodate}}.
-In the next section, we formally define the description complexity $K_d$ and the expected complexity $K_{exp}$ of events and detail how we can compute approximations of these quantities.
+Contrary to the definition of unexpectedness from Equation \ref{eq:unexpected}, we use an absolute value: we do so to acknowledge the fact that events more complex than expected can be memorable as well\footnote{In the original paper \cite{dessalles2011coincidences}, exceptionally complex events are described by considering complexity itself as a way to describe the event: see ``the Pisa Tower effect''\cite{dessalles_pisa_nodate}}.
+In the next section, we define computational approximations for the description complexity $K_d$ and the expected complexity $K_{exp}$ of events.
 
@@
 
-Computing the memorability of events begins with the introduction of formal definitions of events, how they are stored, how we can describe them and how these descriptions can be used to retrieve them.
-
+%Computing the memorability of events begins with the introduction of formal definitions of events, how they are stored, how we can describe them and how these descriptions can be used to retrieve them.
 We define \emph{events} as data points augmented with a \emph{label} indicating their nature (temperature event, failure event, addition/removal of a device) and a timestamp of occurrence. Formally:
@@
 
-To model how humans are able to describe events by using qualifiers, we use \emph{predicates}: function operating on events, eventually taking additional parameters, returning a boolean value: $\pi(e, a_1, a_2, \dots, a_n) \mapsto \{O,1\}$ is a predicate of arity $n$ operating on event $e$.
+To model how humans are able to describe events by using qualifiers, we use \emph{predicates}: Boolean functions operating on events and, possibly, additional parameters: $\pi(e, a_1, a_2, \dots, a_n) \mapsto \{O,1\}$ is a predicate of arity $n$ operating on event $e$.
 In the rest of this paper, we will prefer the equivalent notation $\pi(e, k) \mapsto \{0,1\}$, where $k$ is a binary string encoding the sequence of arguments $a_1, \dots, a_n$.
@@
 \end{equation}
-An example of predicate is to take $\pi = \mathtt{year}$ and $k$ a string encoding the number $1$, thus
+As an example of predicate, consider $\pi = \mathtt{year}$ and $k$ a string encoding the number $1$, thus
 constructing the predicate $\mathtt{year(}e, 1\mathtt{)}$, which tells whether
-the event $e$ occurred $1$ years ago.
+the event $e$ occurred $1$ year ago.
 
-As events occur, they are stored in a \emph{memory} $M_0$.
+As events occur, they are stored in the  \emph{base memory} $M_0$.
 As they are not directly comparable, the memory $M_0$ can be considered as having the structure of an unordered set.
@@
 
-By applying a given predicate $\pi_k$ to all events contained in a memory $M \subseteq M_0$, and selecting only events satisfying $\pi_k$ yields another memory $M_1 \subseteq M \subseteq M_0$.
+By applying a given predicate $\pi$ to all events contained in a memory $M \subseteq M_0$, and selecting only events satisfying $\pi$, one gets another memory $M_1 \subseteq M \subseteq M_0$.
 We call this operation a \emph{filter}:
@@
 \mathcal{M} & \mapsto \mathcal{M}             \\
-  M           & \mapsto \{e \in M | \pi_k(e) \}
+M           & \mapsto \{e \in M | \pi(e) \}
 \end{cases}
@@
 \end{equation}
-and by definition
+By definition
 $p(M) = f_{\pi_{n}, k_{n}}(\dots(f_{\pi_{1}, k_{1}}(M)))$.
@@
 By considering only a finite number of possible predicates $\pi$ and arguments $k$, and a maximum path length, we can construct a finite set $P$ of possible retrieval paths.
-By limiting the search over this set, we can upper bound the description complexity, and use this upper bound as an approximation:
+By limiting the search over this set, we get an upper bound of description complexity, and use this upper bound as an approximation:
 
@@
 a direct implementation, which is shown in Algorithm \ref{alg:complex_iter}. This
-algorithm operates iteratively: starting with the base memory $M_0$
+algorithm operates iteratively: starting from the base memory $M_0$
 (line 1), we apply all possible predicate concepts $\pi$ from a given finite set
@@
 As stated in Equation \ref{eq:memorability}, we define memorability $M(e)$ as the absolute difference between the description complexity of an event and its expected value.
-As we've just defined $K_d(e)$ and provided an approximation in Equation \ref{eq:approx_k_desc}, we now focus on defining the description expected complexity of an event, $K_{exp}(e)$.
+As we've just defined $K_d(e)$ and provided an approximation in Equation \ref{eq:approx_k_desc}, we now focus on defining the \emph{expected} description complexity of an event, $K_{exp}(e)$ that appears in Equation \ref{eq:memorability}
 
-As intended in the Equation \ref{eq:memorability}, this term evaluates the complexity the user, or the system, would expect the event $e$ to have without actually describing it, based on their previous knowledge.
-In our canvas, this prior knowledge consists of the base memory $M_0$.
+$K_{exp}(e)$ evaluates the complexity the user, or the system, would expect the occurrence of event $e$ to have, based on their previous knowledge.
+In our framework, this prior knowledge consists of the base memory $M_0$.
 The expected complexity of the event $e$ can be computed with a simple first-order approximation, i.e.
@@
 
-Still, difficulty remains in the definition of what
-can be considered \emph{similar} events.
-Given that we deal with non comparable events, the only possibility to universally define the notion of similarity is to once again refer to \emph{predicates}.
-Thus, for a given event $e$, and a given predicate $\pi_k$, we define a $\pi_k$ neighborhood of $e$ as the set $N_{\pi, k}(e)$ of all other events satisfying $\pi_k$.
+Still, there is a difficulty in defining what
+should be considered \emph{similar} events.
+Given that we deal with non comparable events, we may define the notion of similarity by referring once again to \emph{predicates}.For a given event $e$ and a given predicate $\pi_k$, we define a $\pi_k$-neighborhood of $e$ as the set $N_{\pi, k}(e)$ of all other events satisfying $\pi_k$.
 
@@
 
-This definition is consistent with the intuitive ideas that more similar events should weight more in the computation. Indeed, if $e'$
-is \emph{very} similar to $e$, it will appear in many neighborhoods, since it
-satisfies mostly the same predicates as $e$. Therefore, it will
-be present in more terms in Equation \ref{eq:expected}, and will weight more in
+This definition is consistent with the intuitive idea that more similar events should weigh more in the computation. Indeed, if $e'$
+is very similar to $e$, it will appear in many neighborhoods, since it
+satisfies most of the predicates that $e$ satisfies. Therefore, it will
+be present in more terms in Equation \ref{eq:expected}, and will weigh more in
 the final result.
@@
 % les prédicats'' `sympathiques` par rapport à ce que l'on questionne.
-The problem of the abductive inference is different from the basic computation
-of the memorability score. Here, \emph{knowing} that we want to find a cause $c$
-to an observed effect $e$, we try to find the most remarkable event in past
-memory in that regard. While our ``memorability'' score aims to identify the
-most remarkable events, it does not take into account the added knowledge about
-the effect $e$ that is available for abductive inference.
+Abductive inference builds on the computation
+of the memorability score. \emph{Knowing} that we want to find a cause $c$
+for an observed effect $e$, we try to find the most remarkable event in past
+memory that is related to $e$. While our ``memorability'' score identifies remarkable past events, it does not take into account their relatedness to $e$.
 
-This added knowledge can be integrated into the description complexity definition by using conditioned predicates.
-That is considering as given, and therefore ``free'' in terms of complexity, information contained in the cause effect $e$.
+The knowledge attached to the occurrence of $e$ can be integrated into the description complexity definition by using conditional complexity $K_d(c | e)$: The information contained in $e$ is considered as given, and therefore ``free'' in terms of complexity.
 For instance, when looking for a cause for
 an anomaly in the living-room, other anomalies occurring in the same
-living-room will be simpler, since the location ``living-room'' is already known
-from knowing the consequence.
+living-room will be simpler, as the location ``living-room'' is already known
+from the observation of the current anomaly.
 
@@
 $\pi_{k::e}(c)$, where $::$ is the append operation.
-The set of paths obtained with such predicates is $P^\infty_e$.
+The set of paths obtained with such predicates is noted $P^\infty_e$.
 This append operation is free in terms of bit-length in the computation of complexity, since the effect event $e$ is an input of the problem.
 Therefore, we have $L'(\pi_{k::e}) = L(\pi_k) = L(\pi) + L(k)$.
-Which finally gives the definition for the conditional description complexity:
+We get a definition for the conditional description complexity:
 
@@
 It can then be averaged over similar events to compute the expected conditional description complexity, $K_{exp}(c|e)$.
-From this, we come to the definition of the conditional memorability, which measures how memorable an event $c$ is, considering the knowledge of another event $e$:
+From this, we come to the definition of the conditional memorability, which measures how memorable an event $c$ turns out to be in the context of the occurrence of another event $e$:
 
@@
 Conditional memorability encapsulates the idea
-presented as the motivation of this paper: when confronted to a surprising
-situation, and with no other source of information, we can identify events that appear more memorable than others, with regards to the target effect event.
-As such, our conditional memorability score provides a ranking that can be used for abductive inference, proposing as best candidates the most memorable ones.
+presented as the motivation of this paper: when confronted with a surprising
+situation, and in the absence of any other source of information, events that appear more memorable than others, with regards to the target event will be selected as potential causes.
+As such, our conditional memorability score provides a ranking that can be used for abductive inference.
+
+\textbf{TODO:} Mettre un passage ici pour dire que ça résoud le dilemne de la récence présenté dans l'intro
 
@@
 \subsection{Setup}
-To test our approach, we set up an experimental smart home setup. This choice of
+We considere here an experimental smart home setup. This choice of
 configuration is motivated by the challenges posed by smart homes for abductive
@@
 recorded, making the detection of memorable events more important; ii) smart
-homes are prone to undergo atypical situations, highly dependent on the context,
+homes are prone to experience atypical situations, highly dependent on the context,
 for which pre-established relations might fail to find good abduction
-candidates. Furthermore, the choice was also motivated by the ease of application: as
-previous works exist on smart homes, it is possible to simulate their behavior
-and quickly generate data to extract events from and test our methods.
+candidates. Our choice was also motivated by
+the existence of previous work involving smart home simulations capable of quickly generating data from which we could extract events and test our methods.
 
@@
 To carry out the simulation, we built custom modules into the existing iCasa
-smart home simulator\cite{lalanda_self-aware_2017}. This simulation platform
-allows for simulating autonomic systems, handling internal communications,
-injection of new components at runtime, deletion or change of existing
-components. As such, we used a basic scenario consisting of a house of 4 rooms,
-a single user, and an outdoor zone. All four rooms of the house are equipped
-with a temperature controller system, monitoring and controlling heaters (fig.
+smart home simulator\cite{lalanda_self-aware_2017}. ICasa
+offers a simulation of autonomic systems that can handle internal communications,
+the possible insertion of new components at runtime, or the deletion or modification of existing
+components. We used a basic scenario consisting of a house with four rooms,
+a single user, and an outdoor zone. All four rooms are equipped
+with a temperature controller system that monitors and controls heaters (fig.
 \ref{fig:view}).
@@
 
-Using this basis, we implemented a scenario spanning over 420 days, and
-comprising a daily cycle of outdoor weather (temperature and sunlight), as well
+Based on this, we implemented a scenario spanning over 420 days, and
+comprising a daily cycle of outdoor weather (temperature and sunlight) fluctuations, as well
 as user's movements. All these daily changes create non-noticeable events,
-serving as a background noise for our experiments. To produce outstanding
-events, we randomly generated around twenty events, spanning over the whole
+serving as a background for our experiments. To produce outstanding
+events, we randomly generated about twenty events, spanning over the whole
 duration of the simulation, of different kinds:
@@
         low temperatures.
-    \item Heater failures: heater can fail, making them turning off regardless
+\item Heater failures: heater may break down, making them turn off regardless
         of the command they receive.
-    \item User's vacation: the user go out of the building for an extended
+\item User's absence: the user goes out of the building for an extended
         period of time.
@@
     controlled temperature deviates from its setting (21°C). We will evaluate if
-  our system finds these deviations as memorable, and if it can propose causal
-  hypotheses.}
+our system finds these deviations as memorable, and if it can turn them into relevant causal
+hypotheses for current anomalies.}
   \label{fig:ts_example}
@@
 
-The values of all devices and zones variables was regularly monitored throughout
-the simulation run, and the resulting data, which an excerpt is shown in figure
-\ref{fig:ts_example} can later be used as a basis for our experiments.
+The values observed from all devices and zones was sampled throughout
+the simulation run. The resulting data (figure
+\ref{fig:ts_example}) was then used as a basis for our experiments.
 
@@
 characterize events from the time series data generated by the iCasa simulation.
-Since this is not the focus point of our present work (see sec.
-\ref{sec:related}), we simply apply threshold and pre-computed conditions based
-detection to create a set of events.
+Since the selection of events is not the focus point of our present work (see sec.
+\ref{sec:related}), we perform event detection merely based on threshold and pre-computed conditions.
 
@@
 for computations.
-We then implemented a small number of predicate concepts:
+Then we implemented a small number of predicate concepts:
 \begin{itemize}
 \item $\mathtt{label(e, k)}$: whether the event $e$ has the label ranked
-        $k^{th}$ in the memory.
-\item $\mathtt{rank(e, a_1, a_2)}$: whether the event $e$ has the rank $a_1$
-along axis $a_2$.
+$k^{th}$ in the label list.
+\item $\mathtt{rank(e, r, a)}$: whether the event $e$ has the rank $r$
+along axis $a$.
 \item $\mathtt{day(e, k)}$: whether the event $e$ occurred $k$ days ago.
@@
 
-With a straightforward implementation of memory, predicates and filters, the
-implementation of alg. \ref{alg:complex_iter} worked, but, as expected, took too
+With a straightforward implementation of memory, predicates and filters, we could run alg. \ref{alg:complex_iter} worked, though, as expected, it took too
 long to be usable in realistic scenarios with hundreds or thousands of events to
@@
 for faster operations during future filtering. For instance, since the memory
-  object keeps a mapping from timestamps to events, which allows to quickly
-  filter by date without having to loop over each stored element. This mapping,
-  however, is not directly used to retrieve an event from the outside, as to
-  preserve the theoretical model of memory as an unordered set presented in
+object keeps a mapping from timestamps to events one can perform a quick
+filtering by date without having to loop over all stored element. This convenient mapping,
+however, is not directly used to retrieve events by their date of occurrence, so as to
+preserve the theoretical model of memory as an unordered set, as presented in
 section \ref{sec:computing}.
@@
 
-  \item Some of the filters, for instance date or rank filters, were
-  hard-written to select from the pre-computed mappings of the memory objects
-  rather than testing a predicate over all memory elements.
+\item Some of the filters, for instance the date and rank filters, were
+hard-written. Events can be selected from these pre-computed mappings over the memory objects
+rather than by testing a predicate over all memory elements.
 \end{itemize}
@@
 
-A main sequence of ``usual'' events, which complexity is roughly a logarithm of
-the elapsed time since their occurrence, is visible. This corresponds to events
+The blue line emerges from a set of ``usual'' events, whose complexity varies as the logarithm of
+the elapsed time since their occurrence. It corresponds to events
 for which the best retrieval path consists of a time description (e.g. ``2
@@
 complexity: some appear simpler, as they can be distinguished by using their
-rank alongside an axis (``the hottest day'', ``the second longest user's
+rank along some axis (``the hottest day'', ``the second longest user's
 absence''), or the rare occurrence of their kind (``the only fault on the heater'').
@@
   \includegraphics[width=\linewidth]{figures/complexities_computed}
-  \caption{The computed complexities of events with at most retrieval paths of
+\caption{Computed description complexities of events with retrieval paths of
     length at most 4. Events of type ``day'' (blue), ``hot'' (green), ``cold''
@@
 
-Computing the ``memorability score'', which is shown in fig. \ref{fig:result1},
-highlights these aforementioned events from the main ``usual'' sequence. Since
-the computation of this measure treats unusually complex or simple events the
+The ``memorability score'' is shown in fig. \ref{fig:result1}.
+% highlights these aforementioned events from the main ``usual'' sequence. Since
+the computation of this measure treats unusually complex or unusually simple events the
 same way (from the absolute value operation in Equation \ref{eq:unexpected}), we
@@
 a temperature anomaly occurring simultaneously to many other anomalies is
-notable, as it is costlier than expected to distinguish it from its neighbors.
+noticeable, as it is costlier than expected to distinguish it from its neighbors.
 
@@
   \includegraphics[width=\linewidth]{figures/complexities_surprises}
-  \caption{Memorability score for events in the memory}
+\caption{Memorability score for events in memory}
   \label{fig:result1}
@@
 flag all perturbation events from the usual daily events and evaluate how a
-detection based on ``memorability'' score would succeed in distinguishing these
+detection based on ``memorability'' score would perform in distinguishing these
 events. The result is presented as a ROC curve in fig. \ref{fig:roc}.
 
+\textbf{TOOD:} Il faudrait commenter la courbe
 \begin{figure}[ht]
@@
 Rate) for a classifier based on our memorability score. Measures consider 23
-   manually flagged events as memorable (events added to the background noise
+manually flagged events as memorable (events added to the normal background
 as described in Section~\ref{sec:example}.)}
@@
 
-As an illustration of the abductive inference possible by using the memorability
-score, we used as a target consequence the temperature drop visible in the raw
+As an illustration of the abductive inference made possible our memorability
+score, we used as a target anomaly the temperature drops visible in the raw
 time series data from our scenario (fig \ref{fig:ts_example}).

### 2021-10-14 new_paper.tex
switching to MDPI template

--- 2021-10-04 paper.tex
@@
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{url}
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage[ruled,linesnumbered]{algorithm2e}
-\usepackage{xcolor}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
-\begin{document}
-\title{
-What should I notice? Evaluating the memorability of events for abductive inference.
-}
-\author{
-  \IEEEauthorblockN{
-\IEEEauthorrefmark{1}\IEEEauthorrefmark{2} Étienne Houzé,
-\IEEEauthorrefmark{1} Jean-Louis Dessalles,
-\IEEEauthorrefmark{1} Ada Diaconescu,
-\IEEEauthorrefmark{2} David Menga
+%  LaTeX support: latex@mdpi.com 
+%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.
+
+%=================================================================
+\documentclass[entropy,article,submit,moreauthors,pdftex]{Definitions/mdpi}
+
+% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal and change "submit" to "accept".
+The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
+This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
+For preprints.org, the editorial staff will make this change immediately prior to posting.
+
+%--------------------
+% Class Options:
+%--------------------
+%----------
+% journal
+%----------
+% Choose between the following MDPI journals:
+% acoustics, actuators, addictions, admsci, adolescents, aerospace, agriculture, agriengineering, agronomy, ai, algorithms, allergies, analytica, animals, antibiotics, antibodies, antioxidants, appliedchem, applmech, applmicrobiol, applnano, applsci, arts, asi, atmosphere, atoms, audiolres, automation, axioms, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomechanics, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, brainsci, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coatings, colloids, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, crops, cryptography, crystals, curroncol, cyber, dairy, data, dentistry, dermato, dermatopathology, designs, diabetology, diagnostics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entropy, environments, environsciproc, epidemiologia, epigenomes, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forensicsci, forests, fractalfract, fuels, futureinternet, futuretransp, futurepharmacol, futurephys, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, hazardousmatters, healthcare, hearts, hemato, heritage, highthroughput, histories, horticulturae, humanities, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijtm, ijtpp, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jzbg, kidney, land, languages, laws, life, liquids, literature, livers, logistics, lubricants, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, metabolites, metals, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, minerals, mining, modelling, molbank, molecules, mps, mti, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photochem, photonics, physchem, physics, physiolsci, plants, plasma, pollutants, polymers, polysaccharides, proceedings, processes, prosthesis, proteomes, psych, psychiatryint, publications, quantumrep, quaternary, qubs, radiation, reactions, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, solids, sports, standards, stats, stresses, surfaces, surgeries, suschem, sustainability, symmetry, systems, taxonomy, technologies, telecom, textiles, thermo, tourismhosp, toxics, toxins, transplantology, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wevj, women, world
+
+%---------
+% article
+%---------
+% The default type of manuscript is "article", but can be replaced by: 
+% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
+% supfile = supplementary materials
+
+%----------
+% submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
+Also, line numbering will be removed.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+
+%---------
+% pdftex
+%---------
+% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
+
+%=================================================================
+% MDPI internal commands
+\firstpage{1} 
+\makeatletter 
+\setcounter{page}{\@firstpage} 
+\makeatother
+\pubvolume{1}
+\issuenum{1}
+\articlenumber{0}
+%\doinum{}
+\pubyear{2021}
+\copyrightyear{2021}
+%\externaleditor{Academic Editor: Firstname Lastname} % For journal Automation, please change Academic Editor to "Communicated by"
+\datereceived{} 
+\dateaccepted{} 
+\datepublished{} 
+\hreflink{https://doi.org/} % If needed use \linebreak
+%------------------------------------------------------------------
+% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
+%\pdfoutput=1
+
+%=================================================================
+% Add packages and commands here.
+The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
+\usepackage{algorithm2e}
+
+%=================================================================
+%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
+
+%=================================================================
+% Full title of the paper (Capitalized)
+\Title{What should I notice? Evaluating the memorability of events for abductive inference.
 }
 
-  \IEEEauthorblockA{\IEEEauthorrefmark{1}
-    Télécom Paris, IP Paris, \emph{Palaiseau}, France\\
-email: \{first\}.\{second\}@telecom-paris.fr}
-  \IEEEauthorblockA{\IEEEauthorrefmark{2} EDF R\&{}D, \emph{Palaiseau}, France\\
-email: \{first\}.\{second\}@edf.fr}
-}
+% MDPI internal command: Title for citation in the left column
+\TitleCitation{What should I notice?}
 
-\maketitle
+% Author Orchid ID: enter ID or remove command
+\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
+%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
 
-\begin{abstract}
-  When confronted to an unprecedented situation, humans typically show good
+% Authors, for the paper (add full first names)
+\Author{Étienne Houzé $^{1,2}$*, Jean-Louis Dessalles $^{2}$, Ada Diaconescu$^{2}$ and David Menga $^{2}$}
+
+% MDPI internal command: Authors, for metadata in PDF
+\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}
+
+% MDPI internal command: Authors, for citation in the left column
+\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
+% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address{%
+$^{1}$ \quad EDF R\&D ; \{first\}.\{last\}@edf.fr;
+7 boulevard Gaspard Monge, 91120 Palaiseau, France\\
+$^{2}$ \quad Télécom Paris; \{first\}.\{last\}@telecom-paris.fr;
+19 place Marguerite Perey, 91120 Palaiseau, France}
+
+% Contact information of the corresponding author
+\corres{Correspondence: etienne.houze@telecom-paris.fr}
+
+% Current address and/or shared authorship
+% \firstnote{Current address: Affiliation 3} 
+% \secondnote{These authors contributed equally to this work.}
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+%\simplesumm{} % Simple summary
+
+%\conference{} % An extended version of a conference paper
+
+% Abstract (Do not insert blank lines, i.e. \\) 
+\abstract{ When confronted to an unprecedented situation, humans typically show good
 performance in quickly identifying noticeable past events and proposing them
@@
 Theory. We also provide an implementation example, a
-smart-home scenario, to show how our approach could translate in practice.
-\end{abstract}
+smart-home scenario, to show how our approach could translate in practice.}
 
-\begin{IEEEkeywords}
-Kolmogorov complexity, Algorithmic Information Theory, Simplicity, Abduction, Surprise
-\end{IEEEkeywords}
+% Keywords
+\keyword{Kolmogorov Complexity; Algorithmic Information Theory; Simplicity; Abduction; Surprise}
+
+% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
+%\PACS{J0101}
+%\MSC{}
+%\JEL{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Diversity
+%\LSID{\url{http://}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Applied Sciences:
+%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Data:
+%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
+If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal.
+In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
+
+%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Toxins
+%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Encyclopedia
+%\encyclopediadef{Instead of the abstract}
+%\entrylink{The Link to this entry published on the encyclopedia platform.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%The order of the section titles is: Introduction, Materials and Methods, Results, Discussion, Conclusions for these journals: aerospace,algorithms,antibodies,antioxidants,atmosphere,axioms,biomedicines,carbon,crystals,designs,diagnostics,environments,fermentation,fluids,forests,fractalfract,informatics,information,inventions,jfmk,jrfm,lubricants,neonatalscreening,neuroglia,particles,pharmaceutics,polymers,processes,technologies,viruses,vision
 
 \section{Introduction}
-
-% As the number of connected devices and sensors grows, so does the amount of
-% produced data. As a consequence, the ability to filter out a few events from past recordings
-% as ``memorable'' becomes prime. While this task is often overlooked by many
-% statistical approaches, which instead benefit from large quantity of
-% information, it is part of the learning process for humans
-% \cite{dehaene_how_2020}. Faced with new situations, humans are prone to infer
-% previously seen unusual events as possible causes, and carry out further testing
-% to assess the existence of a causal relation.
 
@@
 
-\section{Theoretical Background}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Materials and Methods}
+
+\subsection{Theoretical Background}
 \label{sec:theory}
@@
 
-\section{Computing the memorability of events}
+\subsection{Computing the memorability of events}
 \label{sec:computing}
-\subsection{Retrieving an event}
+\subsubsection{Retrieving an event}
 
@@
     \pi : \begin{cases}
-\mathcal{M}\times \{0,1\}^{*} &\mapsto \{0,1\} \\
-(e, k) &\mapsto \pi_{k}(e)
+        \mathcal{M}\times \{0,1\}^{*} & \mapsto \{0,1\}    \\
+        (e, k)                        & \mapsto \pi_{k}(e)
     \end{cases}
@@
 
-\subsection{Description complexity of events}
+\subsubsection{Description complexity of events}
 
@@
 
-\subsection{Computing Memorability}
+\subsubsection{Computing Memorability}
 As stated in Equation \ref{eq:memorability}, we define memorability $M(e)$ as the absolute difference between the description complexity of an event and its expected value.
@@
 
-\subsection{From memorability to abduction}
+\subsubsection{From memorability to abduction}
 % On peut étendre la définition utilisée pour la surprise en faisant intervenir
@@
 
-\section{Implementation Example}
-\label{sec:example}
+This metrics answers the different problematics exposed in the introduction: by using a universal measure for complexity, bits, it allows to compare values from different dimensions.
+For instance, it solves the recency problem exposed above: is a big event a long time ago more memorable than a smaller one only a few minutes ago?
+The answer from conditional memorability is that "it depends".
+It depends on the complexity cost set for the predicates describing time and the magnitude of the event.
+These complexity costs should reflect the user's perception of the different dimensions (see Section \ref{sec:future}).
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
+
 \subsection{Setup}
-We considere here an experimental smart home setup. This choice of
+
+We design two different setups to test our approach. Both are inspired from smart home use cases.  This choice of
 configuration is motivated by the challenges posed by smart homes for abductive
@@
 
-% In order to provide an acceptable setup to our experiments, we used the scenario
-% of smart homes. This kind of scenario is a prime example of situations where
-% innovative methods of abduction can prove useful, for various reasons: i) as the
-% number of sensors grows with the number of equipped devices in the house, not
-% all recorded events are useful and should be remembered over long period of
-% time; ii) atypical situations, which are the ones where abduction is most likely
-% to be used (to explain situations to the user, or to solve a conflict), are also
-% where the lack of past data makes knowledge acquisition hard.
+\subsubsection{Scenario 1}
+In this setup, we aim to reproduce the example mentioned in Section \ref{sec:introduction}: the installation of a brand new smart tv causes unpredicted effects on the light of a room.
+To play this situation, we created a set of events covering a period of 100 days, corresponding to the past knowledge of the house.
+Two kinds of events are recorded: ``tv event'', corresponding to TV use, and luminosity events, describing the luminosity of the room at a given time.
+Low lights occur at night, and can occasionally occur during daytime, with a small probability.
+On the 100\textsuperscript{th} day, a ``tv event'' is recorded with a different ``device'' characteristic.
+Shortly after, the light lowers, which is recorded in the following ``light'' event.
 
+\subsubsection{Scenario 2}
+We considere here an experimental smart home setup, with various sensors, which we simulate over a period of time
 To carry out the simulation, we built custom modules into the existing iCasa
@@
 \subsection{Results}
-
+\label{sec:example}
 \subsubsection{Memorability}
@@
 
-The ``memorability score'' is shown in fig. \ref{fig:result1}.
+The ``memorability score'' is shown in Figure \ref{fig:result1}.
 % highlights these aforementioned events from the main ``usual'' sequence. Since
@@
 detection based on ``memorability'' score would perform in distinguishing these
-events. The result is presented as a ROC curve in fig. \ref{fig:roc}.
+events. The result is presented as a ROC curve in Figure \ref{fig:roc}.
 
-\textbf{TOOD:} Il faudrait commenter la courbe
+\textbf{TODO:} Il faudrait commenter la courbe
 \begin{figure}[ht]
@@
 
+To show the potential application to abductive inference, we use our approach to the first scenario, to see if memorabilty alone can find the brand new TV to be a reasonable cause for the sudden low light.
+
+The result of our algorithm is given in Figure \ref{fig:abduction_res}.
+It shows that the system correctly identifies the TV as being the cause.
+The reason for this choice is that, since the tv's device ID is unique among all other events of type ``tv'', its description complexity is very low.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=.8\linewidth]{./figures/abduction_1_results}
+    \caption{Outputs of the memorability-based abduction module. This represent the 3 events in the memory having the highest memorability score relative to the consequence event.}
+    \label{fig:abduction_res}
+\end{figure}
+
 As an illustration of the abductive inference made possible our memorability
@@
 
-\subsubsection{Discussion}
-\label{ssub:discussion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
 
-\section{Related Works}
+\subsection{Related Works}
 \label{sec:related}
@@
 
-\section{Perspectives}
+\subsection{Perspectives}
 \label{sec:future}
@@
 
-\section{Conclusion}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
+The introduction of description complexity and the subsequent memorability measure allows to compare events of various types with different characteristics.
+As such, it provides a framework to discriminate peculiar events and outliers.
 
-\bibliographystyle{IEEEtran}
-\bibliography{biblio.bib}
+The computation being based on predicates, the memorability score is entirely dependent on their definitions.
+This can be seen as a caveat or a feature of our approach: it makes the measure subjective while enabling the consideration of user's preferences in the definition and choice of predicates.
+This consideration can be the topic of future research and open up potential applications.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}
+
+% Only for the journal Methods and Protocols:
+% If you wish to submit a video article, please do so with any other supplementary material.
+% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontributions{Conceptualization, É.
+Houzé and J-L.
+Dessalles; methodology, software, validation, writing--original draft, É.
+Houzé; supervision, writing--review and editing J-L.
+Dessalles, A.
+Diaconescu and D.
+Menga.
+All authors have read and aggreed to the submitted version of the manuscript.
+}
+
+\funding{\textbf{TODO je ne sais pas quoi mettre ici ? Indiquer le financement de la thèse ?}}
+
+\dataavailability{All code and data used for the experiments can be found at \url{https://github.com/EtienneHouze/memorability_code}.
+The iCasa smart home simulator from the Adele research Group, which was used to generate sensor data, can be found at: \url{http://adeleresearchgroup.github.io/iCasa/snapshot/index.html}}
+
+% \acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
+This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
+
+\conflictsofinterest{The authors declare no conflict of interest.}
+
+%% Optional
+% \sampleavailability{Samples of the compounds ... are available from the authors.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Only for journal Encyclopedia
+%\entrylink{The Link to this entry published on the encyclopedia platform.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+% \abbreviations{Abbreviations}{
+%     The following abbreviations are used in this manuscript:\\
+
+%     \noindent
+%     \begin{tabular}{@{}ll}
+%         MDPI & Multidisciplinary Digital Publishing Institute \\
+%         DOAJ & Directory of open access journals              \\
+%         TLA  & Three letter acronym                           \\
+%         LD   & Linear dichroism
+%     \end{tabular}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+% \appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A").
+If the appendix sections contain a heading then change the argument to "yes".
+% \appendixstart
+% \appendix
+% \section{}
+% \subsection{}
+% The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data.
+Mathematical proofs of results not central to the paper can be added as an appendix.
+
+% \begin{specialtable}[H]
+%     \small
+%     \caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab2}}
+%     \begin{tabular}{ccc}
+%         \toprule
+%         \textbf{Title 1} & \textbf{Title 2} & \textbf{Title 3} \\
+%         \midrule
+%         Entry 1          & Data             & Data             \\
+%         Entry 2          & Data             & Data             \\
+%         \bottomrule
+%     \end{tabular}
+% \end{specialtable}
+
+% \section{}
+% All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\end{paracol}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% To add notes in main text, please use \endnote{} and un-comment the codes below.
+%\begin{adjustwidth}{-5.0cm}{0cm}
+%\printendnotes[custom]
+%\end{adjustwidth}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\reftitle{References}
+
+% Please provide either the correct journal abbreviation (e.g.
+according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
+% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
+
+%=====================================
+% References, variant A: external bibliography
+%=====================================
+\externalbibliography{yes}
+\bibliography{biblio}
+
+%=====================================
+% References, variant B: internal bibliography
+%=====================================
+% \begin{thebibliography}{999}
+%     % Reference 1
+%     \bibitem[Author1(year)]{ref-journal}
+%     Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
+%     % Reference 2
+%     \bibitem[Author2(year)]{ref-book1}
+%     Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor1, F., Editor2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
+%     % Reference 3
+%     \bibitem[Author3(year)]{ref-book2}
+%     Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
+%     % Reference 4
+%     \bibitem[Author4(year)]{ref-unpublish}
+%     Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} stage of publication (under review; accepted; in~press).
+%     % Reference 5
+%     \bibitem[Author5(year)]{ref-communication}
+%     Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
+%     % Reference 6
+%     \bibitem[Author6(year)]{ref-proceeding}
+%     Author 1, A.B.; Author 2, C.D.; Author 3, E.F.
+Title of Presentation.
+In Title of the Collected Work (if available), Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference; Editor 1, Editor 2, Eds.
+(if available); Publisher: City, Country, Year (if available); Abstract Number (optional), Pagination (optional).
+%     % Reference 7
+%     \bibitem[Author7(year)]{ref-thesis}
+%     Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
+%     % Reference 8
+%     \bibitem[Author8(year)]{ref-url}
+%     Title of Site. Available online: URL (accessed on Day Month Year).
+% \end{thebibliography}
+
+% If authors have biography, please use the format below
+%\section*{Short Biography of Authors}
+%\bio
+%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
+%{\textbf{Firstname Lastname} Biography of first author}
+%
+%\bio
+%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
+%{\textbf{Firstname Lastname} Biography of second author}
+
+% The following MDPI journals use author-date citation: Admsci,  Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Jintelligence, JRFM, Languages, Laws, Literature, Religions, Risks, Social Sciences.
+For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
+% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
+% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
+328; \citeyear{ref-journal-3b}, p.475).
+This produces: Wong (1999, p.
+328; 2000, p.
+475)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% for journal Sci
+%\reviewreports{\\
+%Reviewer 1 comments and authors’ response\\
+%Reviewer 2 comments and authors’ response\\
+%Reviewer 3 comments and authors’ response
+%}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2021-10-14 new_paper.tex
Added conclusion, the two examples, and some other minor corrections

@@
 
-\textbf{TODO:} Mettre un passage ici pour dire que ça résoud le dilemne de la récence présenté dans l'intro
-
 This metrics answers the different problematics exposed in the introduction: by using a universal measure for complexity, bits, it allows to compare values from different dimensions.
@@
 \label{sec:example}
-\subsubsection{Memorability}
+
+\subsubsection{The ``TV'' scenario:}
+\paragraph{Memorability}
+The computation of the description complexity measure for the 2400 events recorded in this scenario took around 90 seconds using an i7-8600u laptop.
+The resulting complexities, in Figure \ref{fig:scenar1_complexity}, clearly shows that TV events appear more simple than luminosity events.
+In fact, for most luminosity events, the complexity corresponds to the random choice complexity, 25 bits in this example.
+This pattern comes from the number of luminosity events per day, 24, is too high to consider the retrieval path $day(k) \rightarrow randomPick()$ simpler than $randomPick$ from the entire memory, except for the first more recent days, which appear simpler (on the right-hand side of Figure \ref{fig:scenar1_complexity})
+\begin{figure}
+    \centering
+    \includegraphics[width=.7\linewidth]{figures/memorability_scenar_1.png}
+    \caption{Description complexity for events recorded in the TV scenario. Luminosity events are shown in blue, TV events in green}
+    \label{fig:scenar1_complexity}
+\end{figure}
+\paragraph{Abduction}
+To show the potential application to abductive inference, we use our approach to the first scenario, to see if memorabilty alone can find the brand new TV to be a reasonable cause for the sudden low light.
+
+The result of our algorithm is given in Figure \ref{fig:abduction_res}.
+It shows that the system correctly identifies the TV as being the cause.
+The reason for this choice is that, since the tv's device ID is unique among all other events of type ``tv'', its description complexity is very low.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=.8\linewidth]{./figures/abduction_1_results}
+    \caption{Outputs of the memorability-based abduction module. This represent the 3 events in the memory having the highest memorability score relative to the consequence event.}
+    \label{fig:abduction_res}
+\end{figure}
+\subsubsection{The ``temperature'' scenario:}
+\paragraph{Memorability}
 
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/complexities_computed}
+\includegraphics[width=.8\linewidth]{figures/complexities_computed}
     \caption{Computed description complexities of events with retrieval paths of
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/complexities_surprises}
+\includegraphics[width=.8\linewidth]{figures/complexities_surprises}
     \caption{Memorability score for events in memory}
@@
 detection based on ``memorability'' score would perform in distinguishing these
-events. The result is presented as a ROC curve in Figure \ref{fig:roc}.
-
-\textbf{TODO:} Il faudrait commenter la courbe
+events.
+The result is presented as a ROC curve in Figure \ref{fig:roc}.
+This shows the memorability score's performance as a classifying tool for ``out-of-the norm'' events.
+In this example, memorability alone correctly identified 18 manually generated events with only 2 false-positives.
+While the direct application of memorability for classification or event detection is not within the scope of this paper (see Section \ref{sec:related} for complexity-based detection), this first result is on par with the motivation of memorability being in accordance with the intuitive notion.
+
 \begin{figure}[ht]
     \centering
-    \includegraphics[width=\linewidth]{./figures/roc}
+\includegraphics[width=0.7\linewidth]{./figures/roc}
     \caption{Experimental ROC curve (True Positive Rate against False Positive
@@
 
-To show the potential application to abductive inference, we use our approach to the first scenario, to see if memorabilty alone can find the brand new TV to be a reasonable cause for the sudden low light.
-
-The result of our algorithm is given in Figure \ref{fig:abduction_res}.
-It shows that the system correctly identifies the TV as being the cause.
-The reason for this choice is that, since the tv's device ID is unique among all other events of type ``tv'', its description complexity is very low.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=.8\linewidth]{./figures/abduction_1_results}
-    \caption{Outputs of the memorability-based abduction module. This represent the 3 events in the memory having the highest memorability score relative to the consequence event.}
-    \label{fig:abduction_res}
-\end{figure}
-
+\paragraph{Abduction}
 As an illustration of the abductive inference made possible our memorability
@@
 \label{sec:future}
-The main purpose of the present paper is to show the possible connections
-between existing definitions of simplicity from cognitive science, Algorithmic
-Information Theory and a practical use case in cyber-physical systems. As such,
-many further improvements can be done to pave the way towards a better
-integration and performance for anomaly detection or abduction.
+
+% L'utilisation d'une mesure unique de la mémorabilité d'événements de natures et caracteristiques diverses 
+% La mise en application pratique des notions théoriques de mémorabilité et mémorabilité relative nécessité des développements futurs.
+Dans cette section, nous en avons sélectionné 2 qui nous semblent, à ce jour, les plus préoccupants.
+The practical application of the theoretical notions of memorability and relative memorability requires future developments.
+In this section, we have selected two of them which seem to us, to date, the most challenging.
 
@@
 \section{Conclusions}
-The introduction of description complexity and the subsequent memorability measure allows to compare events of various types with different characteristics.
-As such, it provides a framework to discriminate peculiar events and outliers.
+The introduction of description complexity and the subsequent memorability measure allow to compare events of various types with different characteristics.
+As such, it provides a formal and generic framework to discriminate peculiar events and outliers.
 

### 2021-10-14 new_paper.tex
expanding results and related

@@
     \centering
-    \includegraphics[width=\linewidth]{figures/simulator}
+\includegraphics[width=.8\linewidth]{figures/simulator}
     \caption{View of the simulator's web interface provided by iCasa. The four
@@
 \ref{fig:ts_example}) was then used as a basis for our experiments.
-
-\subsection{Implementing the complexity computation}
-
-For the implementation of our method, we first needed to identify and
-characterize events from the time series data generated by the iCasa simulation.
+We then process the time series data to identify and characterize events.
 Since the selection of events is not the focus point of our present work (see sec.
@@
 
-This set of events constitutes the basis of the initial memory $M_0$ used
-for computations.
-Then we implemented a small number of predicate concepts:
+\subsection{Implementing the complexity computation}
+
+We implemented the computation of both the descritpion complexity and the memorability score into a Python object, called the \texttt{SurpriseAbductionModule}.
+Apart from the base memory of events $M_0$, this module contains a set of predifined predicates $\Pi$ to characterize events.
+For instance, for scenario 2, the predicates we use are the following:
 \begin{itemize}
@@
 \end{itemize}
+
+The description length $L_{\pi, k}$ of using a predicate is computed as follows: since the set of predicates is finite and known, $L(\pi) = \log2(|\Pi|)$ are enough to describe the predicate concept $\pi$\footnote{This approach gives an equal complexity to all predicate concepts.
+While this could be argued when using many concepts, as humans do, we deemed better, for our small examples, to consider all concepts as equal.}.
+To encode the argument $k$ of the predicate, we used the prefix-free Elias delta code, which requires $L(k) = \log2(k) + 2 \log2(\log2(k)+1) + 1]$ bits.
+The total cost of decribing $\pi_{k}$ therefore is
+\begin{equation}
+    \L(\pi, k) = \log2(|\Pi|) + \log2(k) + 2 \log2(\log2(k) + 1) + 1
+\end{equation}
 
@@
 \end{figure}
+
 \subsubsection{The ``temperature'' scenario:}
@@
 for which the best retrieval path consists of a time description (e.g. ``2
-months and 12 days ago''). On the other hand, some events stand out in terms of
+months and 12 days ago'').
+As such, it appears that most days are considered usual by our memorability score: therie is no better way to retrieve them than simply giving the day of their occurence.
+On the other hand, some events stand out in terms of
 complexity: some appear simpler, as they can be distinguished by using their
@@
 The ``memorability score'' is shown in Figure \ref{fig:result1}.
-% highlights these aforementioned events from the main ``usual'' sequence. Since
-the computation of this measure treats unusually complex or unusually simple events the
+Similar to what happened with the description complexity score, most events appear with a low memorability: this corresponds mostly to events from the ``main sequence'' from Figure.
+\ref{fig:computed_cplx}.
+On the other hand, some events stand out : for instance event \textbf{TODO}, which are the hottest and coldest days recorded, or event \textbf{TODO} which correspond to the rare type \emph{device\_removal}.
+Since our memorability measure treats unusually complex or unusually simple events the
 same way (from the absolute value operation in Equation \ref{eq:unexpected}), we
 observe that some events are memorable due to their context only. For instance,
-a temperature anomaly occurring simultaneously to many other anomalies is
-noticeable, as it is costlier than expected to distinguish it from its neighbors.
+the group \textbf{TODO} appears more complex than expected: the same event occuring simultaneously in all four rooms of the house make each instance harder to discern. Table \ref{tab:predicates}.
 
@@
 \end{figure}
+
+\begin{table}
+    \centering
+    \begin{tabular}{c|c|c}
+        Event iD & Event Type    & Retrieval Path                         \\
+        \hline
+        1        & hot           & Label("hot"), AxisRank(1)              \\
+        2        & hot           & Label("hot"), Day(21), Location(room2) \\
+        102      & deviceRemoval & Label("deviceRemoval), Day(2)          \\
+    \end{tabular}
+    \caption{Selected events from the ``temperature example'', with their shortest retrieval path.}
+\end{table}
 
@@
 \end{figure}
-\subsubsection{Abduction}
 
@@
 
-\textbf{TODO: the experiment still needs to be done: some additional data is required here (creating the scenario and finding causes.)}
+\textbf{TODO: the figure is still missing!}
 
@@
 \label{sec:related}
-\textbf{TODO: this section is still WIP: I've added most references and short comments, but I still need to write the story linking them.}
 Our work is intended to be integrated into larger-scale frameworks to monitor
@@
 
+The relation between complexity, compression and causality was used in \cite{tatti_finding_2008} to devise the PACK algorithm.
+It models a dataset by using a family of decision trees where each tree describes how one variable can be expressed given the others.
+By chosing the model minimizing the total description length (i.e.
+description of the model and description of the errors), PACK compresses the dataset while finding relations between variables which can be further analyzed.
 More recently, \cite{marx_causal_2018} used Minimum Description Lenght to determine, given a joint probability distribution over $(X,Y)$, whether $X$ causes $Y$ or $Y$ causes $X$.
@@
 
-\cite{tatti_finding_2008} PACK algorithm to cluster data using the simplest possible trees: maybe link this to isolation forests?
-
-The topic of finding events from streams of time series data has already been explored in many ways.
-\cite{aggarwal_outlier_2017} provides good review of modern approaches and techniques
+Another topic where AIT can provide original approaches is event mining in data streams.
+\cite{aggarwal_outlier_2017} provides a good review of modern approaches and techniques
 in the field.
@@
 All these approaches are interesting, and can be integrated into our canvas as tools to detect events using only complexity.
-As such, one could acheive a purely complexity-driven process for detecting and qualifying events.
-
-Among various methods, the approach of ``Isolation forests''\cite{liu_isolation_2008,hariri_extended_2021} is closely related, by design, to ours.
+Using this, one could acheive a purely complexity-driven process for detecting and qualifying events.
+
+The philosphy of our approach can be closely related to the ``Isolation forests'' method\cite{liu_isolation_2008,hariri_extended_2021}.
 It evaluates the isolation of data points by constructing random binary tree classifiers.

### 2021-10-18 new_paper.tex
better examples

@@
 
-% The difficulty of doing so comes from the wide variety of events, parameters and
-% characteristics to take into account. How can one assess if a small recent
-% perturbation is more important than a larger but older one? In addition, one
-% needs a common metrics to evaluate the relative importance of data coming from
-% various devices. In the case of a smart home, for instance, devices range from
-% presence sensors to humidity and temperature sensors, power meters, etc. To this
-% day, no current system proposes to analyze all events coming from such various
-% devices and come to a unified metrics of ``memorability''.
-
-% To tackle this challenge, we use as a starting point the observation that while
-% all events, regardless of their characteristics or nature, can be uniquely
-% described using a combination of qualifiers, the most memorable ones are likely
-% to require less words to be described. Think, for instance, of ``last year's
-% hottest day'' and ``the 182th day of 7 years ago''. By evaluating the lenght of
-% each description, taking into account both the complexity of concept words (a
-% time ranking, a temperature ranking), and the arguments used in both
-% descriptions (the hottest, the 182th, 7), we can assign each event a complexity
-% score. Memorable events, as they stand out, would then differ from their
-% neighbors by being much simpler (or more complex).
-
 To address this issue, we started from the following intuition: while all
@@
 
-% A possible approach would be to rely on the naming complexity of events:
-% memorable events are more likely to be shorter to be named than boring usual
-% events. Think, for instance, of how ``last year's hottest day'' description
-% appears much simpler than ``the 182th day of 7 years
-% ago''\cite{robles_applications_2010}. Our idea is to rely on this metrics to
-% estimate the description complexity of events. Then, by comparing the
-% description complexity of a given event $e$ to the average complexity of similar
-% events, we induce a measure of unexpectedness that can be used to assess the
-% most memorable events.
-
 \begin{figure}[ht]
@@
     \caption{Retrieving an event through successive predicative
-        filters. From the base memory (yellow), successive filters select events satisfying the associated predicate (grey arrows). For example,
+        filters. From the base memory (yellow), successive filters select events satisfying the associated predicate (gray arrows). For example,
         filter $f_1$ selects events from last year, i.e. which satisfy the
@@
 
-% To programmatically achieve this computation, we rely on principles from
-% Algorithmic Information Theory, which provides formal definitions of complexity
-% and simplicity of objects, which appear to be consistent with human perception
-% \cite{li_introduction_2008, dessalles2011coincidences, delahaye_numerical_2012}.
-%  Figure \ref{fig:filters} shows our solution applied to the example description
-%  ``last year's hottest day''. We assume that
-% a memory $\mathcal{M}$ is a unordered set of various distinct objects, upon
-% which can be applied boolean predicates $\pi_{k}$ to describe them. These predicates thus
-% provide filters $f_{\pi, k}$ that can be applied to the memory, selecting all events
-% satisfying predicate $\pi_{k}$. By applying successive
-% filters, we narrow down the size of the memory, until a single event remains.
-% All the successive filters used to come to this singleton memory constitutes a
-% \emph{retrieval path}, whose length, in bits, gives an upper bound of the
-% description complexity of the retrieved event.
-
 We will first briefly introduce
@@
 choice of the Turing machine $U$ used for the computations appears in the
-definition of eq. \ref{eq:kolmo_def}, all results hold, up to an additional
+definition of Equation \ref{eq:kolmo_def}, all results hold, up to an additional
 constant, if we change the machine. Think how any Turing-complete programming language can
@@
 incomputable\cite{li_introduction_2008},
-%The sketch of the proof is the
-%following: since the definition of complexity from Equation. \ref
-%{eq:kolmo_def} requires to find the shortest of all programs on a given Turing
-%machine outputting $s$, it requires to run all programs up to a given length
-%and compare their result to $s$. However, being able to do so means that we
-%would be able to tell whether these programs terminate. Since the termination
-%of programs is famously undecidable, by extension, the computation of
-%Kolmogorov complexity is impossible.
 it can be approximated with
@@
 \end{equation}
-where $l$ is the label, $t$ the timestamp and $\mathcal{D}$ a multi-dimensional data point representing the various characteristics of $e$: its duration, the maximum temperature reached, the sensor name, its position, etc.
+where $l$ is the label, $t$ the timestamp and $\mathcal{D}$ a multidimensional data point representing the various characteristics of $e$: its duration, the maximum temperature reached, the sensor name, its position, etc.
 Labels can also be considered as classes of events, of which each event is a particular instance.
@@
 \subsubsection{From memorability to abduction}
-% On peut étendre la définition utilisée pour la surprise en faisant intervenir
-% les prédicats'' `sympathiques` par rapport à ce que l'on questionne.
 Abductive inference builds on the computation
@@
 
-This metrics answers the different problematics exposed in the introduction: by using a universal measure for complexity, bits, it allows to compare values from different dimensions.
-For instance, it solves the recency problem exposed above: is a big event a long time ago more memorable than a smaller one only a few minutes ago?
+This metrics answers the different problems exposed in the introduction: by using a universal measure for complexity, bits, it allows to compare values from different dimensions.
+For instance, it solves the dilemma of recent events: is a big event a long time ago more memorable than a smaller one only a few minutes ago?
 The answer from conditional memorability is that "it depends".
@@
 \subsubsection{Scenario 1}
-In this setup, we aim to reproduce the example mentioned in Section \ref{sec:introduction}: the installation of a brand new smart tv causes unpredicted effects on the light of a room.
+In this setup, we aim to reproduce the example mentioned in Section \ref{sec:introduction}: the installation of a brand new smart TV causes unpredictable effects on the light of a room.
 To play this situation, we created a set of events covering a period of 100 days, corresponding to the past knowledge of the house.
-Two kinds of events are recorded: ``tv event'', corresponding to TV use, and luminosity events, describing the luminosity of the room at a given time.
+Two kinds of events are recorded: ``TV event'', corresponding to TV use, and luminosity events, describing the luminosity of the room at a given time.
 Low lights occur at night, and can occasionally occur during daytime, with a small probability.
-On the 100\textsuperscript{th} day, a ``tv event'' is recorded with a different ``device'' characteristic.
+On the 100\textsuperscript{th} day, a ``TV event'' is recorded with a different ``device'' characteristic.
 Shortly after, the light lowers, which is recorded in the following ``light'' event.
@@
 \subsubsection{Scenario 2}
-We considere here an experimental smart home setup, with various sensors, which we simulate over a period of time
+We consider an experimental smart home setup, with various sensors, which we simulate over a period of time
 To carry out the simulation, we built custom modules into the existing iCasa
@@
     \caption{Time series data from the simulation: outdoor temperature (blue) and
-        controller temperature of a room (orange). On different occasion, the
-        controlled temperature deviates from its setting (21°C). We will evaluate if
-        our system finds these deviations as memorable, and if it can turn them into relevant causal
-        hypotheses for current anomalies.}
+        controller temperature of a room (orange). To be used in our framework, these time series data are
+        processed by a simple threshold-based event detection.}
     \label{fig:ts_example}
@@
 Since the selection of events is not the focus point of our present work (see sec.
-\ref{sec:related}), we perform event detection merely based on threshold and pre-computed conditions.
+\ref{sec:related}), we perform event detection merely based on threshold and precomputed conditions.
 
@@
 
-We implemented the computation of both the descritpion complexity and the memorability score into a Python object, called the \texttt{SurpriseAbductionModule}.
-Apart from the base memory of events $M_0$, this module contains a set of predifined predicates $\Pi$ to characterize events.
+We implemented the computation of both the description complexity and the memorability score into a Python object, called the \texttt{SurpriseAbductionModule}.
+Apart from the base memory of events $M_0$, this module contains a set of predefined predicates $\Pi$ to characterize events.
 For instance, for scenario 2, the predicates we use are the following:
@@
     \item $\mathtt{day(e, k)}$: whether the event $e$ occurred $k$ days ago.
-    \item $\mathtt{month(e, k)}$: whether the event $e$ occurred $k$ months
-          ago.
+    % \item $\mathtt{randomChoice(e, k)}$: whether the event $e$ id matches the program $k$, which can be regarded as a random seed.
+    \item $\mathtt{month(e, k)}$: whether the event $e$ occured $k$ months ago.
     \item $\mathtt{location(e, k)}$: whether the event $e$ occurred in zone
@@
 
+Similar predicates were used for scenario 1, with the notable exception of the addition of a \texttt{RandomPick} predicate, which is true if the id of the event matches a random program seed $k$.
+Using this predicate and the associated filter in only one of the two illustrative examples shows the importance of allowing random selection in the complexity computation.
+Furthermore, it could be argued that in some case, random selection of an event is not possible.
+
 The description length $L_{\pi, k}$ of using a predicate is computed as follows: since the set of predicates is finite and known, $L(\pi) = \log2(|\Pi|)$ are enough to describe the predicate concept $\pi$\footnote{This approach gives an equal complexity to all predicate concepts.
 While this could be argued when using many concepts, as humans do, we deemed better, for our small examples, to consider all concepts as equal.}.
-To encode the argument $k$ of the predicate, we used the prefix-free Elias delta code, which requires $L(k) = \log2(k) + 2 \log2(\log2(k)+1) + 1]$ bits.
-The total cost of decribing $\pi_{k}$ therefore is
+To encode the argument $k$ of the predicate, we used the prefix-free Elias delta code\cite{elias_universal_1975}, which requires $L(k) = \log2(k) + 2 \log2(\log2(k)+1) + 1]$ bits.
+The total cost of describing $\pi_{k}$ therefore is
 \begin{equation}
@@
           for faster operations during future filtering. For instance, since the memory
-          object keeps a mapping from timestamps to events one can perform a quick
+          object keeps a mapping from timestamp to events one can perform a quick
           filtering by date without having to loop over all stored element. This convenient mapping,
@@
     \item Some of the filters, for instance the date and rank filters, were
-          hard-written. Events can be selected from these pre-computed mappings over the memory objects
+          hard-written. Events can be selected from these precomputed mappings over the memory objects
           rather than by testing a predicate over all memory elements.
@@
 The computation of the description complexity measure for the 2400 events recorded in this scenario took around 90 seconds using an i7-8600u laptop.
-The resulting complexities, in Figure \ref{fig:scenar1_complexity}, clearly shows that TV events appear more simple than luminosity events.
+The resulting complexities, in Figure \ref{fig:scenar1_complexity}, show that TV events appear more simple than luminosity events.
 In fact, for most luminosity events, the complexity corresponds to the random choice complexity, 25 bits in this example.
@@
 \paragraph{Abduction}
-To show the potential application to abductive inference, we use our approach to the first scenario, to see if memorabilty alone can find the brand new TV to be a reasonable cause for the sudden low light.
-
-The result of our algorithm is given in Figure \ref{fig:abduction_res}.
+To show the potential application to abductive inference, we use our approach to the first scenario, to see if memorability alone can find the brand new TV to be a reasonable cause for the sudden low light.
+
+The result of our algorithm is given in Table \ref{tab:abduction_res}.
 It shows that the system correctly identifies the TV as being the cause.
-The reason for this choice is that, since the tv's device ID is unique among all other events of type ``tv'', its description complexity is very low.
-
-\begin{figure}
+The reason for this choice is that, since the smart TV's device ID is unique among all other events of type ``TV'', its description complexity is very low.
+
+% \begin{figure}
+%     \centering
+%     \includegraphics[width=.8\linewidth]{./figures/abduction_1_results}
+%     \caption{Outputs of the memorability-based abduction module. This represent the 3 events in the memory having the highest memorability score relative to the consequence event.}
+%     \label{fig:abduction_res}
+% \end{figure}
+
+\begin{table}
     \centering
-    \includegraphics[width=.8\linewidth]{./figures/abduction_1_results}
-    \caption{Outputs of the memorability-based abduction module. This represent the 3 events in the memory having the highest memorability score relative to the consequence event.}
-    \label{fig:abduction_res}
-\end{figure}
+    \begin{tabular}{l|c|c}
+        Event Id & Description & Relative memorability (bits)\\
+        \hline
+        2513 & Use of smart TV & 16.76 \\
+        2427 & Last use of the old TV & 14.81 \\
+        2411 & Second-last use of the old TV & 11.21 
+    \end{tabular}
+    \caption{Output of the memorability-based abduction module: top 3 events for the relative memorability metrics.}
+    \label{tab:abduction_res}
+\end{table}
 
 \subsubsection{The ``temperature'' scenario:}
-\paragraph{Memorability}
+% \paragraph{Memorability}
 
@@
 months and 12 days ago'').
-As such, it appears that most days are considered usual by our memorability score: therie is no better way to retrieve them than simply giving the day of their occurence.
+As such, it appears that most days are considered usual by our memorability score: there is no better way to retrieve them than simply giving the day of their occurrence.
 On the other hand, some events stand out in terms of
@@
     \centering
-\includegraphics[width=.8\linewidth]{figures/complexities_computed}
+    \includegraphics[width=.9\linewidth]{figures/complexity_scenar_2.png}
     \caption{Computed description complexities of events with retrieval paths of
-length at most 4. Events of type ``day'' (blue), ``hot'' (green), ``cold''
-(red), ``device removal''(cyan) are shown.}
+        length at most 4. Events of type ``day'' (blue), ``hot'' (red), ``cold''
+        (gray), ``device removal''(orange) and ``device addition'' (cyan) are shown. Events mentioned in the text are specified.}
     \label{fig:computed_cplx}
@@
 \ref{fig:computed_cplx}.
-On the other hand, some events stand out : for instance event \textbf{TODO}, which are the hottest and coldest days recorded, or event \textbf{TODO} which correspond to the rare type \emph{device\_removal}.
+On the other hand, some events stand out : for instance events 20 and 329, which are respectively the hottest and coldest days recorded, or event 183 which correspond to the rare type \emph{device\_removal}.
 Since our memorability measure treats unusually complex or unusually simple events the
@@
 observe that some events are memorable due to their context only. For instance,
-the group \textbf{TODO} appears more complex than expected: the same event occuring simultaneously in all four rooms of the house make each instance harder to discern. Table \ref{tab:predicates}.
+the group to which event 149 belongs appears more complex than expected: the same event occurring simultaneously in all four rooms of the house make each instance harder to discern. 
+Table \ref{tab:paths} illustrates this by exhibiting the retrieval paths used for complexity computation for these events.
 
@@
     \centering
-\includegraphics[width=.8\linewidth]{figures/complexities_surprises}
+    \includegraphics[width=.8\linewidth]{figures/memo_scenar_2.png}
     \caption{Memorability score for events in memory}
@@
         \hline
-        1        & hot           & Label("hot"), AxisRank(1)              \\
-        2        & hot           & Label("hot"), Day(21), Location(room2) \\
-        102      & deviceRemoval & Label("deviceRemoval), Day(2)          \\
+        20       & day           & Label("day"), AxisRank(0, "max\_temp")    \\
+        329      & day           & Label("day"), AxisRank(0, "min\_temp")    \\
+        183      & deviceRemoval & Label("deviceRemoval"), Day(0)            \\
+        149      & cold          & Label("cold"), Day(2), Device("thermo\_2") \\
     \end{tabular}
-    \caption{Selected events from the ``temperature example'', with their shortest retrieval path.}
+    \caption{Selected events from the ``temperature example'' with their shortest retrieval path.}
+    \label{tab:paths}
 \end{table}
@@
 
-\paragraph{Abduction}
-As an illustration of the abductive inference made possible our memorability
-score, we used as a target anomaly the temperature drops visible in the raw
-time series data from our scenario (fig \ref{fig:ts_example}).
-
-\textbf{TODO: the figure is still missing!}
+% \paragraph{Abduction}
+% As an illustration of the abductive inference made possible our memorability
+% score, we used as a target anomaly the temperature drops visible in the raw
+% time series data from our scenario (fig \ref{fig:ts_example}).
+
+% \textbf{TODO: the figure is still missing!}
 
@@
 It models a dataset by using a family of decision trees where each tree describes how one variable can be expressed given the others.
-By chosing the model minimizing the total description length (i.e.
+By choosing the model minimizing the total description length (i.e.
 description of the model and description of the errors), PACK compresses the dataset while finding relations between variables which can be further analyzed.
-More recently, \cite{marx_causal_2018} used Minimum Description Lenght to determine, given a joint probability distribution over $(X,Y)$, whether $X$ causes $Y$ or $Y$ causes $X$.
+More recently, \cite{marx_causal_2018} used Minimum Description Length to determine, given a joint probability distribution over $(X,Y)$, whether $X$ causes $Y$ or $Y$ causes $X$.
 Their method is based on tree models, and implying that a model respecting the causal relation will be simpler to describe.
@@
 All these approaches are interesting, and can be integrated into our canvas as tools to detect events using only complexity.
-Using this, one could acheive a purely complexity-driven process for detecting and qualifying events.
-
-The philosphy of our approach can be closely related to the ``Isolation forests'' method\cite{liu_isolation_2008,hariri_extended_2021}.
+Using this, one could achieve a purely complexity-driven process for detecting and qualifying events.
+
+The philosophy of our approach can be closely related to the ``Isolation forests'' method\cite{liu_isolation_2008,hariri_extended_2021}.
 It evaluates the isolation of data points by constructing random binary tree classifiers.
@@
 By using predicates as a proxy for complexity computation, our methods is more general, as it is agnostic regarding the nature of events.
-However, the introdcution of predicates adds a subjectivity in the determination of memorable points, as we will discuss later.
+However, the introduction of predicates adds a subjectivity in the determination of memorable points, as we will discuss later.
 
@@
 
-% L'utilisation d'une mesure unique de la mémorabilité d'événements de natures et caracteristiques diverses 
-% La mise en application pratique des notions théoriques de mémorabilité et mémorabilité relative nécessité des développements futurs.
-Dans cette section, nous en avons sélectionné 2 qui nous semblent, à ce jour, les plus préoccupants.
 The practical application of the theoretical notions of memorability and relative memorability requires future developments.

### 2021-10-18 new_paper.tex
Short update
Before sending it to JL, Ada and David

@@
 % Authors, for the paper (add full first names)
-\Author{Étienne Houzé $^{1,2}$*, Jean-Louis Dessalles $^{2}$, Ada Diaconescu$^{2}$ and David Menga $^{2}$}
+\Author{Étienne Houzé $^{1,2}$*, Jean-Louis Dessalles $^{2}$, Ada Diaconescu$^{2}$ and David Menga $^{1}$}
 
@@
 which, if given as input to a universal Turing Machine $U$, outputs $s$.
+
 \begin{equation}
@@
 finite program $p_{U}$, we have the following inequality:
+
 \begin{equation}
@@
 machine $U$ taken from the set $\text{TM}$ of Turing machines:
+
 \begin{equation}
@@
 \end{equation}
+
 where the additional constant $C_{U}$ does not depend on $s$.
@@
     (e)$.
+
 \begin{equation}
@@
 Using these insights from AIT, we define the memorability $M(e)$ of an event as the absolute difference between the description complexity $K_d(e)$ of an event and its expected description complexity $K_{exp}(e)$:
+
 \begin{equation}
@@
 \end{equation}
+
 Contrary to the definition of unexpectedness from Equation \ref{eq:unexpected}, we use an absolute value: we do so to acknowledge the fact that events more complex than expected can be memorable as well\footnote{In the original paper \cite{dessalles2011coincidences}, exceptionally complex events are described by considering complexity itself as a way to describe the event: see ``the Pisa Tower effect''\cite{dessalles_pisa_nodate}}.
@@
 We define \emph{events} as data points augmented with a \emph{label} indicating their nature (temperature event, failure event, addition/removal of a device) and a timestamp of occurrence. Formally:
+
 \begin{equation}
@@
 \end{equation}
+
 where $l$ is the label, $t$ the timestamp and $\mathcal{D}$ a multidimensional data point representing the various characteristics of $e$: its duration, the maximum temperature reached, the sensor name, its position, etc.
@@
 Using this notation, the predicate $\pi$ becomes a boolean function operating on $\mathbf{E} \times \{0,1\}^*$:
+
 \begin{equation}
@@
 \end{equation}
+
 As an example of predicate, consider $\pi = \mathtt{year}$ and $k$ a string encoding the number $1$, thus
@@
 We call this operation a \emph{filter}:
+
 \begin{equation}
@@
 \end{equation}
+
 For instance, using the same $\pi = \mathtt{year}$ and $k=1$ as above, we can
@@
 functions. A sequence of such filters is called a \emph{retrieval path}
+
 \begin{equation}
@@
 \end{equation}
+
 By definition
@@
 If we limit ourselves to prefix-free strings encoding predicates and arguments, the total bit length is given by:
+
 \begin{align}
@@
           of the command they receive.
-    \item User's absence: the user goes out of the building for an extended
-          period of time.
+    % \item User's absence: the user goes out of the building for an extended
+    %       period of time.
     \item Device removal/addition: a device is removed, or another one is added
@@
 The total cost of describing $\pi_{k}$ therefore is
+
 \begin{equation}
@@
 In fact, for most luminosity events, the complexity corresponds to the random choice complexity, 25 bits in this example.
-This pattern comes from the number of luminosity events per day, 24, is too high to consider the retrieval path $day(k) \rightarrow randomPick()$ simpler than $randomPick$ from the entire memory, except for the first more recent days, which appear simpler (on the right-hand side of Figure \ref{fig:scenar1_complexity})
+This pattern comes from the number of luminosity events per day, 24, is too high to consider the retrieval path $day,  randomPick$ simpler than $randomPick$ from the entire memory, except for the most recent days, which appear simpler (on the right-hand side of Figure \ref{fig:scenar1_complexity})
 \begin{figure}
@@
 \end{figure}
+
 \paragraph{Abduction}
@@
 The results of the description complexity evaluation, for the described setup,
-are shown in fig. \ref{fig:computed_cplx}. The entire computation, over 4
+are shown in Figure \ref{fig:computed_cplx}. The entire computation, over 4
 iterations (meaning that retrieval paths contained at most 4 filters), took
@@
 
-The blue line emerges from a set of ``usual'' events, whose complexity varies as the logarithm of
-the elapsed time since their occurrence. It corresponds to events
-for which the best retrieval path consists of a time description (e.g. ``2
-months and 12 days ago'').
-As such, it appears that most days are considered usual by our memorability score: there is no better way to retrieve them than simply giving the day of their occurrence.
+The general trend appears different from the ``TV example'' presented in Figure \ref{fig:scenar1_complexity} : instead of having a capped complexity score, a logarithmic trend appears.
+This phenomenon is due to the absence, in this example, of the \texttt{randomPick} predicate; instead, usual events, with outstanding particularity, can be retrieved by using the \texttt{day} predicate and a \texttt{label} predicate.
+As such, it appears that most days are considered usual by our memorability score, creating the main logarithmic sequence of blue dots: there is no better way to retrieve them than simply giving the day of their occurrence.
 On the other hand, some events stand out in terms of
 complexity: some appear simpler, as they can be distinguished by using their
-rank along some axis (``the hottest day'', ``the second longest user's
-absence''), or the rare occurrence of their kind (``the only fault on the heater'').
+rank along some axis (``the hottest day'', ``the second coldest day''), or the rare occurrence of their kind (``the only fault on the heater'').
 

### 2021-10-25 new_paper.tex
Ada's comments, pt. 1

@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{ When confronted to an unprecedented situation, humans typically show good
+\abstract{ When confronted with an unprecedented situation, humans typically show good
 performance in quickly identifying noticeable past events and proposing them
@@
 The approach is meant to be complementary to more traditional
-learning-focused abduction techniques. We provide theoretical ground for our approach
-based on Algorithmic Information
-Theory. We also provide an implementation example, a
-smart-home scenario, to show how our approach could translate in practice.}
+learning-focused abduction techniques.
+We provide a theoretical framework based on Algorithmic Information Theory to define the notion of memorability of events.
+Then, we implement our approaches in two examples based on smart-home scenario to validate our approach in toy cases and illustrate how it could be applied.}
 
@@
 
-As a user has just turned on the TV in her all-equipped living-room, the
+As a user has just switch on the TV in her all-equipped living-room, the
 lights dim and the window blinds go down. Intrigued by this behavior, she
@@
 observed consequences, they infer the possible causes.
+% TODO: answer in the next paragraph?
 
@@
 suggests that human beings are able to use distinct methods to perform abductive tasks
-and infer new knowledge. While the first two mechanisms can be
-automated using knowledge bases and statistical methods, ``memorability-based'' abduction needs to be investigated in more depth.
-
-% For instance, it is
-% possible, for humans without any prior knowledge about
-% physics or electrical engineering, to infer a memorable thunderstorm as a
-% possible cause for a general black-out just by using the memorability of
-% the former. While this kind of abductive inference (i.e. inference to the cause
-% of an observation \cite{magnani_abduction_2011}) can yield many
-% false-positives, it can be considered in situations where the lack of knowledge
-% from previous occurrences penalizes other approaches.
-
-% The difficulty of this approach to reasoning is, however, to quickly identify
-% the relevant candidate hypotheses. In this regards, we may use a basic
-% instinct: without prior knowledge, the most relevant hypothesis might simply be
-% the most memorable recent event. However this definition is highly subjective,
-% and does not seem to fit well into the canvas of computing and AI.
+and infer new knowledge, depending on the situation. While the first two mechanisms can be
+automated using knowledge bases and statistical methods,the third approach, which can be used without any knowledge of the occurring phenomenon or past occurrences.
+Instead, it relies on the observation that some events appear more ``memorable'' than others and, as such, can be considered as possible hypotheses if need be.
 
@@
 ``memorability''. 
-
+% ! Addiction of the follwoing sentence, to show what we can use memorability for.
+In addition to the aforementionned use for abductive inference, having access to a computation of memorability would allow a system to summarize a situation to its user by presenting only the most memorable events: for instance, a summary of notable events that occurred during the home owner's absence.
+
+% ? What to do with the "intuition" thing? Keep it? Leave it?
 To address this issue, we started from the following intuition: while all
@@
 of occurrence, a temperature ranking), and of the arguments (1st hottest,
-182 and 7). This approach, The resulting values define the  \emph{description complexity}
+182 and 7). The resulting values define the  \emph{description complexity}
 of events. Our intuition is that memorable events require simpler and less numerous qualifiers to be
 unambiguously described than unremarkable ones.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=\linewidth]{figures/filters}
-    \caption{Retrieving an event through successive predicative
-        filters. From the base memory (yellow), successive filters select events satisfying the associated predicate (gray arrows). For example,
-        filter $f_1$ selects events from last year, i.e. which satisfy the
-        predicate $\mathtt{date}(event, 1\_year)$. In this case, successively
-        applying filters $f_1$, $f_2$ and $f_3$ yields a unique event $e_h$, last
-        year's hottest day. The complexity of this event can then be upper-bounded
-        by the complexity of the three filters, since they give an unambiguous way
-        to describe the event within the memory.}
-    \label{fig:filters}
-\end{figure}
+% ? Response to Ada's remark: added a thing in the 1st paragraph to show what we want ot do with the memorable events.
 
@@
 (\cdot, 1\_year)$ is \emph{true} of events that occurred last year. Selecting
+%%%%%%%%%%%%%%%%%%%%%% ! MOVED FIGURE HERE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure}[hb]
+    \centering
+    \includegraphics[width=.95\linewidth]{figures/filters}
+    \caption{Retrieving an event through successive predicative
+        filters. From the base memory (yellow), successive filters select events satisfying the associated predicate (gray arrows). For example,
+        filter $f_1$ selects events from last year, i.e. which satisfy the
+        predicate $\mathtt{date}(event, 1\_year)$. In this case, successively
+        applying filters $f_1$, $f_2$ and $f_3$ yields a unique event $e_h$, last
+        year's hottest day. The complexity of this event can then be upper-bounded
+        by the complexity of the three filters as they give an unambiguous way
+        to describe the event within the memory.}
+    \label{fig:filters}
+\end{figure}
+%%%%%%%%%%%%%%%%%%%%%%! FIGURE END %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 all events, from the memory, that satisfy a given predicate corresponds to
@@
 
+% * Paragraph re-written
 We will first briefly introduce
-some basic notions of Algorithmic Information Theory. Then, in
-section~\ref{sec:computing}, we show how we compute the
-memorability of past events. We illustrate our approach in~
-(section \ref{sec:example}) by describing an implementation
-in a smart home simulation context. In section~\ref{sec:related} we review
-a few studies that use complexity theory to generate explanations for the smart home
-and we discuss how our work can be linked with them. We eventually explore potential
-extensions of our work in section \ref{sec:future}
+some basic notions of Algorithmic Information Theory in subsection~\ref{sec:theory}.
+We then expose our contribution by presenting in subsection~\ref{sec:computing} our formal definition of memorability and related notions.
+Then, we present an implementation example of these definitions with two smart-home examples in section~\ref{sec:results}.
+The results of these experiments are then presented and discussed.
+Finally we explore other related works in subsection \ref{sec:related} and explore possible extensions of our work in subsection \ref{sec:future}.
+
+% Then, in
+% subsection~\ref{sec:computing}, we show how we compute the
+% memorability of past events. We illustrate our approach in section~\ref{sec:example} by describing an implementation
+% in a smart home simulation context. In section~\ref{sec:related} we review
+% a few studies that use complexity theory to generate explanations for the smart home
+% and we discuss how our work can be linked with them. We explore potential
+% extensions of our work in section \ref{sec:future}
 
@@
 any object represented by a finite binary
-string)\cite{kolmogorov_three_1965,li_introduction_2008}. The complexity $K(s)$ of a (finite) binary string is the length in bits of the shortest program $p$
+string)\cite{kolmogorov_three_1965,li_introduction_2008}. The complexity $K(s)$ of a (finite) binary string is the length in bits $L(p)$ of the shortest program $p$
 which, if given as input to a universal Turing Machine $U$, outputs $s$.
@@
     \label{eq:kolmo_def}
-    K_{U}(s) = \min_{p}\left\{l(p)|U(p)=s\right\}
+    K_{U}(s) = \min_{p}\left\{L(p)|U(p)=s\right\}
 \end{equation}
@@
 \begin{equation}
-    \label{eq:unexpected} U(e) = K_{w}(e) - K(e)
+    \label{eq:unexpected} \textit{Unex}(e) = K_{w}(e) - K(e)
 \end{equation}
@@
 
-Using these insights from AIT, we define the memorability $M(e)$ of an event as the absolute difference between the description complexity $K_d(e)$ of an event and its expected description complexity $K_{exp}(e)$:
+As \cite{dessalles2011coincidences} suggests a link between the unexpectedness and cognitive relevance, we propose to define the memorability of an event in a similar way.  
+Hence, we define the memorability $M(e)$ of an event as the absolute difference between its description complexity $K_d(e)$ and its expected description complexity $K_{exp}(e)$:
 
@@
 
+% ?
+Peut-être ajouter un paragraphe sur pourquoi ne pas directement utiliser la notion de unexpectedness.
+1) pas calculable.
+2) pas forcément bien définie dans un cas pratique d'application (comment définit la complexité causale ???)
 Contrary to the definition of unexpectedness from Equation \ref{eq:unexpected}, we use an absolute value: we do so to acknowledge the fact that events more complex than expected can be memorable as well\footnote{In the original paper \cite{dessalles2011coincidences}, exceptionally complex events are described by considering complexity itself as a way to describe the event: see ``the Pisa Tower effect''\cite{dessalles_pisa_nodate}}.
@@
 \end{equation}
-
-where $l$ is the label, $t$ the timestamp and $\mathcal{D}$ a multidimensional data point representing the various characteristics of $e$: its duration, the maximum temperature reached, the sensor name, its position, etc.
+% ! response to Ada's remark: there is a hidden unique ID in the implementation, to distinguish between events; however this is not used to retrieve the event, as it is not a defining characteristics.
+where $l$ is the label, $t$ the timestamp and $\mathcal{D}$ a vector of properties characterizing $e$: its duration, the maximum temperature reached, the sensor name, its position, etc.
 Labels can also be considered as classes of events, of which each event is a particular instance.
@@
 In the rest of this paper, we will prefer the equivalent notation $\pi(e, k) \mapsto \{0,1\}$, where $k$ is a binary string encoding the sequence of arguments $a_1, \dots, a_n$.
-Using this notation, the predicate $\pi$ becomes a boolean function operating on $\mathbf{E} \times \{0,1\}^*$:
+Using this notation, the predicate $\pi$ becomes a boolean function operating on $\mathbf{E} \times \{0,1\}^*$, where $\mathbf{E}$ denotes the set of all events:
 
@@
     \pi : \begin{cases}
-        \mathcal{M}\times \{0,1\}^{*} & \mapsto \{0,1\}    \\
+        \mathbf{E}\times \{0,1\}^{*} & \mapsto \{0,1\}    \\
         (e, k)                        & \mapsto \pi_{k}(e)
@@
 
+% !
+response to Ada's remark: complexity computations don't consider time complexity, only program length.
+Indeed, using a FILO structure for the memory makes sense implementation-wise, but it does not affect the definitions of complexity and memorability.
 As events occur, they are stored in the  \emph{base memory} $M_0$.
 As they are not directly comparable, the memory $M_0$ can be considered as having the structure of an unordered set.
-We denote by $\mathcal{M}$ the subsets of $M_0$.
+We denote by $\mathcal{M}$ the set of all subsets of $M_0$.
 By extension, elements of $\mathcal{M}$, i.e.
@@
         \mathcal{M} & \mapsto \mathcal{M}           \\
-        M           & \mapsto \{e \in M | \pi(e) \}
+        M           & \mapsto \{e \in M | \pi_k(e) \}
     \end{cases}
@@
 \end{align}
+where $L(\pi_\square)$ and $L(k_\square)$ respectively denote the length, in bits, required to express the predicate's concept and program.
+This length may vary depending of the encoding choice, see Section \ref{sec:results} for an example.
 
@@
 length of the corresponding retrieval path (line 14).
+% !
+response to Ada's comment: yes, this works well ;-) It is even implemented in the code.
+However, given that time complexity does not impact the definitions of complexity, I decided to present the simplest program, not the most optimized, for the sake of clarity.
 
@@
 
-$K_{exp}(e)$ evaluates the complexity the user, or the system, would expect the occurrence of event $e$ to have, based on their previous knowledge.
+$K_{exp}(e)$ evaluates the complexity that the user, or the system, would expect for the occurrence of event $e$ to have, based on their previous knowledge.
 In our framework, this prior knowledge consists of the base memory $M_0$.
@@
 should be considered \emph{similar} events.
-Given that we deal with non comparable events, we may define the notion of similarity by referring once again to \emph{predicates}.For a given event $e$ and a given predicate $\pi_k$, we define a $\pi_k$-neighborhood of $e$ as the set $N_{\pi, k}(e)$ of all other events satisfying $\pi_k$.
+Given that we deal with non comparable events, we may define the notion of similarity by referring once again to \emph{predicates}.
+For a given event $e$ and a given predicate $\pi_k$, we define a $\pi_k$-neighborhood of $e$ as the set $N_{\pi, k}(e)$ of all other events satisfying $\pi_k$.
 
@@
     \label{eq:similar}
-    N_{\pi, k}(e) = \{e'\in M_0, \pi_k(e') \wedge e' \neq e\}
+    N_{\pi, k}(e) = \{e'\in M_0, (e' \neq e) \wedge \pi_k(e')\}
 \end{equation}
@@
 neighborhoods $N_{\pi, k}(e)$, with the convention that $N_{\pi, k}(e) = \emptyset$
-if $e$ does not satisfy $\pi_k$, we can compute an average expected complexity for $e$:
+if $e$ does not satisfy $\pi_k$, we can compute an average expected complexity for $e$, by summing the complexity of events in all neighbourhoods of $e$ and dividing the total by the number of events in the neighborhoods:
 
@@
 the final result.
+% !
+response to Ada's comment: Yes, it is.
+If a point belongs to many neighbourhoods, with many points in each, it will be hard to distinguish from the other points.
+So, its expected complexity is hard.
+Think, for instance, of the expected complexity for a human being: it will be roughly log2(7000000000) bits, for a randomly picked person.
+If this person is known to you, it will be simpler to describe her.
 
 \subsubsection{From memorability to abduction}
-Abductive inference builds on the computation
+% TODO : see ada's comment for this part
+Abductive inference builds upon the computation
 of the memorability score. \emph{Knowing} that we want to find a cause $c$
@@
 The knowledge attached to the occurrence of $e$ can be integrated into the description complexity definition by using conditional complexity $K_d(c | e)$: The information contained in $e$ is considered as given, and therefore ``free'' in terms of complexity.
-For instance, when looking for a cause for
+For instance, when looking for a cause of
 an anomaly in the living-room, other anomalies occurring in the same
@@
 
-Formally, we now consider only predicates where knowledge of the effect event $e$  is appended to all programs $k$:
+Formally, we now consider that knowledge of the effect is given. This consists, for instance, of appending a description of effect $e$  is appended to all programs $k$:
 $\pi_{k::e}(c)$, where $::$ is the append operation.
@@
 \section{Results}
-
+\label{sec:results}
 \subsection{Setup}
@@
     \begin{tabular}{l|c|c}
-        Event Id & Description & Relative memorability (bits)\\
+        Event Id & Description                   & Relative memorability (bits) \\
         \hline
@@
 This shows the memorability score's performance as a classifying tool for ``out-of-the norm'' events.
-In this example, memorability alone correctly identified 18 manually generated events with only 2 false-positives.
+In this example, memorability alone correctly identified 18 manually generated events with only 5 false-positives.
 While the direct application of memorability for classification or event detection is not within the scope of this paper (see Section \ref{sec:related} for complexity-based detection), this first result is on par with the motivation of memorability being in accordance with the intuitive notion.
@@
 temperature axis and naming the difference with the prototypical day.
+
+% ? Maybe a small part on the importance of the introduction of the random pick predicate ?
+% The stark difference observed between the two examples in this work shows the importance of being able to select a random event from the set of
 

### 2021-11-03 new_paper.tex
Fixing following Ada's comments pt2

@@
 \usepackage{algorithm2e}
+\usepackage{caption}
+\usepackage{subcaption}
 
@@
 % Full title of the paper (Capitalized)
-\Title{What should I notice? Evaluating the memorability of events for abductive inference.
+\Title{What should I notice? Using Alogithmic Information Theory to evaluate the memorability of events in smart homes
 }
@@
 
-% Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{ When confronted with an unprecedented situation, humans typically show good
-performance in quickly identifying noticeable past events and proposing them
-as possible causal hypotheses. This kind of abductive inference is widely
-overlooked in modern AI approaches which rely on massive datasets to learn
-associative patterns. Our proposal is to formalize and compute a ``memorability''
-score over a set of events recorded from a cyber-physical system.
-This score can then be used to select relevant information to be
-remembered and to propose causal hypotheses in unusual situations on demand.
-The approach is meant to be complementary to more traditional
-learning-focused abduction techniques.
+\abstract{
+With the increasing number of connected devices, complex systems such as smart homes record a multitude of events of various types, magnitude and characteristics.
+Faced with this variety and number, current systems struggle to identify which events can be considered more memorable than others.
+On the other hand, human beings are able, without knowledge of the system's inner working or large previous datasets, to quickly categorize some events as being more ``memorable'' than others.
+Having this ability would allow the system to identify and summarize a situation to the user or use the most memorable events as possible hypotheses in an abducitve inference process.
+Our proposal is to use Algorithmic Information Theory to formally define a ``memorability'' score based on the concept of retrieving events by using predicative filters.
+After providing a theoretical framework, we provide smart-home examples to illustrate how our approach could be implemented in practive and qualitatively compare our ``memorability'' score with human perception.
+}
+
+% When confronted with an unprecedented situation, humans typically show good
+% performance in quickly identifying noticeable past events and proposing them
+% as possible causal hypotheses. This kind of abductive inference is widely
+% overlooked in modern AI approaches which rely on massive datasets to learn
+% associative patterns. Our proposal is to formalize and compute a ``memorability''
+% score over a set of events recorded from a cyber-physical system.
+% This score can then be used to select relevant information to be
+% remembered and to propose causal hypotheses in unusual situations on demand.
+% The approach is meant to be complementary to more traditional
+% learning-focused abduction techniques.
 We provide a theoretical framework based on Algorithmic Information Theory to define the notion of memorability of events.
@@
 
-% Keywords
 \keyword{Kolmogorov Complexity; Algorithmic Information Theory; Simplicity; Abduction; Surprise}
@@
 \section{Introduction}
+\label{sec:introduction}
 
@@
 automated using knowledge bases and statistical methods,the third approach, which can be used without any knowledge of the occurring phenomenon or past occurrences.
-Instead, it relies on the observation that some events appear more ``memorable'' than others and, as such, can be considered as possible hypotheses if need be.
+Instead, it relies on the observation that some events appear more ``memorable'' than others and, as such, can be considered as possible hypotheses if need be.\cite{magnani_abduction_2011}
 
@@
 % ? What to do with the "intuition" thing? Keep it? Leave it?
-To address this issue, we started from the following intuition: while all
+To address this issue, we started from the following supposition: while all
 events, regardless of their characteristics or nature, can be uniquely
@@
 likely to require less words to be described.
-``Last year's hottest day'' is simpler than ``the 182\textsuperscript{nd} day 7 years ago''.
+This supposition appears to be in line with some observations: for instance, a corrolation has been found between word frequency and length\cite{strauss_word_2007}.
+Thus, ``last year's hottest day'' is simpler than ``the 182\textsuperscript{nd} day 7 years ago''.
 How can we quantify this relative simplicity?
@@
 
-For machines to implement and compute description complexity, we need a formal framework and computation methods that are in line with human intuition.
+For machines to implement and compute description complexity, we need a formal framework and computation methods that incorporate the aforementionned process.
 Algorithmic Information Theory (AIT) appears to be such a framework, as it is consistent with the human perception of complexity\cite{li_introduction_2008,dessalles2011coincidences,delahaye_numerical_2012}.
@@
 (\cdot, 1\_year)$ is \emph{true} of events that occurred last year. Selecting
-%%%%%%%%%%%%%%%%%%%%%% ! MOVED FIGURE HERE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[hb]
@@
 \end{figure}
-%%%%%%%%%%%%%%%%%%%%%%! FIGURE END %%%%%%%%%%%%%%%%%%%%%%%%%%%%
 all events, from the memory, that satisfy a given predicate corresponds to
@@
 % * Paragraph re-written 
+The rest of this article is organized as follows.
 We will first briefly introduce
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Materials and Methods}
-
-\subsection{Theoretical Background}
+\section{Theoretical Framework}
+\label{sec:framework}
+
+\subsection{Background}
 \label{sec:theory}
@@
 As \cite{dessalles2011coincidences} suggests a link between the unexpectedness and cognitive relevance, we propose to define the memorability of an event in a similar way.
+Since we want to use this score in applications, we need a definition that is well-defined and computable in practice.
 Hence, we define the memorability $M(e)$ of an event as the absolute difference between its description complexity $K_d(e)$ and its expected description complexity $K_{exp}(e)$:
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
+\section{Experiments}
 \label{sec:results}
@@
 candidates. Our choice was also motivated by
-the existence of previous work involving smart home simulations capable of quickly generating data from which we could extract events and test our methods.
+the existence of previous work\cite{lalanda_self-aware_2017} involving smart home simulations capable of quickly generating data from which we could extract events and test our methods.
 
@@
 \subsubsection{Scenario 2}
-We consider an experimental smart home setup, with various sensors, which we simulate over a period of time
+We consider an experimental smart home setup, with various sensors, which we simulate over a period of time.
 To carry out the simulation, we built custom modules into the existing iCasa
@@
 a single user, and an outdoor zone. All four rooms are equipped
-with a temperature controller system that monitors and controls heaters (fig.
+with a temperature control system that controlling heaters (see Figure.
 \ref{fig:view}).
@@
     \centering
-    \includegraphics[width=.8\linewidth]{figures/simulator}
+    \includegraphics[width=.8\linewidth]{figures/icasa_interface.png}
     \caption{View of the simulator's web interface provided by iCasa. The four
@@
     \includegraphics[width=\linewidth]{figures/ts_example}
-    \caption{Time series data from the simulation: outdoor temperature (blue) and
-        controller temperature of a room (orange). To be used in our framework, these time series data are
+    \caption{Time series data from the simulation: outdoor temperature (red) and
+        controller temperature of a room (blue). To be used in our framework, these time series data are
         processed by a simple threshold-based event detection.}
@@
 
-The values observed from all devices and zones was sampled throughout
-the simulation run. The resulting data (figure
+Values observed from all devices and zones were sampled throughout
+the simulation. The resulting data (figure
 \ref{fig:ts_example}) was then used as a basis for our experiments.
 We then process the time series data to identify and characterize events.
-Since the selection of events is not the focus point of our present work (see sec.
+Since how events are detected is not the focus point of our present work (see Section.
 \ref{sec:related}), we perform event detection merely based on threshold and precomputed conditions.
@@
 \begin{itemize}
-    \item $\mathtt{label(e, k)}$: whether the event $e$ has the label ranked
-          $k^{th}$ in the label list.
-    \item $\mathtt{rank(e, r, a)}$: whether the event $e$ has the rank $r$
-          along axis $a$.
+    \item $\mathtt{label(e, k)}$: whether the event $e$ has the $k^{th}$ most frequent label (meaning that frequent labels are simpler to express than rare ones)
+    \item $\mathtt{rank(e, r, a)}$: whether the event $e$ is ranked $r^{th}$ for characteristic $a$, where characteristics are encoded by their frequency (again, common characteristics are the simplest ones)
     \item $\mathtt{day(e, k)}$: whether the event $e$ occurred $k$ days ago.
@@
 
-Similar predicates were used for scenario 1, with the notable exception of the addition of a \texttt{RandomPick} predicate, which is true if the id of the event matches a random program seed $k$.
+% Similar predicates were used for scenario 1.
+However, to show the impact of allowing random selection of events, we added a \texttt{RandomPick} predicate, which is true if the id of the event matches a random program seed $k$.
+As it could be argued that in some cases, randomly selecting an event is not possible, we decided to implement this predicate in only one scenario.
+
+% with the notable exception of the addition of a \texttt{RandomPick} predicate, which is true if the id of the event matches a random program seed $k$.
 Using this predicate and the associated filter in only one of the two illustrative examples shows the importance of allowing random selection in the complexity computation.
@@
 
-The description length $L_{\pi, k}$ of using a predicate is computed as follows: since the set of predicates is finite and known, $L(\pi) = \log2(|\Pi|)$ are enough to describe the predicate concept $\pi$\footnote{This approach gives an equal complexity to all predicate concepts.
+The description length $L(\pi, k)$ of a predicate $\pi_k$ is computed as follows: since the set of predicates is finite and known, $L(\pi) = \log2(|\Pi|)$ are enough to describe the predicate concept $\pi$\footnote{This approach gives an equal complexity to all predicate concepts.
 While this could be argued when using many concepts, as humans do, we deemed better, for our small examples, to consider all concepts as equal.}.
-To encode the argument $k$ of the predicate, we used the prefix-free Elias delta code\cite{elias_universal_1975}, which requires $L(k) = \log2(k) + 2 \log2(\log2(k)+1) + 1]$ bits.
+To encode the argument $k$ of the predicate, we used the widely used prefix-free Elias delta code\cite{elias_universal_1975}, which requires $L(k) = \log2(k) + 2 \log2(\log2(k)+1) + 1]$ bits.
 The total cost of describing $\pi_{k}$ therefore is
@@
 \begin{equation}
-    \L(\pi, k) = \log2(|\Pi|) + \log2(k) + 2 \log2(\log2(k) + 1) + 1
+    \label{eq:pred_cost}
+    L(\pi, k) = \log2(|\Pi|) + \log2(k) + 2 \log2(\log2(k) + 1) + 1
 \end{equation}
@@
 
+Our code is written in Python.
+Examples are presented in the form of Jupyter Notebooks, which allow to quickly reproduce our results and figures.
+All code is available on our Github: \url{https://github.com/EtienneHouze/memorability_code}.
+
 \subsection{Results}
@@
 The computation of the description complexity measure for the 2400 events recorded in this scenario took around 90 seconds using an i7-8600u laptop.
-The resulting complexities, in Figure \ref{fig:scenar1_complexity}, show that TV events appear more simple than luminosity events.
+The resulting complexities, in Figure \ref{fig:scenar1_complexity}, show that TV events appear simpler than luminosity events.
 In fact, for most luminosity events, the complexity corresponds to the random choice complexity, 25 bits in this example.
-This pattern comes from the number of luminosity events per day, 24, is too high to consider the retrieval path $day,  randomPick$ simpler than $randomPick$ from the entire memory, except for the most recent days, which appear simpler (on the right-hand side of Figure \ref{fig:scenar1_complexity})
+This is due to the frequency of luminosity events: given that there are 24 of them per day, considering the retrieval path $day(k), randomPick(24)$ costs more bits than $randomPick(N)$, where $N$ is the total number of events, with the notable exception of recent days, for which the value $k$ is low (see the right-hand side of Figure \ref{fig:scenar1_complexity}).
+
+% TODO add some annotations to the figure?
 \begin{figure}
     \centering
-    \includegraphics[width=.7\linewidth]{figures/memorability_scenar_1.png}
+    \includegraphics[width=\linewidth]{figures/memorability_scenar_1.png}
     \caption{Description complexity for events recorded in the TV scenario. Luminosity events are shown in blue, TV events in green}
@@
 
+% !
+Response to Ada's comment: Indeed, the jump in memorability does not seem that high.
+However, we may consider that the smart TV event will remain memorable through time, as it will remain the first use of the device 'smart TV', whereas the two other events will become less memorable as time passes by...
+Maybe add some paragraph in the discussions section to talk about the phenomenon of considering recent events memorable?
 \begin{table}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Discussion}
-
-\subsection{Related Works}
+\subsection{Discussion: the subjectivity of predicates}
+For individuals, the notion of memorability and complexity of events is highly subjective: the same event may appear usual for a person, while being exceptional for her neighbor.
+Our approach to memorbaility embraces this subjectivity, incorporating it through the definiton of predicates.
+
+In Figure \ref{fig:3preds}, we show this subjectivity by comparing the analyses of the same base memory of events, generated using the same setup as for Example 1, using different three different sets of predicates.
+The resulting figures show different phenomena based on the predicates at hand by the module.
+In figure \ref{fig:pred_1}, only time and labels were recognized by predicates.
+As a result, the general trend of the curve is logarithmic, as events from $T$ days ago require $O(\log(T)))$ bits of information.
+Two main sequences of light events appear: as some light events were recorded simultaneously as TV events, they require the additional information of their label to be retreived.
+In Figure \ref{fig:pred_2}, we added predicates qualifying the light intensity, along with a day/night distinction.
+This added capacity isolated some light events as much simpler: they appear as a green horizontal line.
+These events are times when the light was low during day, or high during night.
+As such, they are outliers, and therefore require less inforamtion to be retrieved.
+
+Finally, Figure \ref{fig:pred_3} comes from a module which has the ability to retrieve any event from a memory of size $N$, at the expense of $O(\log(N))$ bits of information.
+This added capacity has the immediate effect of setting a clear upper limit to the description complexity of items, since any item can be retrieved using $\l(N)$ bits (in this example, this limit is around 25 bits).
+While this kind of direct retrieval is trivial in computer science (one could use the memory address of any stored event), its correlation with human cognition is not obvious: can humans be considered to have this possibility of selecting any event from their entire memory, without restriction regarding their nature, their age, their magnitude?
+However, this highlights an interesting phenomenon: using this direct retrieval, the module makes no distinction, complexity-wise, between events past a given threshold.
+In other words, there is a generic "uninteresting" category of events, in which the module makes no distinction.
+
+In addition to the number of available predicate concepts, one might also tweak their complexity.
+In our example scenarios, we used only a handful of predicates, hence we chose to assign all predicate concepts the same bit description length, $\log(|\Pi|)$ (see Equation.
+\ref{eq:pred_cost}).
+When using more predicate concepts, we may instead give different costs to some of them, to take into account the complexity difference between them: for instance, the generic time concept $\texttt{year}$ would likely require less bits than a predicate $\texttt{tun}$ (an old Mayan time unit, corresponding to roughly 360 days).
+Similar to the selection of predicates to be used, the complexity assigned to each predicate concept, that is, their encoding, denotes the subjectivity of the user.
+\newpage
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}{\linewidth}
+        \centering
+        \includegraphics[width=.7\linewidth]{figures/preds_1.png}
+        \caption{}
+        \label{fig:pred_1}
+    \end{subfigure}
+    \begin{subfigure}{\linewidth}
+        \centering
+        \includegraphics[width=.7\linewidth]{figures/preds_2.png}
+        \caption{}
+        \label{fig:pred_2}
+    \end{subfigure}
+    \begin{subfigure}{\linewidth}
+        \centering
+        \includegraphics[width=.7\linewidth]{figures/preds_3.png}
+        \caption{}
+        \label{fig:pred_3}
+    \end{subfigure}
+    \caption{The same memory of events, analyzed through three different sets of predicates.
+Light events are in blue, TV events in salmon.
+Figure.
+\ref{fig:pred_1} uses only time-related predicates (days and hours ago), while Figure.
+\ref{fig:pred_2} adds label predicates alongside with "dark" and day/night predicates.
+Figure \ref{fig:pred_3} adds the possibility to select directly select any event from the memory of size $N$, at the cost of $l(N)$ bits.}
+    \label{fig:3preds}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Related Works}
 \label{sec:related}
@@
 
-\subsection{Perspectives}
+\section{Perspectives}
 \label{sec:future}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusions}
-The introduction of description complexity and the subsequent memorability measure allow to compare events of various types with different characteristics.
+\section{Conclusion}
+% TODO Re-write the conclusion!
+The ability to identify some events as memorable is useful in the current context of computing, where connected devices record many events of different characteristics, magnitudes, types.
+This is particularly true in complex cyber-physical systems such as smart homes.
+We propose an approach to evaluate the memorability as a difference between the expected description complexity of an event and its actual value.
+With this definition, something is memorable if it requires much less or much more information to be described.
+To formalize this notion, we used the minimum Minimum Description Length principle exposed in Algorithmic Information Theory.
+By defining filter operations from predicates and successively applying these filters, we define formal ways of describing events.
+The length of the shortest of such way can be used as a measure of description complexity.
+
+We provided an implementation algorithm to compute this measure on events and showed some of its results in two smart home examples.
+These application scenarios qualitatively show how our measure fares in comparision with the human notion of memorability, and how this measure can be used to propose relevant hypotheses in an abductive inference process without having further knowledge of the system.
+
+Finally, we discussed the inherent subjectivity of our approach, embodied in the choice of available predicates and their encoding.
+We link our approach to existing works which use Algorithmic Information Theory principles to analyze time series data or detect outlier events.
+
+% The introduction of description complexity and the subsequent memorability measure allow to compare events of various types with different characteristics.
 As such, it provides a formal and generic framework to discriminate peculiar events and outliers.
 
-The computation being based on predicates, the memorability score is entirely dependent on their definitions.
+% The computation being based on predicates, the memorability score is entirely dependent on their definitions.
 This can be seen as a caveat or a feature of our approach: it makes the measure subjective while enabling the consideration of user's preferences in the definition and choice of predicates.

### 2022-01-12 new_paper.tex
fixes after comments pt.3

@@
 The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, soul, multirow, microtype, tikz, totcount, changepage, paracol, attrib, upgreek, cleveref, amsthm, hyphenat, natbib, hyperref, footmisc, url, geometry, newfloat, caption
-\usepackage{algorithm2e}
+\usepackage[linesnumbered, ruled]{algorithm2e}
 \usepackage{caption}
@@
 With the increasing number of connected devices, complex systems such as smart homes record a multitude of events of various types, magnitude and characteristics.
-Faced with this variety and number, current systems struggle to identify which events can be considered more memorable than others.
-On the other hand, human beings are able, without knowledge of the system's inner working or large previous datasets, to quickly categorize some events as being more ``memorable'' than others.
-Having this ability would allow the system to identify and summarize a situation to the user or use the most memorable events as possible hypotheses in an abducitve inference process.
-Our proposal is to use Algorithmic Information Theory to formally define a ``memorability'' score based on the concept of retrieving events by using predicative filters.
-After providing a theoretical framework, we provide smart-home examples to illustrate how our approach could be implemented in practive and qualitatively compare our ``memorability'' score with human perception.
+Current systems struggle to identify which events can be considered more memorable than others.
+In contrast, human are able to quickly categorize some events as being more ``memorable'' than others.
+They do so wihtout relying on knowledge of the system's inner working or large previous datasets.
+Having this ability would allow the system to: i) identify and summarize a situation to the user; ii) suggest the most memorable events as possible hypotheses in an abductive inference process.
+Our proposal is to use Algorithmic Information Theory to define a ``memorability'' score by retrieving events using predicative filters.
+We use smart-home examples to illustrate how our theoretical approach can be implemented in practice.
 }
 
-% When confronted with an unprecedented situation, humans typically show good
-% performance in quickly identifying noticeable past events and proposing them
-% as possible causal hypotheses. This kind of abductive inference is widely
-% overlooked in modern AI approaches which rely on massive datasets to learn
-% associative patterns. Our proposal is to formalize and compute a ``memorability''
-% score over a set of events recorded from a cyber-physical system.
-% This score can then be used to select relevant information to be
-% remembered and to propose causal hypotheses in unusual situations on demand.
-% The approach is meant to be complementary to more traditional
-% learning-focused abduction techniques.
-We provide a theoretical framework based on Algorithmic Information Theory to define the notion of memorability of events.
-Then, we implement our approaches in two examples based on smart-home scenario to validate our approach in toy cases and illustrate how it could be applied.}
-
-\keyword{Kolmogorov Complexity; Algorithmic Information Theory; Simplicity; Abduction; Surprise}
+\keyword{Kolmogorov Complexity; Algorithmic Information Theory; Simplicity; Abduction; Memorability}
 
@@
 
-As a user has just switch on the TV in her all-equipped living-room, the
+As a user has just switched on the TV for the first time in her new  all-equipped living-room, the
 lights dim and the window blinds go down. Intrigued by this behavior, she
@@
 By performing \emph{abductive inference} \cite{magnani_abduction_2011}. This mental operation
-is a key element of humans' ability to understand the world: from the
-observed consequences, they infer the possible causes.
-% TODO: answer in the next paragraph?
-
-In this example, there are mainly three possibilities to come to the
+is a key element of the human ability to understand the world: from the
+observation, they infer the possible causes.
+
+In this example, there are mainly three ways through which the user could come to the
 conclusion. (1) If the user knows how the smart living-room system works, if she
@@
 knowledge to perform abduction. (2) If she has no knowledge
-about the system but made several observation of the same behavior, she may examine past
+about the system but made several observations of the same behavior, she may examine past
 correlations and figure out that turning on the TV set often leads the blinds to
@@
 suspect that the TV is a possible cause for the observed event, just because it appears to her as a memorable recent event (as it is its first occurrence). This example
-suggests that human beings are able to use distinct methods to perform abductive tasks
+suggests that human beings are able to use at least three distinct methods to perform abductive tasks
 and infer new knowledge, depending on the situation. While the first two mechanisms can be
-automated using knowledge bases and statistical methods,the third approach, which can be used without any knowledge of the occurring phenomenon or past occurrences.
-Instead, it relies on the observation that some events appear more ``memorable'' than others and, as such, can be considered as possible hypotheses if need be.\cite{magnani_abduction_2011}
-
-Defining a score of memorability is not straightforward. First, events can be of different nature, and not directly
-comparable. Even for restricted systems such as smart homes, noticeable events
+automated using knowledge bases and statistical methods, the third approach, which can be used without any previous knowledge of the occurring phenomenon or of its past occurrences, remains, to the best of our knowledge, not implemented in current systems.
+Doing so would require the system to have a way of distinguishing some events as more ``memorable'' than others and then consider them as possible hypotheses if need be\cite{magnani_abduction_2011}.
+
+Defining a memorability score is not straightforward. First, events can be of different nature, and not directly
+comparable. For systems such as smart homes, noticeable events
 range from device removal to presence detection or unusually high
@@
 ``memorability''. 
-% ! Addiction of the follwoing sentence, to show what we can use memorability for.
 In addition to the aforementionned use for abductive inference, having access to a computation of memorability would allow a system to summarize a situation to its user by presenting only the most memorable events: for instance, a summary of notable events that occurred during the home owner's absence.
 
-% ? What to do with the "intuition" thing? Keep it? Leave it?
 To address this issue, we started from the following supposition: while all
@@
 likely to require less words to be described.
-This supposition appears to be in line with some observations: for instance, a corrolation has been found between word frequency and length\cite{strauss_word_2007}.
-Thus, ``last year's hottest day'' is simpler than ``the 182\textsuperscript{nd} day 7 years ago''.
+This supposition appears to be in line with observations of human cognition: for instance, a correlation has been found between word frequency and length\cite{strauss_word_2007}, the shortest words being the most common; humans seem to be sensitive to the complexity of events when assessing a coincidence\cite{griffiths_probability_2003,dessalles_coincidences_2011}.
+To illustrate this, think, for instance, of the ``182$^{nd}$ day of 7 years ago'' compared to ``the hottest day ever recorded'': the latter seems more memorable than the former.
 How can we quantify this relative simplicity?
-We could evaluate the complexity of each
+We propose to evaluate the complexity of each
 description, taking into account both the complexity of the concept words (a date
-of occurrence, a temperature ranking), and of the arguments (1st hottest,
+of occurrence, a temperature ranking), and of the arguments (``hottest'' vs
 182 and 7). The resulting values define the  \emph{description complexity}
-of events. Our intuition is that memorable events require simpler and less numerous qualifiers to be
-unambiguously described than unremarkable ones.
-% ? Response to Ada's remark: added a thing in the 1st paragraph to show what we want ot do with the memorable events.
-
-For machines to implement and compute description complexity, we need a formal framework and computation methods that incorporate the aforementionned process.
-Algorithmic Information Theory (AIT) appears to be such a framework, as it is consistent with the human perception of complexity\cite{li_introduction_2008,dessalles2011coincidences,delahaye_numerical_2012}.
-Our method is as follows: we
-consider events as being elements stored in what we call \emph{base memory}. To
-reproduce the language features applicable to events, we use \emph
-{predicates}, i.e. functions assigning a boolean value to events. For instance, the predicate $\mathtt{date}
-(\cdot, 1\_year)$ is \emph{true} of events that occurred last year. Selecting
-\begin{figure}[hb]
+of events. Following our supposition, we would define memorable events as requiring simpler and less numerous qualifiers to be unambiguously described than unremarkable ones.
+
+\begin{figure}[!ht]
     \centering
@@
 \end{figure}
-all events, from the memory, that satisfy a given predicate corresponds to
+
+For machines to implement and compute description complexity, we need a formal framework and computation methods that incorporate the aforementionned process.
+Algorithmic Information Theory (AIT) appears to be such a framework, as it is consistent with the human perception of complexity\cite{li_introduction_2008,dessalles_coincidences_2011,delahaye_numerical_2012}.
+Our method is as follows: we
+consider events as being elements stored in what we call \emph{base memory}. To
+reproduce the language features applicable to events, we use \emph
+{predicates}, i.e. functions assigning a boolean value to events. For instance, the predicate $\mathtt{date}
+(\cdot, 1\_year)$ is \emph{true} of events that occurred last year. Selecting
+all events from the memory that satisfy a given predicate corresponds to
 a \emph{filter} operation. It generates another memory that is a subset of the previous one.
@@
 
-% * Paragraph re-written 
 The rest of this article is organized as follows.
 We will first briefly introduce
-some basic notions of Algorithmic Information Theory in subsection~\ref{sec:theory}.
-We then expose our contribution by presenting in subsection~\ref{sec:computing} our formal definition of memorability and related notions.
-Then, we present an implementation example of these definitions with two smart-home examples in section~\ref{sec:results}.
+some relevant notions of Algorithmic Information Theory in subsection~\ref{sec:theory}.
+We then expose our contribution in subsections~\ref{sec:computing} with a formal definition of memorability.
+We present an implementation of these definitions with two smart-home examples in section~\ref{sec:results}.
 The results of these experiments are then presented and discussed.
-Finally we explore other related works in subsection \ref{sec:related} and explore possible extensions of our work in subsection \ref{sec:future}.
-
-% Then, in
-% subsection~\ref{sec:computing}, we show how we compute the
-% memorability of past events. We illustrate our approach in section~\ref{sec:example} by describing an implementation
-% in a smart home simulation context. In section~\ref{sec:related} we review
-% a few studies that use complexity theory to generate explanations for the smart home
-% and we discuss how our work can be linked with them. We explore potential
-% extensions of our work in section \ref{sec:future}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+We explore other related works in subsection \ref{sec:related} and explore possible extensions of our work in subsection \ref{sec:future}.
+
 \section{Theoretical Framework}
@@
 \end{equation}
-
 where the additional constant $C_{U}$ does not depend on $s$.
@@
 simplicity or complexity of events.
+Hence, the term $K_w(e)$, which designates the causal complexity, models the cost of information a hypothetical World Machine would require to produce the observed outcome.
+This can be, for instance, the cost of different parameters in a physical model.
+As such, this quantity is highly dependent on the knowledge the human subject has of its surrounding environment.
 
@@
 trivial: the complexity of describing this person will be equivalent to
-distinguishing it from the global population, which is also roughly
+distinguishing her from the global population, which is also roughly
 equivalent to the (causal) complexity of describing the circumstances having brought this person to be in
@@
 lower, the description complexity $K(e)$ drops while the causal complexity
-$K_W (e)$ remains unchanged. This is why this latter  event appears unexpected.
-
-As \cite{dessalles2011coincidences} suggests a link between the unexpectedness and cognitive relevance, we propose to define the memorability of an event in a similar way.
+$K_W (e)$ remains unchanged.
+This is why this latter  event appears unexpected.
+By contrast, if you knew beforehand that your best friend used to walk in this park, the causal complexity $K_w(e)$ would be significantly lower, hence reducing the surprise.
+
+As \cite{dessalles2011coincidences} suggests a link between unexpectedness and cognitive relevance, we propose to define the memorability of an event in a similar way.
 Since we want to use this score in applications, we need a definition that is well-defined and computable in practice.
-Hence, we define the memorability $M(e)$ of an event as the absolute difference between its description complexity $K_d(e)$ and its expected description complexity $K_{exp}(e)$:
+We define the memorability $M(e)$ of an event as the absolute difference between its description complexity $K_d(e)$ and its expected description complexity $K_{exp}(e)$:
 
@@
 
-% ?
-Peut-être ajouter un paragraphe sur pourquoi ne pas directement utiliser la notion de unexpectedness.
-1) pas calculable.
-2) pas forcément bien définie dans un cas pratique d'application (comment définit la complexité causale ???)
 Contrary to the definition of unexpectedness from Equation \ref{eq:unexpected}, we use an absolute value: we do so to acknowledge the fact that events more complex than expected can be memorable as well\footnote{In the original paper \cite{dessalles2011coincidences}, exceptionally complex events are described by considering complexity itself as a way to describe the event: see ``the Pisa Tower effect''\cite{dessalles_pisa_nodate}}.
@@
 
-\subsection{Computing the memorability of events}
+\subsection{Defining and retrieving events}
 \label{sec:computing}
-\subsubsection{Retrieving an event}
 
@@
 
-\subsubsection{Description complexity of events}
+\subsection{Description complexity of events}
 
@@
 $p(M_0) = e$, this path provides a possible unambiguous description for $e$.
-We therefore define the description complexity of $e$ as the minimum complexity of a path $p$ retrieving $e$ from the base memory $M_0$.
+We define the description complexity of $e$ as the minimum complexity of a path $p$ retrieving $e$ from the base memory $M_0$.
 
@@
 \end{equation}
-
 where the bit-length $L(p)$ of a retrieval path is defined as the number of bits of a string encoding the path.
@@
 \end{align}
-where $L(\pi_\square)$ and $L(k_\square)$ respectively denote the length, in bits, required to express the predicate's concept and program.
+where $L(\pi_i)$ and $L(k_i)$ denote the length, in bits, required to express the predicate's concept and program, respectively.
 This length may vary depending of the encoding choice, see Section \ref{sec:results} for an example.
@@
 memories which will be explored during the next iteration(lines 21--23). Each
-pass thus explore retrieval paths of increasing length. When a singleton memory
+pass thus explores retrieval paths of increasing length. When a singleton memory
 is reached, the complexity of its unique element is upper-bounded with the
@@
 
-\subsubsection{Computing Memorability}
+\subsection{Computing Memorability}
 As stated in Equation \ref{eq:memorability}, we define memorability $M(e)$ as the absolute difference between the description complexity of an event and its expected value.
@@
 neighborhoods $N_{\pi, k}(e)$, with the convention that $N_{\pi, k}(e) = \emptyset$
-if $e$ does not satisfy $\pi_k$, we can compute an average expected complexity for $e$, by summing the complexity of events in all neighbourhoods of $e$ and dividing the total by the number of events in the neighborhoods:
+if $e$ does not satisfy $\pi_k$, we can compute an average expected complexity for $e$, by summing the complexity of events in all neighborhoods of $e$ and dividing the total by the number of events in the neighborhoods:
 
@@
 
-\subsubsection{From memorability to abduction}
+This metrics answers to the different problems exposed in the introduction: by using a universal measure for complexity, bits, it allows to compare values from different dimensions.
+For instance, it solves the dilemma of recent events: is a big event a long time ago more memorable than a smaller one that occurred only a few minutes ago?
+With our approach to complexity, each one of these dimensions will scale logarithmically.
+The balance between them depends on the subjectivity of the system, which is encoded in the intrinsic complexity of predicates for magnitude and dimension.
+
+\subsection{Defining relative memorability for abduction}
 % TODO : see ada's comment for this part
@@
 
-The knowledge attached to the occurrence of $e$ can be integrated into the description complexity definition by using conditional complexity $K_d(c | e)$: The information contained in $e$ is considered as given, and therefore ``free'' in terms of complexity.
+The knowledge attached to the occurrence of $e$ can be integrated into the description complexity definition by using conditional complexity $K_d(c | e)$: The information contained in $e$ is considered as given, and therefore as ``free'' in terms of complexity.
 For instance, when looking for a cause of
@@
 
-Formally, we now consider that knowledge of the effect is given. This consists, for instance, of appending a description of effect $e$  is appended to all programs $k$:
-$\pi_{k::e}(c)$, where $::$ is the append operation.
+Formally, we now consider that knowledge of the effect is given. This consists, for instance, of appending a description of effect $e$  to all programs $k$:
+$\pi_{e::k}(c)$, where $::$ is the \emph{append} operation.
 The set of paths obtained with such predicates is noted $P^\infty_e$.
-This append operation is free in terms of bit-length in the computation of complexity, since the effect event $e$ is an input of the problem.
+This \emph{append} operation is free in terms of bit-length in the computation of complexity, since the effect event $e$ is an input of the problem.
 Therefore, we have $L'(\pi_{k::e}) = L(\pi_k) = L(\pi) + L(k)$.
@@
 It can then be averaged over similar events to compute the expected conditional description complexity, $K_{exp}(c|e)$.
-From this, we come to the definition of the conditional memorability, which measures how memorable an event $c$ turns out to be in the context of the occurrence of another event $e$:
+From this, we come to the definition of the \emph{conditional memorability}, which measures how memorable an event $c$ turns out to be in the context of the occurrence of another event $e$:
 
@@
 presented as the motivation of this paper: when confronted with a surprising
-situation, and in the absence of any other source of information, events that appear more memorable than others, with regards to the target event will be selected as potential causes.
+situation, and in the absence of any other source of information, events that appear more memorable than others with regards to the target event will be selected as potential causes.
 As such, our conditional memorability score provides a ranking that can be used for abductive inference.
-
-This metrics answers the different problems exposed in the introduction: by using a universal measure for complexity, bits, it allows to compare values from different dimensions.
-For instance, it solves the dilemma of recent events: is a big event a long time ago more memorable than a smaller one only a few minutes ago?
-The answer from conditional memorability is that "it depends".
-It depends on the complexity cost set for the predicates describing time and the magnitude of the event.
-These complexity costs should reflect the user's perception of the different dimensions (see Section \ref{sec:future}).
 
@@
 \subsection{Setup}
+\label{sec:setup}
 
@@
 
-\subsubsection{Scenario 1}
-In this setup, we aim to reproduce the example mentioned in Section \ref{sec:introduction}: the installation of a brand new smart TV causes unpredictable effects on the light of a room.
+% \subsubsection{Scenario 0}
+% First, we study a simple situation where the frequency of an event, and not its magnitude, makes it remarkable.
+Consider a delivery truck servicing a store.
+One apparition of this truck in the street is registered as an event.
+Over a period of 200 days, on most days, the truck appears at most once per day, to deliver the store.
+However, on one particular day, the truck appears over a dozen times in the street!
+The sudden high frequency of appearance makes observations of the truck on that particular day more memorable, while remaining individually the same.
+
+\subsubsection{The ``TV'' scenario}
+In this setup, we aim to reproduce the example mentioned in Section \ref{sec:introduction}: the installation and first use of a brand new smart TV causes unpredictable effects on the light of a room.
 To play this situation, we created a set of events covering a period of 100 days, corresponding to the past knowledge of the house.
-Two kinds of events are recorded: ``TV event'', corresponding to TV use, and luminosity events, describing the luminosity of the room at a given time.
-Low lights occur at night, and can occasionally occur during daytime, with a small probability.
-On the 100\textsuperscript{th} day, a ``TV event'' is recorded with a different ``device'' characteristic.
-Shortly after, the light lowers, which is recorded in the following ``light'' event.
-
-\subsubsection{Scenario 2}
-We consider an experimental smart home setup, with various sensors, which we simulate over a period of time.
-To carry out the simulation, we built custom modules into the existing iCasa
-smart home simulator\cite{lalanda_self-aware_2017}. ICasa
+Two kinds of events are recorded: ``TV event'', corresponding to TV use (old and new); and luminosity events, describing the luminosity of the room at a given time.
+Low lights occur at night, and can occasionally occur during daytime (for instance if the blinds are down).
+On the 100\textsuperscript{th} day, a ``TV event'' is recorded with a different ``device'' characteristic: it corresponds to the first use of the new smart TV.
+Shortly after, the light dims, which is recorded in a ``light'' event.
+
+\subsubsection{The ``temperature'' scenario}
+We consider an experimental smart home setup with various sensors, which we simulate over a period of time.
+To carry out the simulation, we used the iCasa smart home simulator platform\cite{lalanda_self-aware_2017} 
+to which we added custom modules. iCasa
 offers a simulation of autonomic systems that can handle internal communications,
@@
 a single user, and an outdoor zone. All four rooms are equipped
-with a temperature control system that controlling heaters (see Figure.
+with a temperature control system in charge of heaters (see Figure.
 \ref{fig:view}).
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
     \centering
@@
 serving as a background for our experiments. To produce outstanding
-events, we randomly generated about twenty events, spanning over the whole
+events, we randomly generated about twenty special events, spanning over the whole
 duration of the simulation, of different kinds:
@@
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
     \includegraphics[width=\linewidth]{figures/ts_example}
@@
 We then process the time series data to identify and characterize events.
-Since how events are detected is not the focus point of our present work (see Section.
-\ref{sec:related}), we perform event detection merely based on threshold and precomputed conditions.
+Since the ways events are detected is not the focus point of our present work (see Section.
+\ref{sec:related}), we perform event detection merely based on threshold comparison.
 
@@
 
-The description length $L(\pi, k)$ of a predicate $\pi_k$ is computed as follows: since the set of predicates is finite and known, $L(\pi) = \log2(|\Pi|)$ are enough to describe the predicate concept $\pi$\footnote{This approach gives an equal complexity to all predicate concepts.
-While this could be argued when using many concepts, as humans do, we deemed better, for our small examples, to consider all concepts as equal.}.
-To encode the argument $k$ of the predicate, we used the widely used prefix-free Elias delta code\cite{elias_universal_1975}, which requires $L(k) = \log2(k) + 2 \log2(\log2(k)+1) + 1]$ bits.
+The description length $L(\pi, k)$ of a predicate $\pi_k$ is computed as follows: since the set of predicates is finite and known, $L(\pi) = \log_2(|\Pi|)$ bits are enough to describe the predicate concept $\pi$\footnote{This approach gives an equal complexity to all predicate concepts.
+Though this choice may be questionable when using many concepts, as humans do, we we used this simplification as our examples rely on few predicates.}.
+To encode the argument $k$ of the predicate, we used the widely used prefix-free Elias delta code\cite{elias_universal_1975}, which requires $L(k) = \log_2(k) + 2 \log_2(\log_2(k)+1) + 1]$ bits.
 The total cost of describing $\pi_{k}$ therefore is
@@
     \label{eq:pred_cost}
-    L(\pi, k) = \log2(|\Pi|) + \log2(k) + 2 \log2(\log2(k) + 1) + 1
-\end{equation}
-
-With a straightforward implementation of memory, predicates and filters, we could run alg. \ref{alg:complex_iter} worked, though, as expected, it took too
+    L(\pi, k) = \log_2(|\Pi|) + \log_2(k) + 2 \log_2(\log_2(k) + 1) + 1
+\end{equation}
+
+With a straightforward implementation of memory, predicates and filters, we could run Algorithm \ref{alg:complex_iter}. However, it took too
 long to be usable in realistic scenarios with hundreds or thousands of events to
-consider. In order to facilitate and speed up computations, we also implemented
+consider. In order to facilitate and speed up computations, we implemented
 the following improvements:
@@
     \item The memory object was augmented with various built-in rankings, allowing
-          for faster operations during future filtering. For instance, since the memory
+          for faster operations during filtering. For instance, since the memory
           object keeps a mapping from timestamp to events one can perform a quick
-          filtering by date without having to loop over all stored element. This convenient mapping,
+          filtering by date without having to loop over all stored elements. This convenient mapping,
           however, is not directly used to retrieve events by their date of occurrence, so as to
@@
 All code is available on our Github: \url{https://github.com/EtienneHouze/memorability_code}.
+Figures from the code are interactive: hovering the mouse above points displays the iD of the event, as well as the predicates used in the optimal retrieval path.
 
@@
 
-\subsubsection{The ``TV'' scenario:}
+% \subsubsection{The ``Truck'' scenario (scenario 0)}
+% For this scenario, a random sampling of events resulted in 57 seeings of the delivery truck in the street, 21 of them occurring in the same ``remarkable'' day.
+The results of the complexity computations of our methods are displayed in Figure \ref{fig:freq_complex}.
+They show that events from the ``remarkable'' day are more complex than other appearances of the delivery truck.
+This happens because the amount of information required to describe an event from that particular day is higher: it requires additional temporal information, such as the hour and minutes of appearance.
+As such, our method shows its potential in defining memorability not only in terms of magnitude of events, but also in terms of their frequency of apparition.
+
+% \begin{figure}[!ht]
+%     \centering
+%     \includegraphics[width=.9\linewidth]{figures/complexity_frequency.png}
+%     \caption{Complexity of the truck events in Scenario 0. Observations from the ``remarkable'' day, in red, stand out as more complex than the rest.}
+%     \label{fig:freq_complex}
+% \end{figure}
+
+\subsubsection{The ``TV'' scenario}
 \paragraph{Memorability}
-The computation of the description complexity measure for the 2400 events recorded in this scenario took around 90 seconds using an i7-8600u laptop.
-The resulting complexities, in Figure \ref{fig:scenar1_complexity}, show that TV events appear simpler than luminosity events.
-In fact, for most luminosity events, the complexity corresponds to the random choice complexity, 25 bits in this example.
-This is due to the frequency of luminosity events: given that there are 24 of them per day, considering the retrieval path $day(k), randomPick(24)$ costs more bits than $randomPick(N)$, where $N$ is the total number of events, with the notable exception of recent days, for which the value $k$ is low (see the right-hand side of Figure \ref{fig:scenar1_complexity}).
-
-% TODO add some annotations to the figure?
-\begin{figure}
+
+The computation of the description complexity measure for the 2500 recorded events took around 90 seconds using an i7-8565u-equipped laptop.
+The resulting memorability scores are shown in Figure \ref{fig:scenar1_complexity}.
+We can observe that, on average, recent events are given a higher score: this reflects the cost of designating an event by the time elapsed since its occurrence.
+Furthermore, we can see that some light events, in blue, are more memorable than the main sequence.
+These events correspond to either events that occurred simultaneously to TV events, in salmon: as they are simultaneous to another event, they require additional information to be singled out, temporal information not being enough.
+Thus, they appear as ``surprisingly'' more complex than the rest of their king, hence more memorable.
+
+Some light events also appear more memorable than the rest: they are events when light was surprisingly low given the hour and therefore are easier to retrieve.
+While these general observations are consistent with an a-priori intuition, the results are dependent of the choice of predicates used for the computation.
+See section \ref{sec:discussion} for a discussion on this dependence.
+
+\begin{figure}[!ht]
     \centering
-    \includegraphics[width=\linewidth]{figures/memorability_scenar_1.png}
-    \caption{Description complexity for events recorded in the TV scenario. Luminosity events are shown in blue, TV events in green}
+    \includegraphics[width=.9\linewidth]{figures/memo_scenar_1.png}
+    \caption{Computed memorability for events recorded in the TV scenario. Luminosity events are shown in blue, TV events in salmon.}
     \label{fig:scenar1_complexity}
@@
 \paragraph{Abduction}
-To show the potential application to abductive inference, we use our approach to the first scenario, to see if memorability alone can find the brand new TV to be a reasonable cause for the sudden low light.
-
-The result of our algorithm is given in Table \ref{tab:abduction_res}.
-It shows that the system correctly identifies the TV as being the cause.
-The reason for this choice is that, since the smart TV's device ID is unique among all other events of type ``TV'', its description complexity is very low.
-
-% \begin{figure}
-%     \centering
-%     \includegraphics[width=.8\linewidth]{./figures/abduction_1_results}
-%     \caption{Outputs of the memorability-based abduction module. This represent the 3 events in the memory having the highest memorability score relative to the consequence event.}
-%     \label{fig:abduction_res}
-% \end{figure}
-
-% !
-Response to Ada's comment: Indeed, the jump in memorability does not seem that high.
-However, we may consider that the smart TV event will remain memorable through time, as it will remain the first use of the device 'smart TV', whereas the two other events will become less memorable as time passes by...
-Maybe add some paragraph in the discussions section to talk about the phenomenon of considering recent events memorable?
+To show the potential application to abductive inference, we use our approach to the first scenario, to see if memorability alone can find the brand new TV to be a reasonable cause for the sudden light dimming.
+
+The result of our algorithm is given in Table \ref{tab:abduction_res}: the system correctly identifies the new TV as being the cause.
+The reason for this choice is that, since the smart TV's device ID is unique among all other events of type ``TV'', its description complexity stands out as being significantly lower than the others, and therefore entails a high memorability score.
+
 \begin{table}
@@
 
-\subsubsection{The ``temperature'' scenario:}
+\subsubsection{The ``temperature'' scenario}
 % \paragraph{Memorability}
 
-The results of the description complexity evaluation, for the described setup,
-are shown in Figure \ref{fig:computed_cplx}. The entire computation, over 4
-iterations (meaning that retrieval paths contained at most 4 filters), took
-around 30 seconds on a commercial laptop with an i7-8700u CPU.
-
-The general trend appears different from the ``TV example'' presented in Figure \ref{fig:scenar1_complexity} : instead of having a capped complexity score, a logarithmic trend appears.
-This phenomenon is due to the absence, in this example, of the \texttt{randomPick} predicate; instead, usual events, with outstanding particularity, can be retrieved by using the \texttt{day} predicate and a \texttt{label} predicate.
-As such, it appears that most days are considered usual by our memorability score, creating the main logarithmic sequence of blue dots: there is no better way to retrieve them than simply giving the day of their occurrence.
+For this scenario, 578 events were recorded from the setup described in Subsection \ref{sec:setup}.
+THe computation of the memorability and complexity scores took around 30 seconds on a commercial laptop equipped with an i7-8565u CPU.
+Four loops of Algorithm \ref{alg:complex_iter} where required (as additional loops did not improve scores).
+
+Similar to the previous scenario, the general trend of events appears as a time-dependent logarithmic score for most events: this corresponds to events for which the simplest retrieval path consists of describing the elapsed duration since their occurrence using the \texttt{day} predicate.
+As such, it appears that most days are considered ``usual'' by our memorability score.
+This effect produces the main logarithmic sequence of blue dots.
 On the other hand, some events stand out in terms of
@@
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
     \centering
     \includegraphics[width=.9\linewidth]{figures/complexity_scenar_2.png}
-    \caption{Computed description complexities of events with retrieval paths of
+    \caption{Description complexities of events with retrieval paths of
         length at most 4. Events of type ``day'' (blue), ``hot'' (red), ``cold''
-        (gray), ``device removal''(orange) and ``device addition'' (cyan) are shown. Events mentioned in the text are specified.}
+        (gray), ``device removal''(orange) and ``device addition'' (cyan) are shown.}
     \label{fig:computed_cplx}
@@
 Table \ref{tab:paths} illustrates this by exhibiting the retrieval paths used for complexity computation for these events.
-
-\begin{figure}[ht]
-    \centering
-    \includegraphics[width=.8\linewidth]{figures/memo_scenar_2.png}
-    \caption{Memorability score for events in memory}
-    \label{fig:result1}
-\end{figure}
 
@@
 
+\begin{figure}[!ht]
+    \centering
+    \includegraphics[width=.9\linewidth]{figures/memo_scenar_2.png}
+    \caption{Memorability score for events in memory. Events mentioned in the text are highlighted.}
+    \label{fig:result1}
+\end{figure}
+
 Given that we generated the data used for this experiment, it is possible to
-flag all perturbation events from the usual daily events and evaluate how a
+flag all perturbation events apart from the usual daily events and evaluate how a
 detection based on ``memorability'' score would perform in distinguishing these
 events.
-The result is presented as a ROC curve in Figure \ref{fig:roc}.
+The result is presented as a ROC curve in Figure \ref{fig:roc}, obtained by varying the memorability threshold.
 This shows the memorability score's performance as a classifying tool for ``out-of-the norm'' events.
@@
 
-\begin{figure}[ht]
+\begin{figure}[!ht]
     \centering
@@
     \caption{Experimental ROC curve (True Positive Rate against False Positive
-        Rate) for a classifier based on our memorability score. Measures consider 23
+        Rate) for a classifier which compares the memorability of events to a given threshold, which we vary to change the sensitivity of our detector. Measures consider 23
         manually flagged events as memorable (events added to the normal background
@@
 
-% \paragraph{Abduction}
-% As an illustration of the abductive inference made possible our memorability
-% score, we used as a target anomaly the temperature drops visible in the raw
-% time series data from our scenario (fig \ref{fig:ts_example}).
-
-% \textbf{TODO: the figure is still missing!}
-
 \subsection{Discussion: the subjectivity of predicates}
+\label{sec:discussion}
 For individuals, the notion of memorability and complexity of events is highly subjective: the same event may appear usual for a person, while being exceptional for her neighbor.
-Our approach to memorbaility embraces this subjectivity, incorporating it through the definiton of predicates.
-
-In Figure \ref{fig:3preds}, we show this subjectivity by comparing the analyses of the same base memory of events, generated using the same setup as for Example 1, using different three different sets of predicates.
-The resulting figures show different phenomena based on the predicates at hand by the module.
-In figure \ref{fig:pred_1}, only time and labels were recognized by predicates.
-As a result, the general trend of the curve is logarithmic, as events from $T$ days ago require $O(\log(T)))$ bits of information.
-Two main sequences of light events appear: as some light events were recorded simultaneously as TV events, they require the additional information of their label to be retreived.
-In Figure \ref{fig:pred_2}, we added predicates qualifying the light intensity, along with a day/night distinction.
-This added capacity isolated some light events as much simpler: they appear as a green horizontal line.
-These events are times when the light was low during day, or high during night.
-As such, they are outliers, and therefore require less inforamtion to be retrieved.
-
-Finally, Figure \ref{fig:pred_3} comes from a module which has the ability to retrieve any event from a memory of size $N$, at the expense of $O(\log(N))$ bits of information.
-This added capacity has the immediate effect of setting a clear upper limit to the description complexity of items, since any item can be retrieved using $\l(N)$ bits (in this example, this limit is around 25 bits).
-While this kind of direct retrieval is trivial in computer science (one could use the memory address of any stored event), its correlation with human cognition is not obvious: can humans be considered to have this possibility of selecting any event from their entire memory, without restriction regarding their nature, their age, their magnitude?
-However, this highlights an interesting phenomenon: using this direct retrieval, the module makes no distinction, complexity-wise, between events past a given threshold.
-In other words, there is a generic "uninteresting" category of events, in which the module makes no distinction.
-
-In addition to the number of available predicate concepts, one might also tweak their complexity.
-In our example scenarios, we used only a handful of predicates, hence we chose to assign all predicate concepts the same bit description length, $\log(|\Pi|)$ (see Equation.
-\ref{eq:pred_cost}).
-When using more predicate concepts, we may instead give different costs to some of them, to take into account the complexity difference between them: for instance, the generic time concept $\texttt{year}$ would likely require less bits than a predicate $\texttt{tun}$ (an old Mayan time unit, corresponding to roughly 360 days).
-Similar to the selection of predicates to be used, the complexity assigned to each predicate concept, that is, their encoding, denotes the subjectivity of the user.
-\newpage
-
-\begin{figure}[H]
+Our approach to memorability embraces this subjectivity, incorporating it through the definition of predicates.
+
+\begin{figure}[!ht]
     \centering
@@
 
+In Figure \ref{fig:3preds}, we show this subjectivity by comparing the analyses of the same base memory of events, generated using the same setup as for Example 1, using three different sets of predicates.
+The resulting figures show different phenomena based on the predicates available to the module.
+In figure \ref{fig:pred_1}, only time and labels were captured by predicates.
+As a result, the general trend of the curve is logarithmic, as events from $T$ days ago require $O(L(T)))$ bits of information.
+Two main sequences of light events appear: as some light events were recorded simultaneously as TV events, they require the additional information of their label to be retrieved.
+In Figure \ref{fig:pred_2}, we added predicates qualifying the light intensity, along with a day/night distinction.
+This added capacity isolated some light events as much simpler: they all appear as aligned green points.
+These events are times when the light was low during day, or high during night.
+As such, they are outliers, and therefore require less information to be retrieved.
+
+Finally, Figure \ref{fig:pred_3} comes from a module which has the ability to retrieve any event from a memory of size $N$, at the expense of $O(L(N))$ bits of information.
+This added capacity has the immediate effect of setting a clear upper limit to the description complexity of items, since any item can be retrieved using $\L(N)$ bits (in this example, this limit is around 25 bits).
+While this kind of direct retrieval is trivial in computer science (one could use the memory address of any stored event), its correlation with human cognition is not obvious: can humans be considered to have this possibility of selecting any event from their entire memory, without restriction regarding their nature, their time of occurrence, their magnitude?
+However, this highlights an interesting phenomenon: using this direct retrieval, the module makes no distinction, complexity-wise, between events past a given threshold.
+In other words, there is a generic "uninteresting" category of events, among which the module makes no distinction.
+
+In addition to the number of available predicate concepts, one might also tweak their complexity.
+In our example scenarios, we used only a handful of predicates, hence we chose to assign all predicate concepts the same bit description length, $\log(|\Pi|)$ (see Equation.
+\ref{eq:pred_cost}).
+When using more predicate concepts, we may instead give different costs to some of them, to take into account the complexity difference between them: for instance, the generic time concept $\texttt{year}$ would likely require less bits than a predicate $\texttt{tun}$ (an old Mayan time unit, corresponding to roughly 360 days).
+Similar to the selection of predicates used, the complexity assigned to each predicate concept, that is, their encoding, denotes the subjectivity of the user.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Our work is intended to be integrated into larger-scale frameworks to monitor
-and detect events in complex environment such as smart homes. In these works, the
-approach to smart homes is often regarded as self-organizing systems \cite{kramer_rigorous_2009,kounev_notion_2017}. As such, they present capacities of adaptation to new goals,
+and detect events in complex environment such as smart homes.
+Smart homes are often regarded as self-organizing systems \cite{kramer_rigorous_2009,kounev_notion_2017}.
+As such, they present capacities of adaptation to new goals,
 new components, new environment. A commonly used approach is the principle of
-autonomic system, which minimizes user's intervention for management of the
+autonomic system, which minimizes users' interventions for the management of the
 system \cite{kounev_notion_2017,kephart_vision_2003}.
 
-We want to inscribe our work among other classical concurrent approaches to
-abduction. In situations where more data is available, we could for instance rely
-on correlation or causal inference from known relations \cite{peters_elements_2017,fadiga_or_2021}.
-Previous relations between inference and complexity have been studied. In fact, the
+In situations where more data is available, we could rely
+on correlation or causal inference from known relations \cite{peters_elements_2017,fadiga_or_2021}. Relations between inference and complexity have already been studied. The
 case of inference was one of the motivations for R. Solomonoff to introduce his
 universal algorithmic probability \cite{solomonoff_formal_1964} as a tool to
-reach an idealized inference machine, creating the notion of complexity concomitantly
+reach an idealized inference machine, creating the notion of complexity simultaneously
 to Kolmogorov. Subsequently, notions of complexity re-emerged in causal inference:
-\cite{janzing_causal_2010} found, when a causal link exists between two random variables,
-considering the direct joint probability is simpler, in terms of Kolmogorov complexity,
-than the inverse direction.
+\cite{janzing_causal_2010} found than, when a causal link exists between two random variables $A \rightarrow B$, the decomposition of the joint probability is simpler in the direct than the inverse direction: $K(P(A)) + K(P(B|A)) < K(B) + K(A|B)$.
 
@@
 More recently, \cite{marx_causal_2018} used Minimum Description Length to determine, given a joint probability distribution over $(X,Y)$, whether $X$ causes $Y$ or $Y$ causes $X$.
-Their method is based on tree models, and implying that a model respecting the causal relation will be simpler to describe.
+Their method is based on tree models and it implies that a model respecting the causal relation will be simpler to describe.
 
@@
 \cite{hu_discovering_2011} proposes a MDL approach to find the intrinsic dimensions of time series.
-All these approaches are interesting, and can be integrated into our canvas as tools to detect events using only complexity.
-Using this, one could achieve a purely complexity-driven process for detecting and qualifying events.
-
-The philosophy of our approach can be closely related to the ``Isolation forests'' method\cite{liu_isolation_2008,hariri_extended_2021}.
+
+The philosophy of our approach can be related to the ``Isolation forests'' method\cite{liu_isolation_2008,hariri_extended_2021}.
 It evaluates the isolation of data points by constructing random binary tree classifiers.
@@
 Using the average height of leaves in the tree as a metrics, this approach succeeds in identifying outlier points without having to define a ``typical'' point.
-This approach can be understood in terms of complexity: each node of binary tree classifier needs a fixed amount of information to be described (which variable and threshold are used).
-So nodes that place higher need less information to be described.
+This approach can be understood in terms of complexity: each node of a binary tree classifier needs a fixed amount of information to be described (it must indicates which variable and threshold are used).
+So nodes that are located higher in the tree need less information to be described.
 As such, outliers need less information to be singled out.
@@
 By using predicates as a proxy for complexity computation, our methods is more general, as it is agnostic regarding the nature of events.
-However, the introduction of predicates adds a subjectivity in the determination of memorable points, as we will discuss later.
 
 While all these works advocate in favor of a strong link between complexity and the
-discovery of causes, not other works extends the notion up to the point we propose in this paper, using complexity only as a tool to express the intuitive notion of memorability, and using it for inference.
+discovery of causes, they do not extend the notion up to the point we propose in this paper, namely using the sole complexity as a tool to express the intuitive notion of memorability, and using it for inference.
 
@@
 
-The practical application of the theoretical notions of memorability and relative memorability requires future developments.
-In this section, we have selected two of them which seem to us, to date, the most challenging.
-
-First, the main limitation of the current approach is the requirement of
+The practical application of the theoretical notions of memorability and relative memorability requires further developments. We highlight two of them which seem to us, to date, the most challenging.
+
+First, one limitation of the current approach is the requirement of
 predefined predicate concepts, from which the different filters are constructed.
-As an extension, we suggest that in the future, we could explore online
-generation of such predicate. A possibility would be to analyze discriminating
-dimensions of incoming data and create predicate as to name these differences,
+As an extension, we suggest the possibility of exploring such predicates dynamically.
+One mya analyze discriminating
+dimensions of incoming data and create predicates to name these differences,
 similar to the contrast operations proposed in \cite{dessalles_conceptual_2015,
@@
 
-% ? Maybe a small part on the importance of the introduction of the random pick predicate ?
-% The stark difference observed between the two examples in this work shows the importance of being able to select a random event from the set of
-
-While the execution time is not part of the theoretical view of complexity, it
+Second, execution time is not part of the theoretical view of complexity, it
 is of prime importance for practical applications, especially when one considers
 implementation into real-time systems or embedded devices. While the computation
-we propose appear to be heavy, and possibly heavier as the number of allowed
-predicates grows, significant time savings can be achieve by trimming the base
-memory of past events deemed the most ``non memorable''. For instance, one can
-only retains the 100 most memorable events from the past. The difficulty with
-this approach is that such operations should be done in a manner to not
-interfere with the complexity computations for new elements: by forgetting some
+we propose appears to be heavy, and possibly heavier as the number of allowed
+predicates grows, significant time savings can be achieved by trimming the base
+memory of past events deemed the most ``non memorable''. For instance, one could retain
+the 100 most memorable events from the past. The difficulty with
+this approach is that such operations should be done without interfering with the complexity computations of new elements: by forgetting some
 past events, even uninteresting ones, one should make sure to keep track of what
@@
 the way towards practical implementations and dynamic selection of interesting
-events and help reducing the memory and computation cost of data-driven applications.
+events and help reduce the memory and computation cost of data-driven applications.
 
@@
 \section{Conclusion}
-% TODO Re-write the conclusion!
-The ability to identify some events as memorable is useful in the current context of computing, where connected devices record many events of different characteristics, magnitudes, types.
+
+We proposed an approach to evaluate memorability as a difference between the expected description complexity of an event and its actual value.
+With our definition, something is memorable if it requires less information or more information than expected to be described.
+To formalize this notion, we used principles of minimal description exposed in Algorithmic Information Theory.
+By defining filter operations from predicates and successively applying these filters, we defined formal ways of describing events, whose length can then be measured to evaluate their description length.
+From this, memorability is defined as the absolute difference between the average complexity of similar events and the actual description complexity of the event.
+
+% The ability to identify some events as memorable is useful in the current context of computing, where connected devices record many events of different characteristics, magnitudes and types.
 This is particularly true in complex cyber-physical systems such as smart homes.
-We propose an approach to evaluate the memorability as a difference between the expected description complexity of an event and its actual value.
+We proposed an approach to evaluate the memorability as a difference between the expected description complexity of an event and its actual value.
 With this definition, something is memorable if it requires much less or much more information to be described.
-To formalize this notion, we used the minimum Minimum Description Length principle exposed in Algorithmic Information Theory.
-By defining filter operations from predicates and successively applying these filters, we define formal ways of describing events.
-The length of the shortest of such way can be used as a measure of description complexity.
+To formalize this notion, we used the Minimum Description Length principle exposed in Algorithmic Information Theory.
+By defining filter operations from predicates and successively applying these filters, we defined formal ways of describing events.
+The length of the shortest of these descriptions can be used as a measure of description complexity.
+From this, we defined memorability as the absolute difference between the expected and the actual description complexity.
 
 We provided an implementation algorithm to compute this measure on events and showed some of its results in two smart home examples.
-These application scenarios qualitatively show how our measure fares in comparision with the human notion of memorability, and how this measure can be used to propose relevant hypotheses in an abductive inference process without having further knowledge of the system.
-
-Finally, we discussed the inherent subjectivity of our approach, embodied in the choice of available predicates and their encoding.
-We link our approach to existing works which use Algorithmic Information Theory principles to analyze time series data or detect outlier events.
-
-% The introduction of description complexity and the subsequent memorability measure allow to compare events of various types with different characteristics.
-As such, it provides a formal and generic framework to discriminate peculiar events and outliers.
-
-% The computation being based on predicates, the memorability score is entirely dependent on their definitions.
-This can be seen as a caveat or a feature of our approach: it makes the measure subjective while enabling the consideration of user's preferences in the definition and choice of predicates.
-This consideration can be the topic of future research and open up potential applications.
+These application scenarios qualitatively show how our measure fares in comparison with the human notion of memorability, and how this measure can be used to propose relevant hypotheses in an abductive inference process without having further knowledge of the system.
+We discussed the inherent subjectivity of our approach by exhibiting the impact of the choice of predicates for complexity computation in a toy example.
+We consider extending our approach by including online learning of predicates that would make our approach coincide with the subjectivity of the system's user.
+
+The ability to identify some events as memorable is useful in the current context of computing where connected devices record many events with heterogeneous characteristics, magnitude and types.
+In this context, our approach of memorability aims to bring a unifying measure to sort out some events as ``memorable''.
+This ability can pave the way towards memorability-based abduction or selective memory.
 
@@
 
-\funding{\textbf{TODO je ne sais pas quoi mettre ici ? Indiquer le financement de la thèse ?}}
+\funding{This work is part of a PhD thesis on Explainable AI for Smart Homes conducted at EDF R\&D and Télécom Paris.
+It is financed by EDF R\&D and the Agence Nationale pour la Recherche Technologique (ANRT).}
 
@@
 
-% \begin{specialtable}[H]
+% \begin{specialtable}[!h]
 %     \small

### 2022-01-26 new_paper.tex
including ada's comments

@@
 They do so wihtout relying on knowledge of the system's inner working or large previous datasets.
-Having this ability would allow the system to: i) identify and summarize a situation to the user; ii) suggest the most memorable events as possible hypotheses in an abductive inference process.
+Having this ability would allow the system to: i) identify and summarize a situation to the user by presenting only memorable events; ii) suggest the most memorable events as possible hypotheses in an abductive inference process.
 Our proposal is to use Algorithmic Information Theory to define a ``memorability'' score by retrieving events using predicative filters.
@@
 
-As a user has just switched on the TV for the first time in her new  all-equipped living-room, the
+Let us consider the following scenario. As a user has just switched on the TV for the first time in her new  all-equipped living-room, the
 lights dim and the window blinds go down. Intrigued by this behavior, she
 quickly infers that both light dimming and blind closing occurred as a
-consequence of the TV being turned on. How did she come to this conclusion?
-By performing \emph{abductive inference} \cite{magnani_abduction_2011}. This mental operation
+consequence of the TV being switched on. How did she come to this conclusion?
+By performing \emph{abductive inference}~\cite{magnani_abduction_2011}. This mental operation
 is a key element of the human ability to understand the world: from the
@@
 correlations and figure out that turning on the TV set often leads the blinds to
-close and the lights to dim. (3) If there are no previous occurrences of the event (e.g. it is the first time she turns on the TV
+close and the lights to dim. (3) If there are no previous occurrences of the event (e.g. it is the first time she switches on the TV
 in the living-room), she may still be able to
@@
 suggests that human beings are able to use at least three distinct methods to perform abductive tasks
-and infer new knowledge, depending on the situation. While the first two mechanisms can be
-automated using knowledge bases and statistical methods, the third approach, which can be used without any previous knowledge of the occurring phenomenon or of its past occurrences, remains, to the best of our knowledge, not implemented in current systems.
-Doing so would require the system to have a way of distinguishing some events as more ``memorable'' than others and then consider them as possible hypotheses if need be\cite{magnani_abduction_2011}.
+and infer new knowledge, depending on the situation.
+While the first two mechanisms are commonly used in Explainable AI literature~\cite{arrieta_explainable_2020} to identify causes and explanations of agents' decisions, they require preliminary knowledge or data.
+On the other hand,  the third approach can be used without any previous knowledge of the occurring phenomenon or of its past occurrences but remains, to the best of our knowledge, not implemented in current systems.
+Doing so would require the system to have a way of distinguishing some events as more ``memorable'' than others and then consider them as possible hypotheses if need be~\cite{magnani_abduction_2011}.
 
@@
 likely to require less words to be described.
-This supposition appears to be in line with observations of human cognition: for instance, a correlation has been found between word frequency and length\cite{strauss_word_2007}, the shortest words being the most common; humans seem to be sensitive to the complexity of events when assessing a coincidence\cite{griffiths_probability_2003,dessalles_coincidences_2011}.
+This supposition appears to be in line with observations of human cognition: for instance, a correlation has been found between word frequency and length~\cite{strauss_word_2007}, the shortest words being the most common; also, humans seem to be sensitive to the complexity of events when assessing a coincidence~\cite{griffiths_probability_2003,dessalles_coincidences_2011}.
 To illustrate this, think, for instance, of the ``182$^{nd}$ day of 7 years ago'' compared to ``the hottest day ever recorded'': the latter seems more memorable than the former.
@@
 
-\begin{figure}[!ht]
-    \centering
-    \includegraphics[width=.95\linewidth]{figures/filters}
-    \caption{Retrieving an event through successive predicative
-        filters. From the base memory (yellow), successive filters select events satisfying the associated predicate (gray arrows). For example,
-        filter $f_1$ selects events from last year, i.e. which satisfy the
-        predicate $\mathtt{date}(event, 1\_year)$. In this case, successively
-        applying filters $f_1$, $f_2$ and $f_3$ yields a unique event $e_h$, last
-        year's hottest day. The complexity of this event can then be upper-bounded
-        by the complexity of the three filters as they give an unambiguous way
-        to describe the event within the memory.}
-    \label{fig:filters}
-\end{figure}
-
 For machines to implement and compute description complexity, we need a formal framework and computation methods that incorporate the aforementionned process.
-Algorithmic Information Theory (AIT) appears to be such a framework, as it is consistent with the human perception of complexity\cite{li_introduction_2008,dessalles_coincidences_2011,delahaye_numerical_2012}.
+Algorithmic Information Theory (AIT) appears to be such a framework, as it is consistent with the human perception of complexity~\cite{li_introduction_2008,dessalles_coincidences_2011,delahaye_numerical_2012}.
 Our method is as follows: we
@@
 
+\begin{figure}[!ht]
+    \centering
+    \includegraphics[width=.95\linewidth]{figures/filters}
+    \caption{Retrieving an event through successive predicative
+        filters. From the memory (yellow), successive filters select events satisfying the associated predicate (gray arrows). For example,
+        filter $f_1$ selects events from last year, i.e. which satisfy the
+        predicate $\mathtt{date}(event, 1\_year)$. In this case, successively
+        applying filters $f_1$, $f_2$ and $f_3$ yields a unique event $e_h$, last
+        year's hottest day. The complexity of this event can then be upper-bounded
+        by the sum of the complexity of the three filters as they give an unambiguous way
+        to describe the event within the memory.}
+    \label{fig:filters}
+\end{figure}
+
 The rest of this article is organized as follows.
-We will first briefly introduce
+First, we briefly introduce
 some relevant notions of Algorithmic Information Theory in subsection~\ref{sec:theory}.
@@
 any object represented by a finite binary
-string)\cite{kolmogorov_three_1965,li_introduction_2008}. The complexity $K(s)$ of a (finite) binary string is the length in bits $L(p)$ of the shortest program $p$
+string)\cite{kolmogorov_three_1965,li_introduction_2008}. The complexity $K(s)$ of a (finite) binary string $s$ is the length in bits $L(p)$ of the shortest program $p$
 which, if given as input to a universal Turing Machine $U$, outputs $s$.
@@
     \label{eq:kolmo_def}
-    K_{U}(s) = \min_{p}\left\{L(p)|U(p)=s\right\}
+    K_{U}(s) = \min_{p}\left\{L(p)|U(p)=s\right\}\,.
 \end{equation}
@@
     \label{eq:inequality_univ}
-    K_{U'}(s) \le l(p_{u}) + K_{U}(s)
+    K_{U'}(s) \le l(p_{u}) + K_{U}(s)\,.
 \end{equation}
@@
 \begin{equation}
-    \forall U\in\text{TM}, \forall s, |K(s) - K_{u}(s)| \le C_{U}
+    \forall U\in\text{TM}, \forall s, |K(s) - K_{u}(s)| \le C_{U}\,,
 \end{equation}
@@
 Note that the notion of Kolmogorov complexity involves no requirement on the execution time of
-programs, only their length in bits matters for the computation of
+programs; only their length in bits matters for the computation of
 complexity. Though Kolmogorov complexity can be shown to be
-incomputable\cite{li_introduction_2008},
+incomputable~\cite{li_introduction_2008},
 it can be approximated with
@@
 notion and perception of complexity from a human standpoint. For instance, the
-complexity of short binary strings evaluated in \cite{delahaye_numerical_2012}
+complexity of short binary strings evaluated in~\cite{delahaye_numerical_2012}
 shows similar results to human perception of complex strings and patterns. More
-recently, \cite{murena_solving_2020} used Kolmogorov complexity to solve
+recently,~\cite{murena_solving_2020} used Kolmogorov complexity to solve
 analogies and showed results close to human expectations.
@@
 complexity can be pushed farther thanks to the notions of simplicity and unexpectedness,
-which are sometimes regarded of uttermost importance in cognitive science\cite
+which are sometimes considered to be of uttermost importance in cognitive science~\cite
 {chater_simplicity_2003}.
-\cite{dessalles2011coincidences} proposes a formal definition of the
+~\cite{dessalles_coincidences_2011} proposes a formal definition of the
 unexpectedness $U(e)$ of an event, as the difference between an a-priori
@@
 \begin{equation}
-    \label{eq:unexpected} \textit{Unex}(e) = K_{w}(e) - K(e)
+    \label{eq:unexpected} \textit{Unex}(e) = K_{w}(e) - K(e)\,.
 \end{equation}
@@
 simplicity or complexity of events.
-Hence, the term $K_w(e)$, which designates the causal complexity, models the cost of information a hypothetical World Machine would require to produce the observed outcome.
+Hence, the term $K_w(e)$, which designates the causal complexity, models the cost of information that a hypothetical World Machine -- a Turing Machine modeling the person's understanding of the world -- would require to produce the observed outcome.
 This can be, for instance, the cost of different parameters in a physical model.
-As such, this quantity is highly dependent on the knowledge the human subject has of its surrounding environment.
+As such, this quantity is highly dependent on the knowledge that the human subject has of its surrounding environment.
 
@@
 lower, the description complexity $K(e)$ drops while the causal complexity
-$K_W (e)$ remains unchanged.
+$K_w(e)$ remains unchanged.
 This is why this latter  event appears unexpected.
@@
 
-As \cite{dessalles2011coincidences} suggests a link between unexpectedness and cognitive relevance, we propose to define the memorability of an event in a similar way.
+As~\cite{dessalles_coincidences_2011} suggests a link between unexpectedness and cognitive relevance, we propose to define the memorability of an event in a similar way.
+However, the proposed definition of unexpectedness, by introducing the hypothetical World Machine, results in an uncomputable metrics.
 Since we want to use this score in applications, we need a definition that is well-defined and computable in practice.
-We define the memorability $M(e)$ of an event as the absolute difference between its description complexity $K_d(e)$ and its expected description complexity $K_{exp}(e)$:
+We therefore introduce the memorability $M(e)$ of an event as the absolute difference between its description complexity $K_d(e)$ and its expected description complexity $K_{exp}(e)$:
 
@@
     \label{eq:memorability}
-    M(e) = |K_{exp}(e) - K_d(e)|
-\end{equation}
-
-Contrary to the definition of unexpectedness from Equation \ref{eq:unexpected}, we use an absolute value: we do so to acknowledge the fact that events more complex than expected can be memorable as well\footnote{In the original paper \cite{dessalles2011coincidences}, exceptionally complex events are described by considering complexity itself as a way to describe the event: see ``the Pisa Tower effect''\cite{dessalles_pisa_nodate}}.
+    M(e) = |K_{exp}(e) - K_d(e)|\,.
+\end{equation}
+
+Contrary to the definition of unexpectedness from Equation \ref{eq:unexpected}, we use an absolute value: we do so to acknowledge the fact that events more complex than expected can be memorable as well\footnote{In the original paper~\cite{dessalles2011coincidences}, exceptionally complex events are described by considering complexity itself as a way to describe the event: see ``the Pisa Tower effect''\cite{dessalles_pisa_nodate}}.
 In the next section, we define computational approximations for the description complexity $K_d$ and the expected complexity $K_{exp}$ of events.
@@
     \label{eq:event}
-    e = (l, t,\mathcal{D})
+    e = (l, t,\mathcal{D})\,,
 \end{equation}
@@
         \mathbf{E}\times \{0,1\}^{*} & \mapsto \{0,1\}    \\
-        (e, k)                        & \mapsto \pi_{k}(e)
+        (e, k)                        & \mapsto \pi_{k}(e)\,.
     \end{cases}
@@
 the event $e$ occurred $1$ year ago.
+Another example would be the predicate $\pi = \mathtt{date}$, which can take as argument $k=\mathtt{year}::\mathtt{month}::\mathtt{day}$, and $\mathtt{date}(e, y, m, d)$ is true if and only if event $e$ occured at the specified date.
 
@@
         \mathcal{M} & \mapsto \mathcal{M}           \\
-        M           & \mapsto \{e \in M | \pi_k(e) \}
+        M           & \mapsto \{e \in M | \pi_k(e) \}\,.
     \end{cases}
@@
     \label{eq:ret_def}
-    p = (f_{\pi_{1}, k_{1}}, \dots, f_{\pi_{n}, k_{n}})
+    p = (f_{\pi_{1}, k_{1}}, \dots, f_{\pi_{n}, k_{n}})\,.
 \end{equation}
@@
     \label{eq:k_desc}
-    K_d(e) = \min_{p \in P_\infty} \{L(p) | p(M_0) = e\}
+    K_d(e) = \min_{p \in P_\infty} \{L(p) | p(M_0) = e\}\,,
 \end{equation}
@@
     L(p) & = L((f_{\pi_1,k_1}, \dots, f_{\pi_n, k_n}))     \\
-         & = L(\pi_1) + L(k_1) + \dots + L(\pi_n) + L(k_n)
+         & = L(\pi_1) + L(k_1) + \dots + L(\pi_n) + L(k_n)\,,
 \end{align}
@@
     \label{eq:approx_k_desc}
-    K_d(e) \leq \min_{p \in P \land p(M_0) = e} L(p) = \min_{p \in P \land p(M_0)=e} \sum_{f_{\pi, k} \in p} L(\pi) + L(k)
-\end{equation}
+    K_d(e) \leq \min_{p \in P \land p(M_0) = e} L(p) = \min_{p \in P \land p(M_0)=e} \sum_{f_{\pi, k} \in p} L(\pi) + L(k)\,.
+\end{equation}
+
+The approximation of description complexity from Equation \ref{eq:approx_k_desc} allows for
+a direct implementation, which is shown in Algorithm \ref{alg:complex_iter}. This
+algorithm operates iteratively: starting from the base memory $M_0$
+(line 1), we apply all possible predicate concepts $\pi$ from a given finite set
+$\Pi$ and programs $k$ (lines 6-7), up to a given length $\mathtt
+{max\_len}$ bits, and apply them: $M' = f_{\pi, k}(M)$ (line 12). We then store
+the pairs $(M', \mathtt{len(}\pi, k\mathtt{)})$ in an array $\mathtt{future_
+{explore}}$. At the end of the iteration, the results of the filters become the
+memories which will be explored during the next iteration(lines 21--23). Each
+pass thus explores retrieval paths of increasing length. When a singleton memory
+is reached, the complexity of its unique element is upper-bounded with the
+length of the corresponding retrieval path (line 14).
 
@@
     \For{$(\mathtt{M_{prev}}, \mathtt{K_{prev}}) \in \mathtt{current_{explore}}$}{
-    \For{$\mathtt{\pi \in \mathcal{P}}$}{
+    \For{$\pi \in \mathcal{P}$}{
     \For{$\mathtt{k \in \{0,1\}^{*}}$}{
-    $\mathtt{K_{current} \leftarrow l(\pi) + l(k) + K_{prev}}$ \;
+    $\mathtt{K_{current}} \leftarrow L(\pi) + L(k) + \mathtt{K_{prev}}$ \;
     \If{$\mathtt{K_{current}} > \mathtt{max_{complex}}$}{
@@
 \end{algorithm}
-
-The approximation of description complexity from Equation \ref{eq:approx_k_desc} allows for
-a direct implementation, which is shown in Algorithm \ref{alg:complex_iter}. This
-algorithm operates iteratively: starting from the base memory $M_0$
-(line 1), we apply all possible predicate concepts $\pi$ from a given finite set
-$\Pi$ and programs $k$ (lines 6-7), up to a given length $\mathtt
-    {max\_len}$ bits, and apply them: $M' = f_{\pi, k}(M)$ (line 12). We then store
-the pairs $(M', \mathtt{len(}\pi, k\mathtt{)})$ in an array $\mathtt{future_
-            {explore}}$. At the end of the iteration, the results of the filters become the
-memories which will be explored during the next iteration(lines 21--23). Each
-pass thus explores retrieval paths of increasing length. When a singleton memory
-is reached, the complexity of its unique element is upper-bounded with the
-length of the corresponding retrieval path (line 14).
 % !
@@
 As stated in Equation \ref{eq:memorability}, we define memorability $M(e)$ as the absolute difference between the description complexity of an event and its expected value.
-As we've just defined $K_d(e)$ and provided an approximation in Equation \ref{eq:approx_k_desc}, we now focus on defining the \emph{expected} description complexity of an event, $K_{exp}(e)$ that appears in Equation \ref{eq:memorability}
+As we've just defined $K_d(e)$ and provided an approximation in Equation \ref{eq:approx_k_desc}, we now focus on defining the \emph{expected} description complexity of an event, $K_{exp}(e)$ that appears in Equation \ref{eq:memorability}.
 
@@
     \sum_{\pi, k} |N_{\pi, k}(e)|
-    }
+    }\,.
 \end{equation}
@@
 
-This metrics answers to the different problems exposed in the introduction: by using a universal measure for complexity, bits, it allows to compare values from different dimensions.
+This metrics solves the different problems exposed in the introduction: by using a universal measure for complexity, bits, it allows to compare values from different dimensions.
 For instance, it solves the dilemma of recent events: is a big event a long time ago more memorable than a smaller one that occurred only a few minutes ago?
-With our approach to complexity, each one of these dimensions will scale logarithmically.
+With our approach to complexity, each one of these dimensions will scale logarithmically, with the bit length of the required predicate parameters.
 The balance between them depends on the subjectivity of the system, which is encoded in the intrinsic complexity of predicates for magnitude and dimension.
@@
 of the memorability score. \emph{Knowing} that we want to find a cause $c$
-for an observed effect $e$, we try to find the most remarkable event in past
+for an observed effect $e$, we try to find the most remarkable event in
 memory that is related to $e$. While our ``memorability'' score identifies remarkable past events, it does not take into account their relatedness to $e$.
@@
     \label{eq:abd_k}
-    K_d(c | e) & = \min_{p \in P^\infty_e} \{L'(p), \quad p(M_0)=c \}                                                \\
-               & = \min_{p \in P^\infty_e} \left\{\sum_{f_{\pi, k::e} \in p} L(\pi) + L(k), \quad p(M_0) = c\right\}
+    K_d(c | e) & = \min_{p \in P^\infty_e} \{L'(p), \quad p(M_0)=c \}\,,                                                \\
+               & = \min_{p \in P^\infty_e} \left\{\sum_{f_{\pi, k::e} \in p} L(\pi) + L(k), \quad p(M_0) = c\right\}\,.
 \end{align}
@@
     \label{eq:cond_mem}
-    M(c|e) = |K_{exp}(c|e) - K_d(c|e)|
+    M(c|e) = |K_{exp}(c|e) - K_d(c|e)|\,.
 \end{equation}
@@
 candidates. Our choice was also motivated by
-the existence of previous work\cite{lalanda_self-aware_2017} involving smart home simulations capable of quickly generating data from which we could extract events and test our methods.
+the existence of previous work~\cite{lalanda_self-aware_2017} involving smart home simulations capable of quickly generating data from which we could extract events and test our methods.
 
@@
 \subsubsection{The ``TV'' scenario}
-In this setup, we aim to reproduce the example mentioned in Section \ref{sec:introduction}: the installation and first use of a brand new smart TV causes unpredictable effects on the light of a room.
+In this setup, we aim to reproduce the example mentioned in Section \ref{sec:introduction}: installing and using for the first time a brand new smart TV causes unpredictable effects on the light of the room where the TV is located.
+Faced with this situation, history-based approaches fail to identify the right hypothesis as there is no previous data for the new TV.
 To play this situation, we created a set of events covering a period of 100 days, corresponding to the past knowledge of the house.
@@
 Low lights occur at night, and can occasionally occur during daytime (for instance if the blinds are down).
-On the 100\textsuperscript{th} day, a ``TV event'' is recorded with a different ``device'' characteristic: it corresponds to the first use of the new smart TV.
+On the 100\textsuperscript{th} day, a ``TV event'' is recorded with a different ``device'' characteristic: it corresponds to the first usage of the new smart TV.
 Shortly after, the light dims, which is recorded in a ``light'' event.
@@
 We consider an experimental smart home setup with various sensors, which we simulate over a period of time.
-To carry out the simulation, we used the iCasa smart home simulator platform\cite{lalanda_self-aware_2017} 
+The smart home simulation data is then processed to identify some predefined events (such as abnormal temperatures).
+To carry out the simulation, we used the iCasa smart home simulator platform~\cite{lalanda_self-aware_2017} 
 to which we added custom modules. iCasa
-offers a simulation of autonomic systems that can handle internal communications,
+allows the simulation of autonomic systems that can handle internal communications,
 the possible insertion of new components at runtime, or the deletion or modification of existing
@@
 comprising a daily cycle of outdoor weather (temperature and sunlight) fluctuations, as well
-as user's movements. All these daily changes create non-noticeable events,
+as user movements. All these daily changes create non-noticeable events,
 serving as a background for our experiments. To produce outstanding
 events, we randomly generated about twenty special events, spanning over the whole
-duration of the simulation, of different kinds:
+duration of the simulation. These outstanding events were of the following kinds:
 
@@
           low temperatures.
-    \item Heater failures: heater may break down, making them turn off regardless
+    \item Heater failures: heaters may break down, making them turn off regardless
           of the command they receive.
@@
 \end{itemize}
+
+Values observed from all devices and zones were sampled throughout
+the simulation. The resulting data (figure
+\ref{fig:ts_example}) was then used as a basis for our experiments.
+We then process the time series data to identify and characterize events.
+Since the ways events are detected is not the focus of our present work (see Section.
+\ref{sec:related}), we perform event detection merely based on threshold comparison, e.g.
+an temperature event is created if temperature measure are above a given threshold for more than a certain amount of time.
 
@@
 
-Values observed from all devices and zones were sampled throughout
-the simulation. The resulting data (figure
-\ref{fig:ts_example}) was then used as a basis for our experiments.
-We then process the time series data to identify and characterize events.
-Since the ways events are detected is not the focus point of our present work (see Section.
-\ref{sec:related}), we perform event detection merely based on threshold comparison.
-
 \subsection{Implementing the complexity computation}
@@
 \begin{itemize}
-    \item $\mathtt{label(e, k)}$: whether the event $e$ has the $k^{th}$ most frequent label (meaning that frequent labels are simpler to express than rare ones)
+    \item $\mathtt{label(e, k)}$: whether the event $e$ has the $k^{th}$ most frequent label (meaning that frequent labels are simpler to express than rare ones)\footnote{In case some labels have the same frequency, an arbitrary rank is used among them.
+However, given the unlikeliness of this occurrence, the impact on complexity is insignificant (this case did not occur in our test examples with a few hundreds events).}.
+In case some labels have the same frequency, a rank is assigned
     \item $\mathtt{rank(e, r, a)}$: whether the event $e$ is ranked $r^{th}$ for characteristic $a$, where characteristics are encoded by their frequency (again, common characteristics are the simplest ones)
@@
 Though this choice may be questionable when using many concepts, as humans do, we we used this simplification as our examples rely on few predicates.}.
-To encode the argument $k$ of the predicate, we used the widely used prefix-free Elias delta code\cite{elias_universal_1975}, which requires $L(k) = \log_2(k) + 2 \log_2(\log_2(k)+1) + 1]$ bits.
+To encode the argument $k$ of the predicate, we used the widely used prefix-free Elias delta code~\cite{elias_universal_1975}, which requires $L(k) = \log_2(k) + 2 \log_2(\log_2(k)+1) + 1]$ bits.
 The total cost of describing $\pi_{k}$ therefore is
@@
     \label{eq:pred_cost}
-    L(\pi, k) = \log_2(|\Pi|) + \log_2(k) + 2 \log_2(\log_2(k) + 1) + 1
+    L(\pi, k) = \log_2(|\Pi|) + \log_2(k) + 2 \log_2(\log_2(k) + 1) + 1\,.
 \end{equation}
@@
 Furthermore, we can see that some light events, in blue, are more memorable than the main sequence.
-These events correspond to either events that occurred simultaneously to TV events, in salmon: as they are simultaneous to another event, they require additional information to be singled out, temporal information not being enough.
+These events correspond to either events that occurred simultaneously to TV events, in pink: as they are simultaneous to another event, they require additional information to be singled out, temporal information not being enough.
 Thus, they appear as ``surprisingly'' more complex than the rest of their king, hence more memorable.
@@
     \includegraphics[width=.9\linewidth]{figures/memo_scenar_1.png}
-    \caption{Computed memorability for events recorded in the TV scenario. Luminosity events are shown in blue, TV events in salmon.}
+    \caption{TV scenario. Computed memorability for events. Luminosity events are shown in blue, TV events in pink. The two circled events are the ones mentioned in the scenario.}
     \label{fig:scenar1_complexity}
@@
 \paragraph{Abduction}
-To show the potential application to abductive inference, we use our approach to the first scenario, to see if memorability alone can find the brand new TV to be a reasonable cause for the sudden light dimming.
+To show the potential application to abductive inference, we use our approach for abduction in the first scenario, to see if memorability alone can find the brand new TV to be a reasonable cause for the sudden light dimming.
 
@@
 The reason for this choice is that, since the smart TV's device ID is unique among all other events of type ``TV'', its description complexity stands out as being significantly lower than the others, and therefore entails a high memorability score.
+While our method does not guarantee the correctness of the hypothesis (in fact, abductive reasoning cannot offer such a guarantee \cite{magnani_abduction_2011}), it provides an alternate hypothesis which corresponds to what a human may have suspected in this case where previous knowledge is unavailable.
 
@@
     \end{tabular}
-    \caption{Output of the memorability-based abduction module: top 3 events for the relative memorability metrics.}
+    \caption{TV Scenario. Output of the memorability-based abduction module: top 3 events for the relative memorability metrics.}
     \label{tab:abduction_res}
@@
 complexity: some appear simpler, as they can be distinguished by using their
-rank along some axis (``the hottest day'', ``the second coldest day''), or the rare occurrence of their kind (``the only fault on the heater'').
+rank along some axis (``the hottest day'', ``the second coldest day''), or the rare occurrence of their kind (``the only recorded failure of the heater'').
 
@@
     \includegraphics[width=.9\linewidth]{figures/complexity_scenar_2.png}
-    \caption{Description complexities of events with retrieval paths of
+    \caption{Temperature scenario. Description complexities of events with retrieval paths of
         length at most 4. Events of type ``day'' (blue), ``hot'' (red), ``cold''
@@
     \end{tabular}
-    \caption{Selected events from the ``temperature example'' with their shortest retrieval path.}
+    \caption{Temperature scenario. Selected events with their shortest retrieval path.}
     \label{tab:paths}
@@
     \includegraphics[width=.9\linewidth]{figures/memo_scenar_2.png}
-    \caption{Memorability score for events in memory. Events mentioned in the text are highlighted.}
+    \caption{Temperature scenario. Memorability score for events in memory. Events mentioned in the text are highlighted.}
     \label{fig:result1}
@@
 
+% todo: reproduction of the ROC curve. see with the python code
 Given that we generated the data used for this experiment, it is possible to
 flag all perturbation events apart from the usual daily events and evaluate how a
-detection based on ``memorability'' score would perform in distinguishing these
+detection based on the ``memorability'' score would perform in distinguishing these
 events.
@@
     \includegraphics[width=0.7\linewidth]{./figures/roc}
-    \caption{Experimental ROC curve (True Positive Rate against False Positive
+    \caption{Temperature scenario. Experimental ROC curve (True Positive Rate against False Positive
         Rate) for a classifier which compares the memorability of events to a given threshold, which we vary to change the sensitivity of our detector. Measures consider 23
@@
 \subsection{Discussion: the subjectivity of predicates}
+% todo : voir une discussion sur la notion d'influence des événements similaires/la notion de voisinage
 \label{sec:discussion}
-For individuals, the notion of memorability and complexity of events is highly subjective: the same event may appear usual for a person, while being exceptional for her neighbor.
-Our approach to memorability embraces this subjectivity, incorporating it through the definition of predicates.
+For humans, the notion of memorability and event complexity is highly subjective: the same event may appear usual for a person, while being exceptional for another.
+Our approach to memorability aims to reproduce this subjectivity while providing a formal canvas for memorability computations.
+Subjectivity is incorporated through the notion of predicates and their complexity, as they reflect the perception a human has of her surrounding environment.
 
@@
     \end{subfigure}
-    \caption{The same memory of events, analyzed through three different sets of predicates.
+    \caption{TV scenario, variation.
+The same memory of events, analyzed through three different sets of predicates.
 Light events are in blue, TV events in salmon.
@@
 \ref{fig:pred_2} adds label predicates alongside with "dark" and day/night predicates.
-Figure \ref{fig:pred_3} adds the possibility to select directly select any event from the memory of size $N$, at the cost of $l(N)$ bits.}
+Figure \ref{fig:pred_3} adds the possibility to directly select any event from the memory of size $N$, at the cost of $l(N)$ bits.}
     \label{fig:3preds}
@@
 This added capacity has the immediate effect of setting a clear upper limit to the description complexity of items, since any item can be retrieved using $\L(N)$ bits (in this example, this limit is around 25 bits).
-While this kind of direct retrieval is trivial in computer science (one could use the memory address of any stored event), its correlation with human cognition is not obvious: can humans be considered to have this possibility of selecting any event from their entire memory, without restriction regarding their nature, their time of occurrence, their magnitude?
+While this kind of direct retrieval is trivial in computer science (one could use the memory address of any stored event), its correlation to human cognition is not obvious: can humans be considered to have this possibility of selecting any event from their entire memory, without restriction regarding their nature, their time of occurrence, their magnitude?
 However, this highlights an interesting phenomenon: using this direct retrieval, the module makes no distinction, complexity-wise, between events past a given threshold.
@@
 Our work is intended to be integrated into larger-scale frameworks to monitor
-and detect events in complex environment such as smart homes.
-Smart homes are often regarded as self-organizing systems \cite{kramer_rigorous_2009,kounev_notion_2017}.
+and detect events in complex environments such as smart homes.
+Smart homes are often regarded as self-organizing systems~\cite{kramer_rigorous_2009,kounev_notion_2017}.
 As such, they present capacities of adaptation to new goals,
-new components, new environment. A commonly used approach is the principle of
-autonomic system, which minimizes users' interventions for the management of the
-system \cite{kounev_notion_2017,kephart_vision_2003}.
+new components, new environment. A commonly framework is 
+autonomic computing which minimizes users' interventions for the management of the
+system~\cite{kounev_notion_2017,kephart_vision_2003}.
 
 In situations where more data is available, we could rely
-on correlation or causal inference from known relations \cite{peters_elements_2017,fadiga_or_2021}. Relations between inference and complexity have already been studied. The
+on correlation or causal inference from known relations~\cite{peters_elements_2017,fadiga_or_2021}. Relations between inference and complexity have already been studied. The
 case of inference was one of the motivations for R. Solomonoff to introduce his
-universal algorithmic probability \cite{solomonoff_formal_1964} as a tool to
+universal algorithmic probability~\cite{solomonoff_formal_1964} as a tool to
 reach an idealized inference machine, creating the notion of complexity simultaneously
@@
 
-The relation between complexity, compression and causality was used in \cite{tatti_finding_2008} to devise the PACK algorithm.
+The relation between complexity, compression and causality was used in~\cite{tatti_finding_2008} to devise the PACK algorithm.
 It models a dataset by using a family of decision trees where each tree describes how one variable can be expressed given the others.
@@
 description of the model and description of the errors), PACK compresses the dataset while finding relations between variables which can be further analyzed.
-More recently, \cite{marx_causal_2018} used Minimum Description Length to determine, given a joint probability distribution over $(X,Y)$, whether $X$ causes $Y$ or $Y$ causes $X$.
+More recently,~\cite{marx_causal_2018} used Minimum Description Length to determine, given a joint probability distribution over $(X,Y)$, whether $X$ causes $Y$ or $Y$ causes $X$.
 Their method is based on tree models and it implies that a model respecting the causal relation will be simpler to describe.
@@
 Some previous work can also be noted for having used AIT techniques to qualify and detect events in time series data.
-For instance, \cite{batista_complexity-invariant_2011,fadlallah_weighted-permutation_2013} propose weighted permutation entropy as a proxy for complexity measures in time series data, and use it to find relations between different time series.
-\cite{hu_discovering_2011} proposes a MDL approach to find the intrinsic dimensions of time series.
-
-The philosophy of our approach can be related to the ``Isolation forests'' method\cite{liu_isolation_2008,hariri_extended_2021}.
+For instance,~\cite{batista_complexity-invariant_2011,fadlallah_weighted-permutation_2013} propose weighted permutation entropy as a proxy for complexity measures in time series data, and use it to find relations between different time series.~\cite{hu_discovering_2011} proposes a MDL approach to find the intrinsic dimensions of time series.
+
+The philosophy of our approach can be related to the ``Isolation forests'' method~\cite{liu_isolation_2008,hariri_extended_2021}.
 It evaluates the isolation of data points by constructing random binary tree classifiers.
@@
 
-The practical application of the theoretical notions of memorability and relative memorability requires further developments. We highlight two of them which seem to us, to date, the most challenging.
+The practical application of the theoretical notions of event memorability $M(e)$ and relative memorability $M(c|e)$ requires further developments.
+We highlight two of them which seem to us, to date, the most challenging.
 
@@
 predefined predicate concepts, from which the different filters are constructed.
-As an extension, we suggest the possibility of exploring such predicates dynamically.
-One mya analyze discriminating
+As an extension, we suggest the possibility of defining such predicates dynamically.
+One may analyze discriminating
 dimensions of incoming data and create predicates to name these differences,
-similar to the contrast operations proposed in \cite{dessalles_conceptual_2015,
+similar to the contrast operations proposed in~\cite{dessalles_conceptual_2015,
     gardenfors2004conceptual}. For instance, the predicate concept $\mathtt{hot}$
 can be discovered by discriminating a recent hot day along the
-temperature axis and naming the difference with the prototypical day.
+temperature axis and naming the difference with the prototypical day\cite{pol_multi-level_2021}.
 
@@
 
-We proposed an approach to evaluate memorability as a difference between the expected description complexity of an event and its actual value.
-With our definition, something is memorable if it requires less information or more information than expected to be described.
+We proposed an approach to evaluate event memorability as a difference between the expected description complexity of an event and its actual value.
+With our definition, something is memorable if its description requires more information than expected; or less information than expected.
 To formalize this notion, we used principles of minimal description exposed in Algorithmic Information Theory.
 By defining filter operations from predicates and successively applying these filters, we defined formal ways of describing events, whose length can then be measured to evaluate their description length.
-From this, memorability is defined as the absolute difference between the average complexity of similar events and the actual description complexity of the event.
+From this, memorability is defined as the absolute difference between the average complexity of similar events (representing the expected complexity) and the actual description complexity of the event.
 
@@
 
-We provided an implementation algorithm to compute this measure on events and showed some of its results in two smart home examples.
-These application scenarios qualitatively show how our measure fares in comparison with the human notion of memorability, and how this measure can be used to propose relevant hypotheses in an abductive inference process without having further knowledge of the system.
-We discussed the inherent subjectivity of our approach by exhibiting the impact of the choice of predicates for complexity computation in a toy example.
+We provided an implementation algorithm to compute this measure on events and showed its application in two smart home examples.
+These scenarios qualitatively illustrate how our measure fares in comparison with the human notion of memorability, and how this measure can be used to propose relevant hypotheses in an abductive inference process without having further knowledge of the system.
+We discussed the inherent subjectivity of our approach by highlighting the impact of the choice of predicates for complexity computation in a toy example.
 We consider extending our approach by including online learning of predicates that would make our approach coincide with the subjectivity of the system's user.
@@
 For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
-% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
-% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
-328; \citeyear{ref-journal-3b}, p.475).
+% To cite two works by the same author:~\citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a},~\citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
+% To cite two works by the same author with specific pages:~\citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
+328;~\citeyear{ref-journal-3b}, p.475).
 This produces: Wong (1999, p.

### 2022-01-27 paper_HOUZE.tex
setting for publication

--- 2022-01-26 new_paper.tex
@@
 In contrast, human are able to quickly categorize some events as being more ``memorable'' than others.
-They do so wihtout relying on knowledge of the system's inner working or large previous datasets.
+They do so without relying on knowledge of the system's inner working or large previous datasets.
 Having this ability would allow the system to: i) identify and summarize a situation to the user by presenting only memorable events; ii) suggest the most memorable events as possible hypotheses in an abductive inference process.
@@
     \label{eq:similar}
-    N_{\pi, k}(e) = \{e'\in M_0, (e' \neq e) \wedge \pi_k(e')\}
+    N_{\pi, k}(e) = \{e'\in M_0, (e' \neq e) \wedge \pi_k(e')\}\,.
 \end{equation}
@@
 events.
-The result is presented as a ROC curve in Figure \ref{fig:roc}, obtained by varying the memorability threshold.
-This shows the memorability score's performance as a classifying tool for ``out-of-the norm'' events.
-In this example, memorability alone correctly identified 18 manually generated events with only 5 false-positives.
-While the direct application of memorability for classification or event detection is not within the scope of this paper (see Section \ref{sec:related} for complexity-based detection), this first result is on par with the motivation of memorability being in accordance with the intuitive notion.
+Even if event detection is not the main purpose of memorability, we conducted the following experiment: we identified 20 events from the ``Temperature scenario'' that were the result of hand-made perturbation in the smart home simulation (such as days where the outdoor temperature was set to a abnormally high temperature, device removal/addition) as ground-truth events.
+Then, we used a memorability-based detector (i.e.
+flagging all events with memorability above a given threshold as ``true'' events) and tested different threshold values to observe the True-Positive and False-Positive rates.
+The result of this experiment is presented as a ROC curve in Figure \ref{fig:roc} This illustrates the memorability score's performance as a classifying tool for ``out-of-the norm'' events.
+In this example, misclassification has been observed to come from different phenomena: i) recent events are memorable with our metrics, while not being ground truth events; ii) as events are defined on a daily basis, this classification may not be suiting for days-long events (e.g.
+events 329 and 339 correspond in fact to the same cold night generated in the data), which adds unnecessary information to their description and therefore hinder their memorability score.
+While the former is a consequence of how our memorability score considers recent events as particularly memorable (this can be understood as a desired feature for such a metrics), the latter comes from event detection and definition and could be improved by further developments.
 
